---
title: "Producer Theory"
subtitle: "EC 201: Principles of Microeconomics"
author: "Kyle Raze"
date: "Winter 2020"
output:
  xaringan::moon_reader:
    css: ['default', 'metropolis', 'metropolis-fonts', 'my-css.css']
    self_contained: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "https://platform.twitter.com/widgets.js"
---
class: inverse, middle

```{r Setup, include = F}
options(htmltools.dir.version = FALSE)
library(pacman)
p_load(ggthemes, readxl, viridis, knitr, dslabs, gapminder, extrafont, Ecdat, wooldridge, tidyverse, magrittr, janitor, kableExtra)
# Define colors
red_pink <- "#e64173"
met_slate <- "#272822" # metropolis font color 
purple <- "#9370DB"
green <- "#007935"
light_green <- "#7DBA97"
orange <- "#FD5F00"
turquoise <- "#44C1C4"
red <- "#b92e34"
# Notes directory
dir_slides <- "~/GitHub/Intro_Microeconomics_W20/Lectures/03-Consumer_Theory/"
# Knitr options
opts_chunk$set(
  comment = "#>",
  fig.align = "center",
  fig.height = 7,
  fig.width = 10.5,
  #dpi = 300,
  #cache = T,
  warning = F,
  message = F
)  
theme_simple <- theme_bw() + theme(
  axis.line = element_line(color = met_slate),
  panel.grid = element_blank(),
  rect = element_blank(),
  strip.text = element_blank(),
  text = element_text(family = "Fira Sans", color = met_slate, size = 17),
  # axis.text.x = element_text(size = 12),
  # axis.text.y = element_text(size = 12),
  axis.ticks = element_blank()
)
theme_market <- theme_bw() + theme(
  axis.line = element_line(color = met_slate),
  panel.grid = element_blank(),
  rect = element_blank(),
  strip.text = element_blank(),
  text = element_text(family = "Fira Sans", color = met_slate, size = 17),
  axis.title.x = element_text(hjust = 1, size = 17),
  axis.title.y = element_text(hjust = 1, angle = 0, size = 17),
  axis.ticks = element_blank()
)
theme_gif <- theme_bw() + theme(
  axis.line = element_line(color = met_slate),
  panel.grid = element_blank(),
  rect = element_blank(),
  text = element_text(family = "Fira Sans", color = met_slate, size = 17),
  axis.text.x = element_text(size = 12),
  axis.text.y = element_text(size = 12),
  axis.ticks = element_blank()
)
theme_goods <- theme_bw() + theme(
  axis.line = element_line(),
  panel.grid = element_blank(),
  rect = element_rect(size = 1),
  strip.text = element_blank(),
  text = element_text(size = 17, family = "Fira Sans"), 
  axis.text.x = element_text(size = 17),
  axis.text.y = element_text(size = 17, angle = 90, hjust = 0.5),
  axis.title.x = element_text(color = purple, size = 19),
  axis.title.y = element_text(color = red_pink, size = 19),
  axis.ticks = element_blank(),
  plot.title = element_blank(),
  legend.position = "none"
)
shift_axis <- function(p, y=0){
  g <- ggplotGrob(p)
  dummy <- data.frame(y=y)
  ax <- g[["grobs"]][g$layout$name == "axis-b"][[1]]
  p + annotation_custom(grid::grobTree(ax, vp = grid::viewport(y=1, height=sum(ax$height))), 
                        ymax=y, ymin=y) +
    geom_hline(aes(yintercept=y), data = dummy, size = 0.5, color = met_slate) +
    theme(axis.text.x = element_blank(), 
          axis.ticks.x = element_blank())
}
wrapper <- function(x, ...) paste(strwrap(x, ...), collapse = "\n")
```

# Prologue

---
# Housekeeping

**Midterm 2:** Grades posted.

- You should have received an email with your score, an approximate grade, and an attachment with your answer choices.

--

**Discussions**

- Upcoming worksheets important for final exam preparation.

--

**Optional Short Essays**

- One percentage point of extra credit per essay (up to four total).
- Submit on Canvas.
- Due by 10:00 on Monday of Week 10 (before lecture).
- See guidelines in the syllabus.


---
# Agenda

**Goal:** Understand the tradeoffs that producers face.

- How do producers make decisions?

- Where does the supply curve come from?

--

**Q:** Why should you care?

- **A:** Useful insights for entrepreneurs, managers, and policymakers.  

--

**Outline**

1. Costs of production (today).

2. Firms in competitive markets (Wednesday).

3. Monopoly (next week).

---
# Profit

**Assumption:** Firms seek to maximize profit.

- Profit .mono[=] total revenue .mono[-] total cost

--

**Q:** Do all businesses try to maximize profit?

- Publicly-traded firms?

- Family businesses?

- Non-profit organizations?

---
# Profit Maximization

To maximize profit, a producer must answer several interrelated questions:

- How much should I produce?

- What price should I set?

- How much labor should I hire? How much capital should I purchase?

- Should I shut down or keep producing?

---
# Profit Maximization

Answers to those questions depend on 

1. Production technology .mono[-->] cost structure.

2. Market structure .mono[-->] ability to set prices.

3. Short-run *vs.* long-run.

---
class: inverse, middle

# Costs of Production

---
# Costs

## Total cost .mono[=] .purple[explicit cost] .mono[+] .pink[implicit cost]

--

## .purple[Explicit cost]

> .purple[Tangible, out-of-pocket expenses.]

- *e.g.,* wages, rent, raw materials, maintenance, taxes, *etc.*

---
# Costs

## Total cost .mono[=] .purple[explicit cost] .mono[+] .pink[implicit cost]

## .pink[Implicit cost]

> .pink[Opportunity cost of already-owned resources.]

- Forgone value of the next-best investment <br> $\quad$ .mono[-->] no out-of-pocket payments.

- *e.g.,* starting a different business, investing in the stock market, *etc.*

---
# Costs

## Example

Need $100,000 to start a business and the interest rate is 5%.

**Option 1:** Borrow $100,000.

- Explicit cost .mono[=] $5,000.

**Option 2:** Use $40,000 from savings and borrow $60,000.

- Explicit cost .mono[=] $3,000 .mono[-->] 5% interest on the loan.

- Implicit cost .mono[=] $2,000 .mono[-->] 5% forgone interest on the $40,000 from savings.

--

.pink[Total cost .mono[=] $5,000 for either option.]

---
# Accounting *vs.* Economic Profits

## Accounting profit <br> $\quad$ .mono[=] total revenue .mono[-] .pink[explicit cost]

--

## Economic profit <br> $\quad$ .mono[=] total revenue .mono[-] (.pink[explicit cost] .mono[+] .purple[implicit cost])

--

<br>

**Q:** Which of the following is true?

> **A.** Accounting profit .mono[>] economic profit.

> **B.** Accounting profit .mono[<] economic profit.

---
count: false
# Accounting *vs.* Economic Profits

## Accounting profit <br> $\quad$ .mono[=] total revenue .mono[-] .pink[explicit cost]

## Economic profit <br> $\quad$ .mono[=] total revenue .mono[-] (.pink[explicit cost] .mono[+] .purple[implicit cost])

<br>

**Q:** Which of the following is true?

> .pink[**A.** Accounting profit .mono[>] economic profit.]

> **B.** Accounting profit .mono[<] economic profit.

---
# Accounting *vs.* Economic Profits

Positive accounting profit doesn't always mean that a firm is doing well.

1. Economic profit .mono[<] 0 <br> $\quad$ .mono[==>] Current business is less profitable than the next-best option.

2. Economic profit .mono[=] 0 <br> $\quad$ .mono[==>] Current business is just as profitable as the next-best option.

3. Economic profit .mono[>] 0 <br> $\quad$ .mono[==>] Current business is more profitable than the next-best option.

---
# Production Technology

**Output:** The product that a firm produces.

**Inputs:** Resources used in a firm's production process to produce output.

- Land: natural resources, site of business, *etc.*

- Labor: workers, managers, *etc.*

- Capital: machinery, computers, vehicles, *etc.*

--

.pink[**Production function:** Mathematical description of the relationship between inputs and output in a firm's production process.]

.center[.pink[Inputs .mono[-->] production technology .mono[-->] output]] 

---
# Total Product

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- 0:10
total_product <- function(x) 3*x - 0.5*x^3 + 4*x^2 

ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 7.5), expand=c(0,0), breaks = seq(0,7,1)) +
  scale_y_continuous(limits = c(0, 65), expand=c(0,0), breaks = seq(0,60,10)) +
  theme_market +
  labs(x = "L", y = "Q") +
  stat_function(fun = total_product, color = purple, size = 1)
```
]

.less-right[
## Definition

> Total amount of output from a given quantity of input.

**Example:** Labor (L), holding other inputs constant.

]

---
count: false
# Total Product

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- 0:10
z <- seq(0, 2.67, 0.01)
w <- seq(2.67, 5.68, 0.01)
u <- seq(5.68, 7.5, 0.01)
total_product <- function(x) 3*x - 0.5*x^3 + 4*x^2 

ggplot() +
  scale_x_continuous(limits = c(0, 7.5), expand=c(0,0), breaks = seq(0,7,1)) +
  scale_y_continuous(limits = c(0, 65), expand=c(0,0), breaks = seq(0,60,10)) +
  theme_market +
  labs(x = "L", y = "Q") +
  stat_function(aes(x), fun = total_product, color = purple, size = 1) +
  geom_vline(xintercept = 2.67, color = met_slate, size = 1, linetype = "dashed") +
  geom_vline(xintercept = 5.68, color = met_slate, size = 1, linetype = "dashed") +
  annotate("label", label = wrapper("Increasing at Increasing Rate", 15), x = 2.67/2, y = 58.5, color = purple, family = "Fira Sans", size = 4) +
  annotate("label", label = wrapper("Increasing at Decreasing Rate", 15), x = (2.67 + 5.68)/2, y = 58.5, color = purple, family = "Fira Sans", size = 4) +
  annotate("label", label = wrapper("Decreasing", 15), x = (7.5 + 5.68)/2, y = 58.5, color = purple, family = "Fira Sans", size = 4) +
  geom_ribbon(aes(x = z, ymin = 0, ymax = 65), fill = purple, alpha = 0.2, linetype = "blank") + # PS
  geom_ribbon(aes(x = u, ymin = 0, ymax = 65), fill = purple, alpha = 0.2, linetype = "blank")
```
]

.less-right[
## Definition

> Total amount of output from a given quantity of input.

**Example:** Labor (L), holding other inputs constant.

]

---
# Marginal Product

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- 0:10
total_product <- function(x) 3*x - 0.5*x^3 + 4*x^2 
q_1 <- total_product(3)
q_2 <- total_product(4)

ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 7.5), expand=c(0,0), breaks = seq(0,7,1)) +
  scale_y_continuous(limits = c(0, 65), expand=c(0,0), breaks = seq(0,60,10)) +
  theme_market +
  labs(x = "L", y = "Q") +
  stat_function(fun = total_product, color = purple, size = 1) +
  geom_point(aes(x = 3, y = q_1), color = met_slate, size = 2) +
  geom_point(aes(x = 4, y = q_2), color = met_slate, size = 2) +
  geom_segment(aes(x = 3, y = 0, xend = 3, yend = q_1), linetype  = "dashed", color = met_slate, size = 1) +
  geom_segment(aes(x = 4, y = 0, xend = 4, yend = q_2), linetype  = "dashed", color = met_slate, size = 1) +
  geom_segment(aes(x = 0, y = q_1, xend = 3, yend = q_1), linetype  = "dashed", color = met_slate, size = 1) +
  geom_segment(aes(x = 0, y = q_2, xend = 4, yend = q_2), linetype  = "dashed", color = met_slate, size = 1) +
  geom_segment(aes(x = 3 + 0.1, y = 10, xend = 4 - 0.1, yend = 10), arrow = arrow(angle = 20, length = unit(0.4,"cm"), type = "closed"), color = met_slate, size = 1) +
  geom_segment(aes(x = 0.5, y = q_1 + 0.75, xend = 0.5, yend = q_2 - 0.75), arrow = arrow(angle = 20, length = unit(0.4,"cm"), type = "closed"), color = met_slate, size = 1) +
  annotate("text", label = expression(Delta*"Q"=="Marginal Product"), x = 2, y = (q_1 + q_2)/2, color = met_slate, family = "Fira Sans", size = 4) +
  annotate("text", label = expression(Delta*"L"==1), x = 3.5, y = 7.5, color = met_slate, family = "Fira Sans", size = 4)
```
]

.less-right[
## Definition

> Change in total product from a one-unit increase in an input.

**Example:** Labor (L), holding other inputs constant.

]

---
count: false
# Marginal Product

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- 0:10
marginal_product <- function(x) 3 + 8*x - 1.5*x^2

mp1 <- ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 7.5), expand=c(0,0), breaks = seq(1,7,1)) +
  scale_y_continuous(limits = c(-6, 16), expand=c(0,0), breaks = seq(-5,15,5)) +
  theme_market +
  #geom_hline(yintercept = 0, color = met_slate, size = 0.5) + 
  labs(x = "L", y = "MP") +
  stat_function(fun = marginal_product, color = purple, size = 1) +
  theme(axis.line.x = element_blank(),
        axis.title.x = element_text(hjust = 1, vjust = 21, size = 17)) 

shift_axis(mp1)
```
]

.less-right[
## Definition

> Change in total product from a one-unit increase in an input.

**Example:** Labor (L), holding other inputs constant.

]

---
count: false
# Marginal Product

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- 0:10
z <- seq(0, 2.67, 0.01)
w <- seq(2.67, 5.68, 0.01)
u <- seq(5.68, 7.5, 0.01)
marginal_product <- function(x) 3 + 8*x - 1.5*x^2

mp1 <- ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 7.5), expand=c(0,0), breaks = seq(1,7,1)) +
  scale_y_continuous(limits = c(-6, 16), expand=c(0,0), breaks = seq(-5,15,5)) +
  theme_market +
  #geom_hline(yintercept = 0, color = met_slate, size = 0.5) + 
  labs(x = "L", y = "MP") +
  stat_function(fun = marginal_product, color = purple, size = 1) +
  theme(axis.line.x = element_blank(),
        axis.title.x = element_text(hjust = 1, vjust = 21, size = 17)) +
  geom_vline(xintercept = 2.67, color = met_slate, size = 1, linetype = "dashed") +
  geom_vline(xintercept = 5.68, color = met_slate, size = 1, linetype = "dashed") +
  annotate("label", label = wrapper("Increasing at Increasing Rate", 22), x = 2.67/2 - 0.0, y = 14.75, color = purple, family = "Fira Sans", size = 4) +
  annotate("label", label = wrapper("Increasing at Decreasing Rate", 22), x = (2.67 + 5.68)/2 + 0.0, y = 14.75, color = purple, family = "Fira Sans", size = 4) +
  annotate("label", label = wrapper("Decreasing", 22), x = (7.5 + 5.68)/2, y = 14.75, color = purple, family = "Fira Sans", size = 4)

shift_axis(mp1)
```
]

.less-right[
## Definition

> Change in total product from a one-unit increase in an input.

**Example:** Labor (L), holding other inputs constant.

]

---
# Diminishing Marginal Product

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- 0:10
marginal_product <- function(x) 3 + 8*x - 1.5*x^2

mp1 <- ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 7.5), expand=c(0,0), breaks = seq(1,7,1)) +
  scale_y_continuous(limits = c(-6, 16), expand=c(0,0), breaks = seq(-5,15,5)) +
  theme_market +
  #geom_hline(yintercept = 0, color = met_slate, size = 0.5) + 
  labs(x = "L", y = "MP") +
  stat_function(fun = marginal_product, color = purple, size = 1) +
  theme(axis.line.x = element_blank(),
        axis.title.x = element_text(hjust = 1, vjust = 21, size = 17)) +
  geom_vline(xintercept = 2.67, color = met_slate, size = 1, linetype = "dashed") +
  annotate("label", label = wrapper("Increasing Marginal Product", 18), x = 2.67/2 - 0.0, y = 14.85, color = purple, family = "Fira Sans", size = 4) +
  annotate("label", label = wrapper("Decreasing Marginal Product", 18), x = (2.67 + 7.5)/2 + 0.0, y = 14.85, color = purple, family = "Fira Sans", size = 4)

shift_axis(mp1)
```
]

.less-right[
## Idea

.purple[Marginal product eventually decreases] as more of one input is used with a fixed quantity of other inputs.

]

---
class: clear-slide

**Q:** Suppose you are a watermelon farmer. If the price of a watermelon is $5 and the market wage is $2000 per month, how many workers should you hire?

```{R, echo = F, escape = F}
data.frame(
  v1 = c(0, 1, 2, 3, 4, 5),
  v2 = c(0, 1000, 1800, 2400, 2800, 3000),
  v3 = c("", "", "", "", "", ""),
  v4 = c("", "", "", "", "", ""),
  v5 = c("$2000", "$2000", "$2000", "$2000", "$2000", "$2000")
) %>% kable(
  escape = F,
  col.names = c("Workers (L)", "Watermelons (Q)", "Marginal Product", "Marginal Product &times; P", "Wage"),
  align = c("c", "c", "c", "c", "c")#,
) %>%
row_spec(1:6, color = met_slate) %>%
row_spec(1:6, extra_css = "line-height: 110%;")
```

---
count: false
class: clear-slide

**Q:** Suppose you are a watermelon farmer. If the price of a watermelon is $5 and the market wage is $2000 per month, how many workers should you hire?

```{R, echo = F, escape = F}
data.frame(
  v1 = c(0, 1, 2, 3, 4, 5),
  v2 = c(0, 1000, 1800, 2400, 2800, 3000),
  v3 = c("&mdash;", 1000, 800, 600, 400, 200),
  v4 = c("", "", "", "", "", ""),
  v5 = c("$2000", "$2000", "$2000", "$2000", "$2000", "$2000")
) %>% kable(
  escape = F,
  col.names = c("Workers (L)", "Watermelons (Q)", "Marginal Product", "Marginal Product &times; P", "Wage"),
  align = c("c", "c", "c", "c", "c")#,
) %>%
row_spec(1:6, color = met_slate) %>%
row_spec(1:6, extra_css = "line-height: 110%;")
```

---
count: false
class: clear-slide

**Q:** Suppose you are a watermelon farmer. If the price of a watermelon is $5 and the market wage is $2000 per month, how many workers should you hire?

```{R, echo = F, escape = F}
data.frame(
  v1 = c(0, 1, 2, 3, 4, 5),
  v2 = c(0, 1000, 1800, 2400, 2800, 3000),
  v3 = c("&mdash;", 1000, 800, 600, 400, 200),
  v4 = c("&mdash;", "$5000", "$4000", "$3000", "$2000", "$1000"),
  v5 = c("$2000", "$2000", "$2000", "$2000", "$2000", "$2000")
) %>% kable(
  escape = F,
  col.names = c("Workers (L)", "Watermelons (Q)", "Marginal Product", "Marginal Product &times; P", "Wage"),
  align = c("c", "c", "c", "c", "c")#,
) %>%
row_spec(1:6, color = met_slate) %>%
row_spec(1:6, extra_css = "line-height: 110%;")
```

--

.purple[**Think at the margin:**] Keep hiring as long as .purple[Marginal Product .mono[&times;] P] $\geq$ .purple[Wage].

--

**A:** .pink[You should hire 4 workers.]

---
class: clear-slide

**Input:** Land .mono[-->] must pay $1000 per month in rent, regardless of the number of watermelons grown.

**Input:** Labor .mono[-->] market wage is $2000 per month.

```{R, echo = F, escape = F}
data.frame(
  v1 = c(0, 1, 2, 3, 4, 5),
  v2 = c(0, 1000, 1800, 2400, 2800, 3000),
  v3 = c("", "", "", "", "", ""),
  v4 = c("", "", "", "", "", ""),
  v5 = c("", "", "", "", "", "")
) %>% kable(
  escape = F,
  col.names = c("Workers (L)", "Watermelons (Q)", "Cost of Land", "Cost of Labor", "Total Cost"),
  align = c("c", "c", "c", "c", "c")#,
) %>%
row_spec(1:6, color = met_slate) %>%
row_spec(1:6, extra_css = "line-height: 110%;")
```

---
count: false
class: clear-slide

**Input:** Land .mono[-->] must pay $1000 per month in rent, regardless of the number of watermelons grown.

**Input:** Labor .mono[-->] market wage is $2000 per month.

```{R, echo = F, escape = F}
data.frame(
  v1 = c(0, 1, 2, 3, 4, 5),
  v2 = c(0, 1000, 1800, 2400, 2800, 3000),
  v3 = c("$1000", "$1000", "$1000", "$1000", "$1000", "$1000"),
  v4 = c("", "", "", "", "", ""),
  v5 = c("", "", "", "", "", "")
) %>% kable(
  escape = F,
  col.names = c("Workers (L)", "Watermelons (Q)", "Cost of Land", "Cost of Labor", "Total Cost"),
  align = c("c", "c", "c", "c", "c")#,
) %>%
row_spec(1:6, color = met_slate) %>%
row_spec(1:6, extra_css = "line-height: 110%;")
```

---
count: false
class: clear-slide

**Input:** Land .mono[-->] must pay $1000 per month in rent, regardless of the number of watermelons grown.

**Input:** Labor .mono[-->] market wage is $2000 per month.

```{R, echo = F, escape = F}
data.frame(
  v1 = c(0, 1, 2, 3, 4, 5),
  v2 = c(0, 1000, 1800, 2400, 2800, 3000),
  v3 = c("$1000", "$1000", "$1000", "$1000", "$1000", "$1000"),
  v4 = c("$0", "$2000", "$4000", "$6000", "$8000", "$10000"),
  v5 = c("", "", "", "", "", "")
) %>% kable(
  escape = F,
  col.names = c("Workers (L)", "Watermelons (Q)", "Cost of Land", "Cost of Labor", "Total Cost"),
  align = c("c", "c", "c", "c", "c")#,
) %>%
row_spec(1:6, color = met_slate) %>%
row_spec(1:6, extra_css = "line-height: 110%;")
```

---
count: false
class: clear-slide

**Input:** Land .mono[-->] must pay $1000 per month in rent, regardless of the number of watermelons grown.

**Input:** Labor .mono[-->] market wage is $2000 per month.

```{R, echo = F, escape = F}
data.frame(
  v1 = c(0, 1, 2, 3, 4, 5),
  v2 = c(0, 1000, 1800, 2400, 2800, 3000),
  v3 = c("$1000", "$1000", "$1000", "$1000", "$1000", "$1000"),
  v4 = c("$0", "$2000", "$4000", "$6000", "$8000", "$10000"),
  v5 = c("$1000", "$3000", "$5000", "$7000", "$9000", "$11000")
) %>% kable(
  escape = F,
  col.names = c("Workers (L)", "Watermelons (Q)", "Cost of Land", "Cost of Labor", "Total Cost"),
  align = c("c", "c", "c", "c", "c")#,
) %>%
row_spec(1:6, color = met_slate) %>%
row_spec(1:6, extra_css = "line-height: 110%;")
```

---
class: clear-slide

**Input:** Land .mono[-->] must pay $1000 per month in rent, regardless of the number of watermelons grown.

**Input:** Labor .mono[-->] market wage is $2000 per month.

.center[.purple[**Total Cost Curve**]]

.pull-left[
```{R, echo = F, escape = F}
cost_data <- data.frame(
  q = c(0, 1000, 1800, 2400, 2800, 3000),
  tc = c("$1000", "$3000", "$5000", "$7000", "$9000", "$11000")
) 
cost_data %>% 
  kable(
  escape = F,
  col.names = c("Watermelons (Q)", "Total Cost"),
  align = c("c", "c")#,
) %>%
row_spec(1:6, color = met_slate) %>%
row_spec(1:6, extra_css = "line-height: 110%;")
```
]

.pull-right[
```{r, echo = FALSE, fig.height = 3.25, fig.width = 4.25, dev = "svg"}
cost_data <- data.frame(
  q = c(0, 1000, 1800, 2400, 2800, 3000),
  tc = c(1000, 3000, 5000, 7000, 9000, 11000)
) 
cost_data %>% 
  ggplot(aes(x = q, y = tc)) +
  scale_x_continuous(limits = c(0, 3500)) +
  scale_y_continuous(limits = c(0, 12000), expand=c(0, 0), breaks = scales::pretty_breaks()) +
  theme_market +
  labs(x = "Q", y = "TC") +
  geom_point(size = 2, color = purple) +
  geom_line(size = 1, color = purple)
```
]

---
# Total Cost

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- seq(0, 6, 0.1)
total_cost <- function(x) 10 + 10*x - 4*x^2 + x^3

ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 5.5), expand=c(0,0), breaks = scales::pretty_breaks()) +
  scale_y_continuous(limits = c(0, 65), expand=c(0,0), breaks = scales::pretty_breaks()) +
  theme_market +
  labs(x = "Q", y = "TC") +
  stat_function(fun = total_cost, color = purple, size = 1) 
```
]

.less-right[
## Origin story

Production function .mono[+] input prices <br> .mono[-->] .purple[**total cost curve**].

**Q:** Why is the total cos curve S-shaped?

**A:** .purple[Diminishing marginal product.]

]

---
# Production Decisions

**Q:** How does a firm decide how much output to produce?

- To increase profit, should it produce more or fewer units?

--

**A:** Think at the margin.

- If the cost of additional units is less than the revenue from selling them, then profits rise as production increases.

- If the cost of additional units exceeds the revenue from selling them, then profits decline as production increases.

---
# Marginal Cost

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- seq(0, 6, 0.1)
total_cost <- function(x) 10 + 10*x - 4*x^2 + x^3
tc_1 <- total_cost(3)
tc_2 <- total_cost(4)

ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 5.5), expand=c(0,0), breaks = scales::pretty_breaks()) +
  scale_y_continuous(limits = c(0, 65), expand=c(0,0), breaks = scales::pretty_breaks()) +
  theme_market +
  labs(x = "Q", y = "TC") +
  stat_function(fun = total_cost, color = purple, size = 1) +
  geom_point(aes(x = 3, y = tc_1), color = met_slate, size = 2) +
  geom_point(aes(x = 4, y = tc_2), color = met_slate, size = 2) +
  geom_segment(aes(x = 3, y = 0, xend = 3, yend = tc_1), linetype  = "dashed", color = met_slate, size = 1) +
  geom_segment(aes(x = 4, y = 0, xend = 4, yend = tc_2), linetype  = "dashed", color = met_slate, size = 1) +
  geom_segment(aes(x = 0, y = tc_1, xend = 3, yend = tc_1), linetype  = "dashed", color = met_slate, size = 1) +
  geom_segment(aes(x = 0, y = tc_2, xend = 4, yend = tc_2), linetype  = "dashed", color = met_slate, size = 1) +
  geom_segment(aes(x = 3 + 0.1, y = 10, xend = 4 - 0.1, yend = 10), arrow = arrow(angle = 20, length = unit(0.4,"cm"), type = "closed"), color = met_slate, size = 1) +
  geom_segment(aes(x = 0.5, y = tc_1 + 0.75, xend = 0.5, yend = tc_2 - 0.75), arrow = arrow(angle = 20, length = unit(0.4,"cm"), type = "closed"), color = met_slate, size = 1) +
  annotate("text", label = expression(Delta*"TC"=="Marginal Cost"), x = 2 - 0.5, y = (tc_1 + tc_2)/2, color = met_slate, family = "Fira Sans", size = 4) +
  annotate("text", label = expression(Delta*"Q"==1), x = 3.5, y = 7.5, color = met_slate, family = "Fira Sans", size = 4)
```
]

.less-right[
## Definition

> Change in total cost that arises from a one-unit increase in output.

]

---
count: false
# Marginal Cost

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- seq(0, 6, 0.1)
marginal_cost <- function(x) 10 - 8*x + 3*x^2

ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 5.5), expand=c(0,0), breaks = scales::pretty_breaks()) +
  scale_y_continuous(limits = c(0, 65), expand=c(0,0), breaks = scales::pretty_breaks()) +
  theme_market +
  labs(x = "Q", y = "MC") +
  stat_function(fun = marginal_cost, color = purple, size = 1) 
```
]

.less-right[
## Definition

> Change in total cost that arises from a one-unit increase in output.

]

---
# Fixed *vs.* Variable Costs

## Total cost .mono[=] .green[fixed costs] .mono[+] .pink[variable costs]

## .green[Fixed costs]

> .green[Costs that **do not vary** with the quantity of output produced.]

- *e.g.,* rent, cost of equipment, loan payments, *etc.*

- Do not contribute to marginal cost.

---
# Fixed *vs.* Variable Costs

## Total cost .mono[=] .green[fixed costs] .mono[+] .pink[variable costs]

## .pink[Variable costs]

> .pink[Costs that **vary** with the quantity of output produced.]

- *e.g.,* wages, cost of raw materials, *etc.*

- Contribute to marginal cost.

---
# Fixed *vs.* Variable Costs

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- seq(0, 6, 0.1)
total_cost <- function(x) 10 + 10*x - 4*x^2 + x^3
vc <- function(x) 10*x - 4*x^2 + x^3
fc <- function(x) 10

ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 5.5), expand=c(0,0), breaks = scales::pretty_breaks()) +
  scale_y_continuous(limits = c(0, 65), expand=c(0,0), breaks = scales::pretty_breaks()) +
  theme_market +
  labs(x = "Q", y = "$") +
  stat_function(fun = fc, color = green, size = 1) +
  stat_function(fun = total_cost, color = purple, size = 1) +
  stat_function(fun = vc, color = red_pink, size = 1) +
  annotate("text", label = "TC", x = 3.6, y = 50, color = purple, family = "Fira Sans", size = 9) +
  annotate("text", label = "VC", x = 4.75, y = 50, color = red_pink, family = "Fira Sans", size = 9) +
  annotate("text", label = "FC", x = 4.75, y = 13, color = green, family = "Fira Sans", size = 9)

```
]

.less-right[
## .green[Fixed costs (FC)]

> .green[Costs that **do not vary** with Q.]

## .pink[Variable costs (VC)]

> .pink[Costs that **vary** with Q.]

]

---
# Average Fixed Cost

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- seq(0, 6, 0.1)
total_cost <- function(x) 10 + 10*x - 4*x^2 + x^3
afc <- function(x) 10/x

ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 5.5), expand=c(0,0), breaks = scales::pretty_breaks()) +
  scale_y_continuous(limits = c(0, 65), expand=c(0,0), breaks = scales::pretty_breaks()) +
  theme_market +
  labs(x = "Q", y = "$") +
  stat_function(fun = afc, color = green, size = 1) +
  annotate("text", label = "AFC", x = 4.75, y = 5.5, color = green, family = "Fira Sans", size = 9)

```
]

.less-right[
## Definition

> AFC .mono[=] FC &divide; Q

Decreases as Q rises <br> .mono[-->] more output spread over same fixed cost.

]

---
# Average Variable Cost

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- seq(0, 6, 0.1)
total_cost <- function(x) 10 + 10*x - 4*x^2 + x^3
avc <- function(x) (10*x - 4*x^2 + x^3)/x

ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 5.5), expand=c(0,0), breaks = scales::pretty_breaks()) +
  scale_y_continuous(limits = c(0, 65), expand=c(0,0), breaks = scales::pretty_breaks()) +
  theme_market +
  labs(x = "Q", y = "$") +
  stat_function(fun = avc, color = red_pink, size = 1) +
  annotate("text", label = "AVC", x = 4.75, y = 18, color = red_pink, family = "Fira Sans", size = 9)

```
]

.less-right[
## Definition

> AVC .mono[=] VC &divide; Q

]

---
# Average Total Cost

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- seq(0, 6, 0.1)
total_cost <- function(x) 10 + 10*x - 4*x^2 + x^3
atc <- function(x) (10 + 10*x - 4*x^2 + x^3)/x

ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 5.5), expand=c(0,0), breaks = scales::pretty_breaks()) +
  scale_y_continuous(limits = c(0, 65), expand=c(0,0), breaks = scales::pretty_breaks()) +
  theme_market +
  labs(x = "Q", y = "$") +
  stat_function(fun = atc, color = purple, size = 1) +
  annotate("text", label = "ATC", x = 0.7, y = 50, color = purple, family = "Fira Sans", size = 9)
```
]

.less-right[
## Definition

> ATC .mono[=] TC &divide; Q <br> .mono[=] AFC .mono[+] AVC

]

---
count: false
# Average Total Cost

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- seq(0, 6, 0.1)
total_cost <- function(x) 10 + 10*x - 4*x^2 + x^3
atc <- function(x) (10 + 10*x - 4*x^2 + x^3)/x

ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 5.5), expand=c(0,0), breaks = scales::pretty_breaks()) +
  scale_y_continuous(limits = c(0, 65), expand=c(0,0), breaks = scales::pretty_breaks()) +
  theme_market +
  labs(x = "Q", y = "$") +
  stat_function(fun = atc, color = purple, size = 1) +
  annotate("text", label = "ATC", x = 0.7, y = 50, color = purple, family = "Fira Sans", size = 9) +
  geom_vline(xintercept = 2.69, color = met_slate, size = 1, linetype = "dashed") +
  annotate("label", label = "Efficient Scale", x = 2.69, y = 60, color = met_slate, family = "Fira Sans", size = 4)
```
]

.less-right[
## Definition

> ATC .mono[=] TC &divide; Q <br> .mono[=] AFC .mono[+] AVC

Efficient scale .mono[=] Q that minimizes ATC.
]

---
# Average Total Cost

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- seq(0, 6, 0.1)
total_cost <- function(x) 10 + 10*x - 4*x^2 + x^3
atc <- function(x) (10 + 10*x - 4*x^2 + x^3)/x
afc <- function(x) 10/x
avc <- function(x) (10*x - 4*x^2 + x^3)/x

ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 5.5), expand=c(0,0), breaks = scales::pretty_breaks()) +
  scale_y_continuous(limits = c(0, 65), expand=c(0,0), breaks = scales::pretty_breaks()) +
  theme_market +
  labs(x = "Q", y = "$") +
  stat_function(fun = atc, color = purple, size = 1) +
  annotate("text", label = "ATC", x = 0.7, y = 50, color = purple, family = "Fira Sans", size = 9) +
  geom_vline(xintercept = 2.69, color = met_slate, size = 1, linetype = "dashed") +
  stat_function(fun = afc, color = green, size = 1) +
  annotate("text", label = "AFC", x = 4.75, y = 5.5, color = green, family = "Fira Sans", size = 9) +
  stat_function(fun = avc, color = red_pink, size = 1) +
  annotate("text", label = "AVC", x = 0.4, y = 5, color = red_pink, family = "Fira Sans", size = 9) +
  annotate("label", label = "Efficient Scale", x = 2.69, y = 60, color = met_slate, family = "Fira Sans", size = 4)
```
]

.less-right[

Efficient scale .mono[=] Q that minimizes ATC.

Q .mono[<] efficient scale: Falling AFC pulls ATC down.

Q .mono[>] efficient scale: Rising AVC pulls ATC up.
]

---
# Average Total Cost .mono[+] Marginal Cost

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- seq(0, 6, 0.1)
total_cost <- function(x) 10 + 10*x - 4*x^2 + x^3
atc <- function(x) (10 + 10*x - 4*x^2 + x^3)/x
marginal_cost <- function(x) 10 - 8*x + 3*x^2
avc <- function(x) (10*x - 4*x^2 + x^3)/x

ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 5.5), expand=c(0,0), breaks = scales::pretty_breaks()) +
  scale_y_continuous(limits = c(0, 65), expand=c(0,0), breaks = scales::pretty_breaks()) +
  theme_market +
  labs(x = "Q", y = "$") +
  stat_function(fun = atc, color = purple, size = 1) +
  annotate("text", label = "ATC", x = 0.7, y = 50, color = purple, family = "Fira Sans", size = 9) +
  stat_function(fun = avc, color = red_pink, size = 1, alpha = 0.5) +
  annotate("text", label = "AVC", x = 4.8, y = 10, color = red_pink, family = "Fira Sans", size = 9, alpha = 0.5) +
  stat_function(fun = marginal_cost, color = red_pink, size = 1) +
  annotate("text", label = "MC", x = 4.8, y = 50, color = red_pink, family = "Fira Sans", size = 9)
```
]

.less-right[
MC .mono[<] ATC .mono[==>] ATC is falling.

MC .mono[>] ATC .mono[==>] ATC is rising.

MC crosses ATC at the Q that minimizes ATC.

]

---
count: false
# Average Total Cost .mono[+] Marginal Cost

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- seq(0, 6, 0.1)
total_cost <- function(x) 10 + 10*x - 4*x^2 + x^3
atc <- function(x) (10 + 10*x - 4*x^2 + x^3)/x
marginal_cost <- function(x) 10 - 8*x + 3*x^2
avc <- function(x) (10*x - 4*x^2 + x^3)/x

ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 5.5), expand=c(0,0), breaks = scales::pretty_breaks()) +
  scale_y_continuous(limits = c(0, 65), expand=c(0,0), breaks = scales::pretty_breaks()) +
  theme_market +
  labs(x = "Q", y = "$") +
  stat_function(fun = atc, color = purple, size = 1) +
  annotate("text", label = "ATC", x = 0.7, y = 50, color = purple, family = "Fira Sans", size = 9) +
  stat_function(fun = marginal_cost, color = red_pink, size = 1) +
  annotate("text", label = "MC", x = 4.8, y = 50, color = red_pink, family = "Fira Sans", size = 9)
```
]

.less-right[
MC .mono[<] ATC .mono[==>] ATC is falling.

MC .mono[>] ATC .mono[==>] ATC is rising.

MC crosses ATC at the Q that minimizes ATC.

]

---
# Short-Run *vs.* Long-Run

## Short-Run

Some inputs are fixed.

- *e.g.,* land, factories, assembly-line robots, *etc.*

- Fixed inputs .mono[-->] fixed cost.

--

## Long-Run

All inputs are variable

- *e.g.,* firms can build new factories or sell existing ones.

- Firms use most efficient mix of inputs for any Q (*e.g.,* factory with lowest ATC).

---
# Long-Run ATC

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
cost_data <- tibble(
  q = seq(0, 6, 0.01),
  atc_s = 18 - 8*q + 3*q^2,
  atc_m = 30- 16*q + 3*q^2,
  atc_l = 60 - 24*q + 3*q^2
) %>% 
  mutate(atc = case_when(atc_s <= atc_m & atc_s <= atc_l ~ atc_s,
                         atc_m <= atc_s & atc_m <= atc_l ~ atc_m,
                         atc_l <= atc_s & atc_l <= atc_m ~ atc_l))

cost_data %>% 
  ggplot() +
  scale_x_continuous(limits = c(0, 5.5), expand=c(0,0), breaks = scales::pretty_breaks()) +
  scale_y_continuous(limits = c(0, 27), expand=c(0,0), breaks = scales::pretty_breaks()) +
  geom_line(aes(x = q, y = atc_s), color = red_pink, size = 1, linetype = "solid") +
  geom_line(aes(x = q, y = atc_m), color = green, size = 1, linetype = "solid") +
  geom_line(aes(x = q, y = atc_l), color = orange, size = 1, linetype = "solid") +
  theme_market +
  annotate("text", label = expression(ATC[S]), x = 0.7, y = 16, color = red_pink, family = "Fira Sans", size = 6) +
  annotate("text", label = expression(ATC[M]), x = 2.7, y = 10, color = green, family = "Fira Sans", size = 6) +
  annotate("text", label = expression(ATC[L]), x = 4.95, y = 18, color = orange, family = "Fira Sans", size = 6) +
  labs(x = "Q", y = "$")
```
]

.less-right[

A firm can choose from three factory sizes: S, M, L.  

Each factory size has its own short-run ATC curve. 

The firm can change to a different factory size in the long run, but not in the short run.

]

---
# Long-Run ATC

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
cost_data %>% 
  ggplot() +
  scale_x_continuous(limits = c(0, 5.5), expand=c(0,0), breaks = scales::pretty_breaks()) +
  scale_y_continuous(limits = c(0, 27), expand=c(0,0), breaks = scales::pretty_breaks()) +
  geom_line(aes(x = q, y = atc_s), color = red_pink, size = 1, linetype = "solid") +
  geom_line(aes(x = q, y = atc_m), color = green, size = 1, linetype = "solid") +
  geom_line(aes(x = q, y = atc_l), color = orange, size = 1, linetype = "solid") +
  geom_line(aes(x = q, y = atc), color = purple, size = 1.5, linetype = "solid") +
  theme_market +
  annotate("text", label = expression(ATC[S]), x = 0.7, y = 16, color = red_pink, family = "Fira Sans", size = 6) +
  annotate("text", label = expression(ATC[M]), x = 2.7, y = 10, color = green, family = "Fira Sans", size = 6) +
  annotate("text", label = expression(LRATC), x = 4, y = 8.5, color = purple, family = "Fira Sans", size = 9) +
  annotate("text", label = expression(ATC[L]), x = 4.95, y = 18, color = orange, family = "Fira Sans", size = 6) +
  labs(x = "Q", y = "$")
```
]

.less-right[

In the long run, a firm will choose the factory size that minimizes long-run ATC given the production level.

]

---
# Long-Run ATC

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- seq(0, 6, 0.1)
atc <- function(x) 10 - 2.75*x + 0.5*x^2

ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 5.5), expand=c(0,0), breaks = scales::pretty_breaks()) +
  scale_y_continuous(limits = c(0, 11), expand=c(0,0), breaks = scales::pretty_breaks()) +
  theme_market +
  labs(x = "Q", y = "$") +
  stat_function(fun = atc, color = purple, size = 1) +
  annotate("text", label = "LRATC", x = 4.5, y = 6.4, color = purple, family = "Fira Sans", size = 9)
```
]

.less-right[
Typical long-run average total cost curve.

- Traced out by numerous short-run ATC curves.

However, LRATC can take a variety of forms.

- Depends on returns to scale.

]

---
# Returns to Scale

## Economies of scale

> Long-run ATC falls as the quantity of output rises.

- Increased specialization among workers.

- More common when Q is low.

- *e.g.,* tech start-ups?

---
# Returns to Scale

## Constant returns to scale

> Long-run ATC stays the same as the quantity of output rises.

- *e.g.,* restaurant chains?

---
# Returns to Scale

## Diseconomies of scale

> Long-run ATC rises as the quantity of output rises.

- Coordination problems in large firms.

- Failure to control costs.

- More common when Q is high.

- *e.g.,* hospitals?



