---
title: "Monopoly & Antitrust"
subtitle: "EC 201: Principles of Microeconomics"
author: "Kyle Raze"
date: "Winter 2020"
output:
  xaringan::moon_reader:
    css: ['default', 'metropolis', 'metropolis-fonts', 'my-css.css']
    self_contained: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "https://platform.twitter.com/widgets.js"
---
class: inverse, middle

```{r Setup, include = F}
options(htmltools.dir.version = FALSE)
library(pacman)
p_load(ggthemes, readxl, viridis, knitr, dslabs, gapminder, extrafont, Ecdat, wooldridge, tidyverse, magrittr, janitor, kableExtra)
# Define colors
red_pink <- "#e64173"
met_slate <- "#272822" # metropolis font color 
purple <- "#9370DB"
green <- "#007935"
light_green <- "#7DBA97"
orange <- "#FD5F00"
turquoise <- "#44C1C4"
red <- "#b92e34"
# Notes directory
dir_slides <- "~/GitHub/Intro_Microeconomics_W20/Lectures/03-Consumer_Theory/"
# Knitr options
opts_chunk$set(
  comment = "#>",
  fig.align = "center",
  fig.height = 7,
  fig.width = 10.5,
  #dpi = 300,
  #cache = T,
  warning = F,
  message = F
)  
theme_simple <- theme_bw() + theme(
  axis.line = element_line(color = met_slate),
  panel.grid = element_blank(),
  rect = element_blank(),
  strip.text = element_blank(),
  text = element_text(family = "Fira Sans", color = met_slate, size = 17),
  # axis.text.x = element_text(size = 12),
  # axis.text.y = element_text(size = 12),
  axis.ticks = element_blank()
)
theme_market <- theme_bw() + theme(
  axis.line = element_line(color = met_slate),
  panel.grid = element_blank(),
  rect = element_blank(),
  strip.text = element_blank(),
  text = element_text(family = "Fira Sans", color = met_slate, size = 17),
  axis.title.x = element_text(hjust = 1, size = 17),
  axis.title.y = element_text(hjust = 1, angle = 0, size = 17),
  axis.ticks = element_blank()
)
theme_gif <- theme_bw() + theme(
  axis.line = element_line(color = met_slate),
  panel.grid = element_blank(),
  rect = element_blank(),
  text = element_text(family = "Fira Sans", color = met_slate, size = 17),
  axis.text.x = element_text(size = 12),
  axis.text.y = element_text(size = 12),
  axis.ticks = element_blank()
)
theme_goods <- theme_bw() + theme(
  axis.line = element_line(),
  panel.grid = element_blank(),
  rect = element_rect(size = 1),
  strip.text = element_blank(),
  text = element_text(size = 17, family = "Fira Sans"), 
  axis.text.x = element_text(size = 17),
  axis.text.y = element_text(size = 17, angle = 90, hjust = 0.5),
  axis.title.x = element_text(color = purple, size = 19),
  axis.title.y = element_text(color = red_pink, size = 19),
  axis.ticks = element_blank(),
  plot.title = element_blank(),
  legend.position = "none"
)
theme_game <- theme_bw() + theme(
  axis.line = element_line(),
  panel.grid = element_blank(),
  rect = element_rect(size = 1),
  strip.text = element_blank(),
  text = element_text(size = 20, family = "Fira Sans"), 
  axis.text.x = element_text(size = 20, color = met_slate),
  axis.text.y = element_text(size = 20, angle = 90, color = met_slate, hjust = 0.5),
  axis.title.x = element_text(color = met_slate, size = 20),
  axis.title.y = element_text(color = met_slate, size = 20),
  axis.ticks = element_blank(),
  plot.title = element_blank(),
  legend.position = "none"
)
shift_axis <- function(p, y=0){
  g <- ggplotGrob(p)
  dummy <- data.frame(y=y)
  ax <- g[["grobs"]][g$layout$name == "axis-b"][[1]]
  p + annotation_custom(grid::grobTree(ax, vp = grid::viewport(y=1, height=sum(ax$height))), 
                        ymax=y, ymin=y) +
    geom_hline(aes(yintercept=y), data = dummy, size = 0.5, color = met_slate) +
    theme(axis.text.x = element_blank(), 
          axis.ticks.x = element_blank())
}
wrapper <- function(x, ...) paste(strwrap(x, ...), collapse = "\n")
```

# Prologue

---
# Housekeeping

**Final exam:** Next week!

- Details next lecture.

--

**Canvas:** New practice problems .mono[+] discussion worksheets.

--

**Wednesday lecture:** Final exam review.

- Come prepared with questions!
- Syllabus update: no podcast.

--

**Friday discussion:** Monopoly worksheet .mono[+] practice problems.

---
# Market Failure

## Causes

1. Absence of property rights.

    - Externalities.
    
    - Public goods.

2. Market power.

    - *e.g.,* monopoly (today).

3. Asymmetric information.

---
class: inverse, middle

# Monopoly

---
# Monopoly

## Conditions

1. One seller.

2. Unique product without close substitutes.

3. Barriers to entry.

--

## Result

**Market power:** A monopolist has the ability to influence market prices.

- A monopolist is a **price maker**.

---
# Monopoly

## Examples?

**Q:** Is Google a monopolist? .purple[Why] or .pink[why not]?

--

- .purple[Most popular search engine by far .mono[+] huge, vertically integrated company .mono[+] ability to purchase potential new competitors.]

- .pink[Competes with Apple, Amazon, and Microsoft.]

--

**A:** .pink[Not a monopolist, but has market power.]

---
# Monopoly

## Examples?

**Q:** Is Comcast a monopolist? .purple[Why] or .pink[why not]?

--

- .purple[Sole internet provider in many areas .mono[+] huge, vertically integrated company .mono[+] high barriers to entry.]

- .pink[Competes with Verizon in some areas.]

--

**A:** .purple[Depending on where you live, yes.]

---
# Monopoly

## Examples?

**Q:** Is the only hardware store in a small town a monopolist? .purple[Why] or .pink[why not]?

--

- .purple[No direct competitors in town .mono[+] small size of market deters entrants.]

- .pink[A grocery store might carry some of the same products.]

--

**A:** .purple[Yes, provided that there sufficiently few indirect competitors.]

---
# Sources of Monopoly Power

## Exclusive control of resources

A firm can keep out potential competitors by buying up essential resources for production.

--

**Examples**  

- Early 1900s: Aluminum manufacturer ALCOA owned 90% of the global bauxite supply.

- De Beers owned most of the world's raw diamonds until the mid-2000s.

---
# Sources of Monopoly Power

## Difficulty raising capital

Incumbent monopolists are often large .mono[-->] new competitors would need a lot of money to compete effectively!

- Chances of competing against entrenched monopolist are low .mono[-->] risky investment for lenders.

--

**Example:** Operating systems.

- Supplanting Windows 10 as the leading operating system for PCs would require vast amounts of capital.

- A lender would pick Microsoft over your start-up to develop the next big operating system.

---
# Sources of Monopoly Power

## Economies of scale

Some industries feature immense upfront fixed costs, but low marginal costs thereafter.

- Tendency for consolidation over time.

--

**Examples:** Natural monopolies.

- Electricity and water.

- Cable internet and television.

- Railroads.

---
# Sources of Monopoly Power

## Licensing

Governments establish monopolies with licensing requirements.

- Licensing requirement .mono[=] legal barrier to entry.

**Rationale:** Minimize negative externalities or exploit economies of scale.

--

**Examples**

- Trash collection.

- Taxi medallions.

- Occupational licenses.

---
# Sources of Monopoly Power

## Patents and copyright law

Governments issue exclusive rights to sell a particular good or service for a fixed amount of time.

- Exclusive rights .mono[-->] monopoly.

**Tradeoff:** Market power *vs.* innovation.

--

**Examples**

- New prescription drugs.

- Books and music.

---
# Monopolists *vs.* Competitive Firms

Consumers of a particular firm's product have .pink[fewer alternatives] in a monopolistic market than in a perfectly competitive market.

.pull-left[
.center[**Monopolist**]
```{r, echo = FALSE, fig.height = 3.75, fig.width = 3.75, dev = "svg"}
demand <- function(x) 8 - 0.6*x
inv_demand <- function(x) (5/3)*(8 - x)
p_1 <- 7
p_2 <- 6
q_1 <- inv_demand(p_1) 
q_2 <- inv_demand(p_2)
ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = NULL) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = NULL) +
  theme_market +
  labs(x = "Q", y = "P") +
  stat_function(fun = demand, color = red_pink, size = 1) +
  annotate("text", label = "D", x = 10, y = 2.65, color = red_pink, family = "Fira Sans", size = 9) 
```
]
.pull-right[
.center[**Perfectly Competitive Firm**]
```{r, echo = FALSE, fig.height = 3.75, fig.width = 3.75,  dev = "svg"}
ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = NULL) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = NULL) +
  theme_market +
  labs(x = "Q", y = "P") +
  geom_hline(yintercept = 5, color = red_pink, size = 1) +
  annotate("text", label = "D", x = 10, y = 5.5, color = red_pink, family = "Fira Sans", size = 9)
```
]

---
# Marginal Revenue

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
demand <- function(x) 10 - x
inv_demand <- function(x) 10 - x
p_1 <- 7
p_2 <- 9
q_1 <- inv_demand(p_1) 
q_2 <- inv_demand(p_2)
ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  labs(x = "Q", y = "P") +
  # geom_rect(aes(xmin = 0, xmax = q_1, ymin = 0, ymax = p_1), alpha = 0.2, fill = green) +
  geom_rect(aes(xmin = 0, xmax = q_2, ymin = 0, ymax = p_2), alpha = 0.2, fill = green) +
  stat_function(fun = demand, color = red_pink, size = 1) +
  # geom_point(x = q_1, y = p_1, color = met_slate, size = 2) +
  geom_point(x = q_2, y = p_2, color = met_slate, size = 2) +
  # geom_segment(x = 0, xend = q_1, y = p_1, yend = p_1, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = 0, xend = q_2, y = p_2, yend = p_2, color = met_slate, linetype = "dashed", size = 1) +
  # geom_segment(x = q_1, xend = q_1, y = 0, yend = p_1, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = q_2, xend = q_2, y = 0, yend = p_2, color = met_slate, linetype = "dashed", size = 1) +
  annotate("text", label = "D", x = 10, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) 
```
]

.less-right[

.pink[**Price effect:**] As price decreases, existing customers pay less.

.purple[**Output effect:**] As price decreases, new customers purchase goods.

]

---
count: false
# Marginal Revenue

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
demand <- function(x) 10 - x
inv_demand <- function(x) 10 - x
p_1 <- 7
p_2 <- 9
q_1 <- inv_demand(p_1) 
q_2 <- inv_demand(p_2)
ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  labs(x = "Q", y = "P") +
  geom_rect(aes(xmin = 0, xmax = q_1, ymin = 0, ymax = p_1), alpha = 0.2, fill = green) +
  # geom_rect(aes(xmin = 0, xmax = q_2, ymin = 0, ymax = p_2), alpha = 0.2, fill = green) +
  stat_function(fun = demand, color = red_pink, size = 1) +
  geom_point(x = q_1, y = p_1, color = met_slate, size = 2) +
  geom_point(x = q_2, y = p_2, color = met_slate, size = 2) +
  geom_segment(x = 0, xend = q_1, y = p_1, yend = p_1, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = 0, xend = q_2, y = p_2, yend = p_2, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = q_1, xend = q_1, y = 0, yend = p_1, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = q_2, xend = q_2, y = 0, yend = p_2, color = met_slate, linetype = "dashed", size = 1) +
  annotate("text", label = "D", x = 10, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) 
```
]

.less-right[

.pink[**Price effect:**] As price decreases, existing customers pay less.

.purple[**Output effect:**] As price decreases, new customers purchase goods.

.purple[Output effect] <br> $\quad$ .mono[>] .pink[price effect] .mono[==>] <br> price cut .green[increases revenue].

]

---
count: false
# Marginal Revenue

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
demand <- function(x) 10 - x
inv_demand <- function(x) 10 - x
p_1 <- 7
p_2 <- 9
q_1 <- inv_demand(p_1) 
q_2 <- inv_demand(p_2)
ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  labs(x = "Q", y = "P") +
  geom_rect(aes(xmin = 0, xmax = q_2, ymin = p_1, ymax = p_2), alpha = 0.2, fill = red_pink) +
  geom_rect(aes(xmin = q_1, xmax = q_2, ymin = 0, ymax = p_1), alpha = 0.2, fill = purple) +
  stat_function(fun = demand, color = red_pink, size = 1) +
  geom_point(x = q_1, y = p_1, color = met_slate, size = 2) +
  geom_point(x = q_2, y = p_2, color = met_slate, size = 2) +
  geom_segment(x = 0, xend = q_1, y = p_1, yend = p_1, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = 0, xend = q_2, y = p_2, yend = p_2, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = q_1, xend = q_1, y = 0, yend = p_1, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = q_2, xend = q_2, y = 0, yend = p_2, color = met_slate, linetype = "dashed", size = 1) +
  annotate("text", label = "D", x = 10, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) 
```
]

.less-right[

.pink[**Price effect:**] As price decreases, existing customers pay less.

.purple[**Output effect:**] As price decreases, new customers purchase goods.

.purple[Output effect] <br> $\quad$ .mono[>] .pink[price effect] .mono[==>] <br> price cut .green[increases revenue].

]

---
# Marginal Revenue

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
demand <- function(x) 10 - x
inv_demand <- function(x) 10 - x
p_1 <- 4
p_2 <- 6
q_1 <- inv_demand(p_1) 
q_2 <- inv_demand(p_2)
ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  labs(x = "Q", y = "P") +
  # geom_rect(aes(xmin = 0, xmax = q_1, ymin = 0, ymax = p_1), alpha = 0.2, fill = green) +
  geom_rect(aes(xmin = 0, xmax = q_2, ymin = 0, ymax = p_2), alpha = 0.2, fill = green) +
  stat_function(fun = demand, color = red_pink, size = 1) +
  # geom_point(x = q_1, y = p_1, color = met_slate, size = 2) +
  geom_point(x = q_2, y = p_2, color = met_slate, size = 2) +
  # geom_segment(x = 0, xend = q_1, y = p_1, yend = p_1, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = 0, xend = q_2, y = p_2, yend = p_2, color = met_slate, linetype = "dashed", size = 1) +
  # geom_segment(x = q_1, xend = q_1, y = 0, yend = p_1, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = q_2, xend = q_2, y = 0, yend = p_2, color = met_slate, linetype = "dashed", size = 1) +
  annotate("text", label = "D", x = 10, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) 
```
]

.less-right[

.pink[**Price effect:**] As price decreases, existing customers pay less.

.purple[**Output effect:**] As price decreases, new customers purchase goods.

]

---
count: false
# Marginal Revenue

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
demand <- function(x) 10 - x
inv_demand <- function(x) 10 - x
p_1 <- 4
p_2 <- 6
q_1 <- inv_demand(p_1) 
q_2 <- inv_demand(p_2)
ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  labs(x = "Q", y = "P") +
  geom_rect(aes(xmin = 0, xmax = q_1, ymin = 0, ymax = p_1), alpha = 0.2, fill = green) +
  # geom_rect(aes(xmin = 0, xmax = q_2, ymin = 0, ymax = p_2), alpha = 0.2, fill = green) +
  stat_function(fun = demand, color = red_pink, size = 1) +
  geom_point(x = q_1, y = p_1, color = met_slate, size = 2) +
  geom_point(x = q_2, y = p_2, color = met_slate, size = 2) +
  geom_segment(x = 0, xend = q_1, y = p_1, yend = p_1, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = 0, xend = q_2, y = p_2, yend = p_2, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = q_1, xend = q_1, y = 0, yend = p_1, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = q_2, xend = q_2, y = 0, yend = p_2, color = met_slate, linetype = "dashed", size = 1) +
  annotate("text", label = "D", x = 10, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) 
```
]

.less-right[

.pink[**Price effect:**] As price decreases, existing customers pay less.

.purple[**Output effect:**] As price decreases, new customers purchase goods.

.purple[Output effect] <br> $\quad$ .mono[=] .pink[price effect] .mono[==>] <br> price cut .green[does not change revenue].

]

---
count: false
# Marginal Revenue

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
demand <- function(x) 10 - x
inv_demand <- function(x) 10 - x
p_1 <- 4
p_2 <- 6
q_1 <- inv_demand(p_1) 
q_2 <- inv_demand(p_2)
ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  labs(x = "Q", y = "P") +
  geom_rect(aes(xmin = 0, xmax = q_2, ymin = p_1, ymax = p_2), alpha = 0.2, fill = red_pink) +
  geom_rect(aes(xmin = q_1, xmax = q_2, ymin = 0, ymax = p_1), alpha = 0.2, fill = purple) +
  stat_function(fun = demand, color = red_pink, size = 1) +
  geom_point(x = q_1, y = p_1, color = met_slate, size = 2) +
  geom_point(x = q_2, y = p_2, color = met_slate, size = 2) +
  geom_segment(x = 0, xend = q_1, y = p_1, yend = p_1, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = 0, xend = q_2, y = p_2, yend = p_2, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = q_1, xend = q_1, y = 0, yend = p_1, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = q_2, xend = q_2, y = 0, yend = p_2, color = met_slate, linetype = "dashed", size = 1) +
  annotate("text", label = "D", x = 10, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) 
```
]

.less-right[

.pink[**Price effect:**] As price decreases, existing customers pay less.

.purple[**Output effect:**] As price decreases, new customers purchase goods.

.purple[Output effect] <br> $\quad$ .mono[=] .pink[price effect] .mono[==>] <br> price cut .green[does not change revenue].

]

---
# Marginal Revenue

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
demand <- function(x) 10 - x
inv_demand <- function(x) 10 - x
p_1 <- 1
p_2 <- 3
q_1 <- inv_demand(p_1) 
q_2 <- inv_demand(p_2)
ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  labs(x = "Q", y = "P") +
  # geom_rect(aes(xmin = 0, xmax = q_1, ymin = 0, ymax = p_1), alpha = 0.2, fill = green) +
  geom_rect(aes(xmin = 0, xmax = q_2, ymin = 0, ymax = p_2), alpha = 0.2, fill = green) +
  stat_function(fun = demand, color = red_pink, size = 1) +
  # geom_point(x = q_1, y = p_1, color = met_slate, size = 2) +
  geom_point(x = q_2, y = p_2, color = met_slate, size = 2) +
  # geom_segment(x = 0, xend = q_1, y = p_1, yend = p_1, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = 0, xend = q_2, y = p_2, yend = p_2, color = met_slate, linetype = "dashed", size = 1) +
  # geom_segment(x = q_1, xend = q_1, y = 0, yend = p_1, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = q_2, xend = q_2, y = 0, yend = p_2, color = met_slate, linetype = "dashed", size = 1) +
  annotate("text", label = "D", x = 10, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) 
```
]

.less-right[

.pink[**Price effect:**] As price decreases, existing customers pay less.

.purple[**Output effect:**] As price decreases, new customers purchase goods.

]

---
count: false
# Marginal Revenue

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
demand <- function(x) 10 - x
inv_demand <- function(x) 10 - x
p_1 <- 1
p_2 <- 3
q_1 <- inv_demand(p_1) 
q_2 <- inv_demand(p_2)
ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  labs(x = "Q", y = "P") +
  geom_rect(aes(xmin = 0, xmax = q_1, ymin = 0, ymax = p_1), alpha = 0.2, fill = green) +
  # geom_rect(aes(xmin = 0, xmax = q_2, ymin = 0, ymax = p_2), alpha = 0.2, fill = green) +
  stat_function(fun = demand, color = red_pink, size = 1) +
  geom_point(x = q_1, y = p_1, color = met_slate, size = 2) +
  geom_point(x = q_2, y = p_2, color = met_slate, size = 2) +
  geom_segment(x = 0, xend = q_1, y = p_1, yend = p_1, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = 0, xend = q_2, y = p_2, yend = p_2, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = q_1, xend = q_1, y = 0, yend = p_1, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = q_2, xend = q_2, y = 0, yend = p_2, color = met_slate, linetype = "dashed", size = 1) +
  annotate("text", label = "D", x = 10, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) 
```
]

.less-right[

.pink[**Price effect:**] As price decreases, existing customers pay less.

.purple[**Output effect:**] As price decreases, new customers purchase goods.

.purple[Output effect] <br> $\quad$ .mono[<] .pink[price effect] .mono[==>] <br> price cut .green[decreases revenue].

]

---
count: false
# Marginal Revenue

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
demand <- function(x) 10 - x
inv_demand <- function(x) 10 - x
p_1 <- 1
p_2 <- 3
q_1 <- inv_demand(p_1) 
q_2 <- inv_demand(p_2)
ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  labs(x = "Q", y = "P") +
  geom_rect(aes(xmin = 0, xmax = q_2, ymin = p_1, ymax = p_2), alpha = 0.2, fill = red_pink) +
  geom_rect(aes(xmin = q_1, xmax = q_2, ymin = 0, ymax = p_1), alpha = 0.2, fill = purple) +
  stat_function(fun = demand, color = red_pink, size = 1) +
  geom_point(x = q_1, y = p_1, color = met_slate, size = 2) +
  geom_point(x = q_2, y = p_2, color = met_slate, size = 2) +
  geom_segment(x = 0, xend = q_1, y = p_1, yend = p_1, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = 0, xend = q_2, y = p_2, yend = p_2, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = q_1, xend = q_1, y = 0, yend = p_1, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = q_2, xend = q_2, y = 0, yend = p_2, color = met_slate, linetype = "dashed", size = 1) +
  annotate("text", label = "D", x = 10, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) 
```
]

.less-right[

.pink[**Price effect:**] As price decreases, existing customers pay less.

.purple[**Output effect:**] As price decreases, new customers purchase goods.

.purple[Output effect] <br> $\quad$ .mono[<] .pink[price effect] .mono[==>] <br> price cut .green[decreases revenue].

]

---
# Marginal Revenue

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
demand <- function(x) 10 - x
inv_demand <- function(x) 10 - x
marginal_revenue <- function(x) 10 - 2*x

ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  labs(x = "Q", y = "P") +
  stat_function(fun = demand, color = red_pink, size = 1) +
  stat_function(fun = marginal_revenue, color = green, size = 1) +
  annotate("text", label = "D", x = 10, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) +
  annotate("text", label = "MR", x = 5.5, y = 0.65, color = green, family = "Fira Sans", size = 9) 
```
]

.less-right[

## Definition 

> Change in total revenue that arises from a one-unit increase in output.


]

---
# Marginal Revenue

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
demand <- function(x) 10 - x
inv_demand <- function(x) 10 - x
marginal_revenue <- function(x) 10 - 2*x

ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  labs(x = "Q", y = "P") +
  stat_function(fun = demand, color = red_pink, size = 1) +
  stat_function(fun = marginal_revenue, color = green, size = 1) +
  annotate("text", label = "D", x = 10, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) +
  annotate("text", label = "MR", x = 5.5, y = 0.65, color = green, family = "Fira Sans", size = 9) 
```
]

.less-right[

A monopolist faces a downward-sloping MR curve.

- MR .mono[>] $0 .mono[-->] increasing revenue.

- MR .mono[=] $0 .mono[-->] maximum revenue.

- MR .mono[<] $0 .mono[-->] decreasing revenue.


]

---
# Profit Maximization

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- seq(0, 8, 0.1)
atc <- function(x) (20 + 12*x - 2*x^2 + (2/3)*x^3)/x
marginal_cost <- function(x) 12 - 4*x + 2*x^2
demand <- function(x) 37 - 5*x
marginal_revenue <- function(x) 37 - 10*x
q <- uniroot(function(x) marginal_revenue(x) - marginal_cost(x), range(x))$root
not_p <- marginal_revenue(q)
p <- demand(q)
cost <- atc(q)
z <- seq(0, q, 0.001)

p_max_1 <- ggplot() +
  scale_x_continuous(limits = c(0, 7.5), expand=c(0,0), breaks = c(q), labels = c(4)) +
  scale_y_continuous(limits = c(0, 37.5), expand=c(0,0), breaks = c(cost, p), labels = c(10, 16)) +
  theme_market +
  labs(x = "Q", y = "$") +
  stat_function(aes(x), fun = atc, color = orange, size = 1) +
  annotate("text", label = "ATC", x = 6.5, y = 35, color = orange, family = "Fira Sans", size = 9) +
  stat_function(aes(x), fun = marginal_cost, color = purple, size = 1) +
  annotate("text", label = "MC", x = 1, y = 8, color = purple, family = "Fira Sans", size = 9)
p_max_1
```
]

.less-right[

**Q:** How does a monopolist maximize profit?

]

---
count: false
# Profit Maximization

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
p_max_1
```
]

.less-right[

**Q:** How does a monopolist maximize profit?

**A:** Two steps!

- **Step 1:** .pink[Pick Q.sub[M] where MR .mono[=] MC.]

]

---
count: false
# Profit Maximization

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
p_max_1 <- p_max_1 +
  stat_function(aes(x), fun = demand, color = red_pink, size = 1) +
  stat_function(aes(x), fun = marginal_revenue, color = green, size = 1) +
  annotate("text", label = "D", x = 6.25, y = 8, color = red_pink, family = "Fira Sans", size = 9) +
  annotate("text", label = "MR", x = 3.5, y = 8, color = green, family = "Fira Sans", size = 9)
p_max_1
```
]

.less-right[

**Q:** How does a monopolist maximize profit?

**A:** Two steps!

- **Step 1:** .pink[Pick Q.sub[M] where MR .mono[=] MC.]

]

---
count: false
# Profit Maximization

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
p_max_1 <- p_max_1 +
  geom_point(aes(x = q, y = not_p), color = met_slate, size = 2) +
  geom_segment(aes(x = q, xend = q, y = 0, yend = not_p), color = met_slate, linetype = "dashed", size = 1)
p_max_1
```
]

.less-right[

**Q:** How does a monopolist maximize profit?

**A:** Two steps!

- **Step 1:** .pink[Pick Q.sub[M] where MR .mono[=] MC.]

]

---
count: false
# Profit Maximization

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
p_max_1
```
]

.less-right[

**Q:** How does a monopolist maximize profit?

**A:** Two steps!

- **Step 1:** .pink[Pick Q.sub[M] where MR .mono[=] MC.]

- **Step 2:** .pink[Set P.sub[M] on the demand curve.]

]

---
count: false
# Profit Maximization

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
p_max_1 <- p_max_1 +
  geom_point(aes(x = q, y = p), color = met_slate, size = 2) +
  geom_segment(aes(x = q, xend = q, y = not_p, yend = p), color = met_slate, linetype = "dashed", size = 1)
p_max_1
```
]

.less-right[

**Q:** How does a monopolist maximize profit?

**A:** Two steps!

- **Step 1:** .pink[Pick Q.sub[M] where MR .mono[=] MC.]

- **Step 2:** .pink[Set P.sub[M] on the demand curve.]

]

---
count: false
# Profit Maximization

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
p_max_1 <- p_max_1 +
  geom_segment(aes(x = 0, xend = q, y = p, yend = p), color = met_slate, linetype = "dashed", size = 1)
p_max_1
```
]

.less-right[

**Q:** How does a monopolist maximize profit?

**A:** Two steps!

- **Step 1:** .pink[Pick Q.sub[M] where MR .mono[=] MC.]

- **Step 2:** .pink[Set P.sub[M] on the demand curve.]

]

---
count: false
# Profit Maximization

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
p_max_1
```
]

.less-right[

**Q:** How does a monopolist maximize profit?

**A:** Two steps!

- **Step 1:** .pink[Pick Q.sub[M] where MR .mono[=] MC.]

- **Step 2:** .pink[Set P.sub[M] on the demand curve.]

Profit .mono[=] (P.sub[M] .mono[-] ATC) &times; Q.sub[M]
]

---
count: false
# Profit Maximization

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
p_max_1 <- p_max_1 +
  geom_point(aes(x = q, y = cost), color = met_slate, size = 2) +
  geom_segment(aes(x = 0, xend = q, y = cost, yend = cost), color = met_slate, linetype = "dashed", size = 1)
p_max_1
```
]

.less-right[

**Q:** How does a monopolist maximize profit?

**A:** Two steps!

- **Step 1:** .pink[Pick Q.sub[M] where MR .mono[=] MC.]

- **Step 2:** .pink[Set P.sub[M] on the demand curve.]

Profit .mono[=] (P.sub[M] .mono[-] ATC) &times; Q.sub[M] <br> $\quad$ .mono[=] ($16 .mono[-] $10) &times; 4
]

---
count: false
# Profit Maximization

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
p_max_1 <- p_max_1 +
  geom_ribbon(aes(x = z, ymin = cost, ymax = p), fill = green, alpha = 0.2, linetype = "blank") +
  annotate("text", label = "Profit", x = q/3, y = (p + cost)/2, color = green, family = "Fira Sans", size = 6) 
p_max_1
```
]

.less-right[

**Q:** How does a monopolist maximize profit?

**A:** Two steps!

- **Step 1:** .pink[Pick Q.sub[M] where MR .mono[=] MC.]

- **Step 2:** .pink[Set P.sub[M] on the demand curve.]

Profit .mono[=] (P.sub[M] .mono[-] ATC) &times; Q.sub[M] <br> $\quad$ .mono[=] ($16 .mono[-] $10) &times; 4 <br> $\quad$ .mono[=] $24.
]

---
# Perfect Competition *vs.* Monopoly

.pull-left[
.center[**Competitive Market**]
```{r, echo = FALSE, fig.height = 4.75, fig.width = 3.75,  dev = "svg"}
x <- 0:10
demand <- function(x) 10 - x
supply <- function(x) 1 + x

q <- uniroot(function(x) demand(x) - supply(x), range(x))$root
p <- demand(q)

ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(q), labels = c(expression(Q[C]))) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(p), labels = c(expression(P[C]))) +
  theme_market +
  labs(x = "Q", y = "P") +
  stat_function(fun = demand, color = red_pink, size = 1) +
  stat_function(fun = supply, color = purple, size = 1) +
  geom_point(x = q, y = p, color = met_slate, size = 2) +
  geom_segment(x = 0, xend = q, y = p, yend = p, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = q, xend = q, y = 0, yend = p, color = met_slate, linetype = "dashed", size = 1) +
  annotate("text", label = "D", x = 10.1, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) +
  annotate("text", label = "S", x = 9.3, y = 9.65, color = purple, family = "Fira Sans", size = 9)
```
]

---
count: false
# Perfect Competition *vs.* Monopoly
.pull-left[
.center[**Competitive Market**]
```{r, echo = FALSE, fig.height = 4.75, fig.width = 3.75,  dev = "svg"}
x <- 0:10
demand <- function(x) 10 - x
supply <- function(x) 1 + x

q <- uniroot(function(x) demand(x) - supply(x), range(x))$root
p <- demand(q)

ggplot(data = data.frame(x = 0), mapping = aes(x = x)) +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(q), labels = c(expression(Q[C]))) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(p), labels = c(expression(P[C]))) +
  theme_market +
  labs(x = "Q", y = "P") +
  stat_function(fun = demand, color = red_pink, size = 1) +
  stat_function(fun = supply, color = purple, size = 1) +
  geom_point(x = q, y = p, color = met_slate, size = 2) +
  geom_segment(x = 0, xend = q, y = p, yend = p, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = q, xend = 10.5, y = p, yend = p, color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(x = q, xend = q, y = 0, yend = p, color = met_slate, linetype = "dashed", size = 1) +
  annotate("text", label = "D", x = 10.1, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) +
  annotate("text", label = "S", x = 9.3, y = 9.65, color = purple, family = "Fira Sans", size = 9)
```
]

.pull-right[
.center[**Monopoly**]
```{r, echo = FALSE, fig.height = 4.75, fig.width = 3.75, dev = "svg"}
x <- 0:10
marginal_cost <- function(x) 1 + x
demand <- function(x) 10 - x
marginal_revenue <- function(x) 10 - 2*x
q <- uniroot(function(x) marginal_revenue(x) - marginal_cost(x), range(x))$root
not_p <- marginal_revenue(q)
p <- demand(q)
q_c <- uniroot(function(x) demand(x) - supply(x), range(x))$root
p_c <- demand(q_c)

ggplot() +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(q, q_c), labels = c(expression(Q[M]), expression(Q[C]))) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(p_c, p), labels = c(expression(P[C]), expression(P[M]))) +
  theme_market +
  labs(x = "Q", y = "P") +
  stat_function(aes(x), fun = marginal_cost, color = purple, size = 1) +
  annotate("text", label = "MC", x = 9.75, y = 9.65, color = purple, family = "Fira Sans", size = 9) +
  stat_function(aes(x), fun = demand, color = red_pink, size = 1) +
  stat_function(aes(x), fun = marginal_revenue, color = green, size = 1) +
  annotate("text", label = "D", x = 10.1, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) +
  annotate("text", label = "MR", x = 5.75, y = 0.65, color = green, family = "Fira Sans", size = 9) +
  geom_point(aes(x = q_c, y = p_c), color = met_slate, size = 2) +
  geom_segment(aes(x = 0, xend = q_c, y = p_c, yend = p_c), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = q_c, xend = q_c, y = 0, yend = p_c), color = met_slate, linetype = "dashed", size = 1) +
  geom_point(aes(x = q, y = p), color = met_slate, size = 2) +
  geom_point(aes(x = q, y = not_p), color = met_slate, size = 2) +
  geom_segment(aes(x = q, xend = q, y = 0, yend = p), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = 0, xend = q, y = p, yend = p), color = met_slate, linetype = "dashed", size = 1)
```
]

---
# Perfect Competition *vs.* Monopoly

.pull-left[
## .purple[Competitive Market]

1. .purple[Many] firms.

2. .purple[No] firm can earn .purple[long-run economic profits].

3. Each firm is a .purple[**price taker**] <br> .mono[-->] no market power!

4. Each firm produces .purple[efficient] level of output <br> .mono[-->] *i.e.,* where P .mono[=] MC.
]

--

.pull-right[
## .pink[Monopoly]

1. .pink[One] firm.

2. Monopolist can earn .pink[long-run economic profits].

3. Monopolist is a .pink[**price maker**] <br> .mono[-->] significant market power!

4. Monopolist produces .pink[inefficient] level of output <br> .mono[-->] *i.e.,* where P .mono[>] MC.
]

---
# Social Consequences

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- 0:10
marginal_cost <- function(x) 1 + x
demand <- function(x) 10 - x
marginal_revenue <- function(x) 10 - 2*x
q <- uniroot(function(x) marginal_revenue(x) - marginal_cost(x), range(x))$root
not_p <- marginal_revenue(q)
p <- demand(q)
q_c <- uniroot(function(x) demand(x) - supply(x), range(x))$root
p_c <- demand(q_c)
w <- seq(0, q, 0.01)
w_2 <- seq(q, q_c, 0.01)

ggplot() +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(q, q_c), labels = c(expression(Q[M]), expression(Q[C]))) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(p_c, p), labels = c(expression(P[C]), expression(P[M]))) +
  theme_market +
  labs(x = "Q", y = "P") +
  stat_function(aes(x), fun = marginal_cost, color = purple, size = 1) +
  annotate("text", label = "MC", x = 9.75, y = 9.65, color = purple, family = "Fira Sans", size = 9) +
  stat_function(aes(x), fun = demand, color = red_pink, size = 1) +
  stat_function(aes(x), fun = marginal_revenue, color = green, size = 1) +
  annotate("text", label = "D", x = 10.1, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) +
  annotate("text", label = "MR", x = 5.75, y = 0.65, color = green, family = "Fira Sans", size = 9) +
  geom_point(aes(x = q_c, y = p_c), color = met_slate, size = 2) +
  geom_segment(aes(x = 0, xend = q_c, y = p_c, yend = p_c), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = q_c, xend = q_c, y = 0, yend = p_c), color = met_slate, linetype = "dashed", size = 1) +
  geom_point(aes(x = q, y = p), color = met_slate, size = 2) +
  geom_point(aes(x = q, y = not_p), color = met_slate, size = 2) +
  geom_segment(aes(x = q, xend = q, y = 0, yend = p), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = 0, xend = q, y = p, yend = p), color = met_slate, linetype = "dashed", size = 1) +
  
  geom_ribbon(aes(x = w, ymin = marginal_cost(w), ymax = p), fill = purple, alpha = 0.2, linetype = "blank") + # PS
  annotate("text", label = "PS", x = q/4, y = p_c - 0.5, color = purple, family = "Fira Sans", size = 5) +
  geom_ribbon(aes(x = w, ymin = p, ymax = demand(w)), fill = red_pink, alpha = 0.2, linetype = "blank") + # CS
  annotate("text", label = "CS", x = q/4, y = p + 0.5, color = red_pink, family = "Fira Sans", size = 5) +
  geom_ribbon(aes(x = w_2, ymin = supply(w_2), ymax = p_c), fill = orange, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w_2, ymin = p_c, ymax = demand(w_2)), fill = orange, alpha = 0.2, linetype = "blank") +
  annotate("text", label = "DWL", x = (q + q_c)/2 - 0.18, y = p_c + 0.3, color = orange, family = "Fira Sans", size = 5)
```
]

.less-right[

## Inefficiency

Monopolies fail to maximize total surplus.

- Q.sub[M] .mono[<] Q.sub[C] .mono[==>] <br> .orange[deadweight loss].

]

---
# Social Consequences

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- 0:10
marginal_cost <- function(x) 1 + x
demand <- function(x) 10 - x
marginal_revenue <- function(x) 10 - 2*x
q <- uniroot(function(x) marginal_revenue(x) - marginal_cost(x), range(x))$root
not_p <- marginal_revenue(q)
p <- demand(q)
q_c <- uniroot(function(x) demand(x) - supply(x), range(x))$root
p_c <- demand(q_c)
w <- seq(0, q, 0.01)
w_2 <- seq(q, q_c, 0.01)

ggplot() +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(q, q_c), labels = c(expression(Q[M]), expression(Q[C]))) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(p_c, p), labels = c(expression(P[C]), expression(P[M]))) +
  theme_market +
  labs(x = "Q", y = "P") +
  stat_function(aes(x), fun = marginal_cost, color = purple, size = 1) +
  annotate("text", label = "MC", x = 9.75, y = 9.65, color = purple, family = "Fira Sans", size = 9) +
  stat_function(aes(x), fun = demand, color = red_pink, size = 1) +
  stat_function(aes(x), fun = marginal_revenue, color = green, size = 1) +
  annotate("text", label = "D", x = 10.1, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) +
  annotate("text", label = "MR", x = 5.75, y = 0.65, color = green, family = "Fira Sans", size = 9) +
  geom_point(aes(x = q_c, y = p_c), color = met_slate, size = 2) +
  geom_segment(aes(x = 0, xend = q_c, y = p_c, yend = p_c), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = q_c, xend = q_c, y = 0, yend = p_c), color = met_slate, linetype = "dashed", size = 1) +
  geom_point(aes(x = q, y = p), color = met_slate, size = 2) +
  geom_point(aes(x = q, y = not_p), color = met_slate, size = 2) +
  geom_segment(aes(x = q, xend = q, y = 0, yend = p), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = 0, xend = q, y = p, yend = p), color = met_slate, linetype = "dashed", size = 1) +
  geom_rect(aes(xmin = 0, xmax = q, ymin = p_c, ymax = p), alpha = 0.2, fill = purple) +
  geom_ribbon(aes(x = w_2, ymin = supply(w_2), ymax = p_c), fill = orange, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w_2, ymin = p_c, ymax = demand(w_2)), fill = orange, alpha = 0.2, linetype = "blank") +
  annotate("text", label = "DWL", x = (q + q_c)/2 - 0.18, y = p_c + 0.3, color = orange, family = "Fira Sans", size = 5)
```
]

.less-right[

## Inefficiency

Monopolies fail to maximize total surplus.

- Q.sub[M] .mono[<] Q.sub[C] .mono[==>] <br> .orange[deadweight loss].

Monopolies .purple[reduce] consumer surplus.

]


---
# Social Consequences

## Limited choices for consumers

Monopolists face few incentives to compete for customers.

- **Result:** Fewer product lines .mono[+] lower quality.

- **Example:** Cable companies and bundled services.

--

## Rent seeking

Monopolists can use political processes to preempt competition or secure new monopolies.

- A form of competition, but not the good kind.

- **Example:** Lobbying Congress for trade protections.

---
# Monopoly

## Solutions?

1. Antitrust law.

    - *e.g,.* breaking up monopolies, blocking mergers, *etc.*

2. Regulation.

    - *e.g.,* price controls.

3. In the case of monopsony, unions?

4. Wait for technological disruptions.

    - *e.g.,* rise of mobile devices reduced Microsoft's market share.

---
class: inverse, middle

# Antitrust

---
# Antitrust

## History

Technological innovations during the 1800s created economies of scale.

- **Result:** Consolidation of industries into national *trusts*..super[.pink[<span>&#8224;</span>]]

.footnote[.pink[<span>&#8224;</span>]: *Trust* .mono[=] An organization of colluding or jointly-owned companies.]

- **Example:** Standard Oil.

--

Sherman Antitrust Act of 1890 made anti-competitive business practices illegal.

- Eventually used to break-up trusts like Standard Oil and block mergers.

---
# Antitrust

## History

Over time, courts started to rule in favor of small incumbent firms over national entrants.

- Small incumbents had local monopoly power!

--

Eventually, federal courts adopted the .pink[Consumer Welfare Standard].

- **Q:** Would a merger increase consumer surplus?

    - If yes, let the merger happen.

    - If no, block the merger.

--

- Still governs antitrust law today.


---
# Mergers

## Market definition

.pink[**Product market**]

- **Q:** Are there .pink[substitutes] for the merging firms' good? 

--

.purple[**Geographic market**]

- **Q:** Is the market .purple[local or national?]

--

**Bottom line** 

- Larger markets make mergers look better.

- Smaller markets make mergers look worse.

---
# Mergers

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- 0:10
marginal_cost <- function(x) 3 + x
demand <- function(x) 10 - x
marginal_revenue <- function(x) 10 - 2*x
q <- uniroot(function(x) marginal_revenue(x) - marginal_cost(x), range(x))$root
not_p <- marginal_revenue(q)
p <- demand(q)
q_c <- uniroot(function(x) demand(x) - marginal_cost(x), range(x))$root
p_c <- demand(q_c)
w <- seq(0, q, 0.01)
w_2 <- seq(q, q_c, 0.01)

ggplot() +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(q, q_c), labels = c(expression(Q[M]), expression(Q[C]))) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(p_c, p), labels = c(expression(P[C]), expression(P[M]))) +
  theme_market +
  labs(x = "Q", y = "P") +
  stat_function(aes(x), fun = marginal_cost, color = purple, size = 1) +
  annotate("text", label = expression(MC[1]), x = 7.75, y = 9.65, color = purple, family = "Fira Sans", size = 9) +
  stat_function(aes(x), fun = demand, color = red_pink, size = 1) +
  stat_function(aes(x), fun = marginal_revenue, color = green, size = 1) +
  annotate("text", label = "D", x = 10.1, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) +
  annotate("text", label = "MR", x = 5.75, y = 0.65, color = green, family = "Fira Sans", size = 9) +
  geom_point(aes(x = q_c, y = p_c), color = met_slate, size = 2) +
  geom_segment(aes(x = 0, xend = q_c, y = p_c, yend = p_c), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = q_c, xend = q_c, y = 0, yend = p_c), color = met_slate, linetype = "dashed", size = 1) +
  geom_point(aes(x = q, y = p), color = met_slate, size = 2) +
  geom_segment(aes(x = q, xend = q, y = 0, yend = p), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = 0, xend = q, y = p, yend = p), color = met_slate, linetype = "dashed", size = 1) +
  geom_ribbon(aes(x = w_2, ymin = marginal_cost(w_2), ymax = p_c), fill = orange, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w_2, ymin = p_c, ymax = demand(w_2)), fill = orange, alpha = 0.2, linetype = "blank")
```
]

.less-right[

All else being equal, industry consolidation creates .orange[deadweight loss].

]

---
count: false
# Mergers

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- 0:10
marginal_cost <- function(x) 3 + x
demand <- function(x) 10 - x
marginal_revenue <- function(x) 10 - 2*x
q <- uniroot(function(x) marginal_revenue(x) - marginal_cost(x), range(x))$root
not_p <- marginal_revenue(q)
p <- demand(q)
q_c <- uniroot(function(x) demand(x) - marginal_cost(x), range(x))$root
p_c <- demand(q_c)

marginal_cost_2 <- function(x) 0.25*x
q_2 <- uniroot(function(x) marginal_revenue(x) - marginal_cost_2(x), range(x))$root
p_2 <- demand(q_2)
q_c2 <- uniroot(function(x) demand(x) - marginal_cost_2(x), range(x))$root
p_c2 <- demand(q_c2)
w <- seq(0, q_c, 0.01)
w_2 <- seq(q_2, q_c2, 0.01)

ggplot() +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(q_2, q_c), labels = c(expression(Q[M]), expression(Q[C]))) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(p_c, p_2), labels = c(expression(P[C]), expression(P[M]))) +
  theme_market +
  labs(x = "Q", y = "P") +
  stat_function(aes(x), fun = marginal_cost, color = purple, size = 1, alpha = 0.5) +
  annotate("text", label = expression(MC[1]), x = 7.75, y = 9.65, color = purple, family = "Fira Sans", size = 9, alpha = 0.5) +
  stat_function(aes(x), fun = marginal_cost_2, color = purple, size = 1) +
  annotate("text", label = expression(MC[2]), x = 9.75, y = 3.15, color = purple, family = "Fira Sans", size = 9) +
  stat_function(aes(x), fun = demand, color = red_pink, size = 1) +
  stat_function(aes(x), fun = marginal_revenue, color = green, size = 1) +
  annotate("text", label = "D", x = 10.1, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) +
  annotate("text", label = "MR", x = 5.75, y = 0.65, color = green, family = "Fira Sans", size = 9) +
  geom_point(aes(x = q_c, y = p_c), color = met_slate, size = 2) +
  geom_segment(aes(x = 0, xend = q_c, y = p_c, yend = p_c), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = q_c, xend = q_c, y = 0, yend = p_c), color = met_slate, linetype = "dashed", size = 1) +
  
  # geom_point(aes(x = q, y = p), color = met_slate, size = 2, alpha = 0.5) +
  # geom_segment(aes(x = q, xend = q, y = 0, yend = p), color = met_slate, linetype = "dashed", size = 1, alpha = 0.5) +
  # geom_segment(aes(x = 0, xend = q, y = p, yend = p), color = met_slate, linetype = "dashed", size = 1, alpha = 0.5) +
  
  geom_point(aes(x = q_2, y = p_2), color = met_slate, size = 2) +
  geom_segment(aes(x = q_2, xend = q_2, y = 0, yend = p_2), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = 0, xend = q_2, y = p_2, yend = p_2), color = met_slate, linetype = "dashed", size = 1) +
  geom_ribbon(aes(x = w, ymin = p_c, ymax = demand(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w_2, ymin = marginal_cost_2(w_2), ymax = p_c2), fill = orange, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w_2, ymin = p_c2, ymax = demand(w_2)), fill = orange, alpha = 0.2, linetype = "blank")
```
]

.less-right[

All else being equal, industry consolidation creates .orange[deadweight loss].

In some cases, consolidation can reduce marginal costs.

Possible to .pink[increase] consumer surplus relative to original competitive equilibrium.

]

---
count: false
# Mergers

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- 0:10
marginal_cost <- function(x) 3 + x
demand <- function(x) 10 - x
marginal_revenue <- function(x) 10 - 2*x
q <- uniroot(function(x) marginal_revenue(x) - marginal_cost(x), range(x))$root
not_p <- marginal_revenue(q)
p <- demand(q)
q_c <- uniroot(function(x) demand(x) - marginal_cost(x), range(x))$root
p_c <- demand(q_c)

marginal_cost_2 <- function(x) 0.25*x
q_2 <- uniroot(function(x) marginal_revenue(x) - marginal_cost_2(x), range(x))$root
p_2 <- demand(q_2)
q_c2 <- uniroot(function(x) demand(x) - marginal_cost_2(x), range(x))$root
p_c2 <- demand(q_c2)
w <- seq(0, q_2, 0.01)
w_2 <- seq(q_2, q_c2, 0.01)

ggplot() +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(q_2, q_c), labels = c(expression(Q[M]), expression(Q[C]))) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(p_c, p_2), labels = c(expression(P[C]), expression(P[M]))) +
  theme_market +
  labs(x = "Q", y = "P") +
  stat_function(aes(x), fun = marginal_cost, color = purple, size = 1, alpha = 0.5) +
  annotate("text", label = expression(MC[1]), x = 7.75, y = 9.65, color = purple, family = "Fira Sans", size = 9, alpha = 0.5) +
  stat_function(aes(x), fun = marginal_cost_2, color = purple, size = 1) +
  annotate("text", label = expression(MC[2]), x = 9.75, y = 3.15, color = purple, family = "Fira Sans", size = 9) +
  stat_function(aes(x), fun = demand, color = red_pink, size = 1) +
  stat_function(aes(x), fun = marginal_revenue, color = green, size = 1) +
  annotate("text", label = "D", x = 10.1, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) +
  annotate("text", label = "MR", x = 5.75, y = 0.65, color = green, family = "Fira Sans", size = 9) +
  geom_point(aes(x = q_c, y = p_c), color = met_slate, size = 2) +
  geom_segment(aes(x = 0, xend = q_c, y = p_c, yend = p_c), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = q_c, xend = q_c, y = 0, yend = p_c), color = met_slate, linetype = "dashed", size = 1) +
  
  # geom_point(aes(x = q, y = p), color = met_slate, size = 2, alpha = 0.5) +
  # geom_segment(aes(x = q, xend = q, y = 0, yend = p), color = met_slate, linetype = "dashed", size = 1, alpha = 0.5) +
  # geom_segment(aes(x = 0, xend = q, y = p, yend = p), color = met_slate, linetype = "dashed", size = 1, alpha = 0.5) +
  
  geom_point(aes(x = q_2, y = p_2), color = met_slate, size = 2) +
  geom_segment(aes(x = q_2, xend = q_2, y = 0, yend = p_2), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = 0, xend = q_2, y = p_2, yend = p_2), color = met_slate, linetype = "dashed", size = 1) +
  geom_ribbon(aes(x = w, ymin = p_2, ymax = demand(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w_2, ymin = marginal_cost_2(w_2), ymax = p_c2), fill = orange, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w_2, ymin = p_c2, ymax = demand(w_2)), fill = orange, alpha = 0.2, linetype = "blank")
```
]

.less-right[

All else being equal, industry consolidation creates .orange[deadweight loss].

In some cases, consolidation can reduce marginal costs.

Possible to .pink[increase] consumer surplus relative to original competitive equilibrium.

]

---
class: clear-slide

**Podcast Question:** True or false? According to Charlotte Slaiman, mergers usually result in lower prices for consumers.

> **A.** True.

> **B.** False.

---
count: false
class: clear-slide

**Podcast Question:** True or false? According to Charlotte Slaiman, mergers usually result in lower prices for consumers.

> **A.** True.

> .pink[**B.** False.]

---
class: inverse, middle

# Regulation

---
# Regulating Monopoly

## Options

1. Mandate marginal cost pricing?

    - Make monopolist choose quantity where price equals marginal cost.

2. Have the government take over the monopoly?

    - Not always better.

3. Subsidize the monopolist?

    - Politically unpopular in most cases.

---
# Regulating Monopoly

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- 0:10
marginal_cost <- function(x) 2 + 1.25*x
demand <- function(x) 10 - x
marginal_revenue <- function(x) 10 - 2*x
q <- uniroot(function(x) marginal_revenue(x) - marginal_cost(x), range(x))$root
not_p <- marginal_revenue(q)
p <- demand(q)
q_c <- uniroot(function(x) demand(x) - marginal_cost(x), range(x))$root
p_c <- demand(q_c)
w <- seq(0, q, 0.01)
w_2 <- seq(q, q_c, 0.01)

ggplot() +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(q, q_c), labels = c(expression(Q[M]), expression(Q[C]))) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(p_c, p), labels = c(expression(P[C]), expression(P[M]))) +
  theme_market +
  labs(x = "Q", y = "P") +
  stat_function(aes(x), fun = marginal_cost, color = purple, size = 1) +
  annotate("text", label = "MC", x = 7.25, y = 9.65, color = purple, family = "Fira Sans", size = 9) +
  stat_function(aes(x), fun = demand, color = red_pink, size = 1) +
  stat_function(aes(x), fun = marginal_revenue, color = green, size = 1) +
  annotate("text", label = "D", x = 10.1, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) +
  annotate("text", label = "MR", x = 5.75, y = 0.65, color = green, family = "Fira Sans", size = 9) +
  geom_point(aes(x = q_c, y = p_c), color = met_slate, size = 2) +
  geom_segment(aes(x = 0, xend = q_c, y = p_c, yend = p_c), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = q_c, xend = q_c, y = 0, yend = p_c), color = met_slate, linetype = "dashed", size = 1) +
  geom_point(aes(x = q, y = p), color = met_slate, size = 2) +
  geom_point(aes(x = q, y = not_p), color = met_slate, size = 2) +
  geom_segment(aes(x = q, xend = q, y = 0, yend = p), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = 0, xend = q, y = p, yend = p), color = met_slate, linetype = "dashed", size = 1) +
  
  geom_ribbon(aes(x = w, ymin = marginal_cost(w), ymax = p), fill = purple, alpha = 0.2, linetype = "blank") + # PS
  geom_ribbon(aes(x = w, ymin = p, ymax = demand(w)), fill = red_pink, alpha = 0.2, linetype = "blank") + # CS
  geom_ribbon(aes(x = w_2, ymin = marginal_cost(w_2), ymax = p_c), fill = orange, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w_2, ymin = p_c, ymax = demand(w_2)), fill = orange, alpha = 0.2, linetype = "blank") +
  annotate("text", label = "PS", x = q/4, y = p_c - 0.5, color = purple, family = "Fira Sans", size = 5) +
  annotate("text", label = "CS", x = q/4, y = p + 0.5, color = red_pink, family = "Fira Sans", size = 5) 
```
]

.less-right[

**Q:** Can a price ceiling at the competitive price eliminate .orange[deadweight loss]?

]

---
count: false
# Regulating Monopoly

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- 0:10
marginal_cost <- function(x) 2 + 1.25*x
demand <- function(x) 10 - x
marginal_revenue <- function(x) 10 - 2*x
q <- uniroot(function(x) marginal_revenue(x) - marginal_cost(x), range(x))$root
not_p <- marginal_revenue(q)
p <- demand(q)
q_c <- uniroot(function(x) demand(x) - marginal_cost(x), range(x))$root
p_c <- demand(q_c)
w <- seq(0, q, 0.01)
w_2 <- seq(q, q_c, 0.01)

ggplot() +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(q, q_c), labels = c(expression(Q[M]), expression(Q[C]))) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(p_c, p), labels = c(expression(P[C]), expression(P[M]))) +
  theme_market +
  labs(x = "Q", y = "P") +
  stat_function(aes(x), fun = marginal_cost, color = purple, size = 1) +
  annotate("text", label = "MC", x = 7.25, y = 9.65, color = purple, family = "Fira Sans", size = 9) +
  stat_function(aes(x), fun = demand, color = red_pink, size = 1) +
  stat_function(aes(x), fun = marginal_revenue, color = green, size = 1) +
  annotate("text", label = "D", x = 10.1, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) +
  annotate("text", label = "MR", x = 5.75, y = 0.65, color = green, family = "Fira Sans", size = 9) +
  geom_point(aes(x = q_c, y = p_c), color = met_slate, size = 2) +
  geom_segment(aes(x = 0, xend = q_c, y = p_c, yend = p_c), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = q_c, xend = q_c, y = 0, yend = p_c), color = met_slate, linetype = "dashed", size = 1) +
  geom_point(aes(x = q, y = p), color = met_slate, size = 2) +
  geom_point(aes(x = q, y = not_p), color = met_slate, size = 2) +
  geom_segment(aes(x = q, xend = q, y = 0, yend = p), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = 0, xend = q, y = p, yend = p), color = met_slate, linetype = "dashed", size = 1)
```
]

.less-right[

**Q:** Can a price ceiling at the competitive price eliminate deadweight loss?

]

---
count: false
# Regulating Monopoly

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- 0:10
marginal_cost <- function(x) 2 + 1.25*x
demand <- function(x) 10 - x
marginal_revenue <- function(x) 10 - 2*x
q <- uniroot(function(x) marginal_revenue(x) - marginal_cost(x), range(x))$root
not_p <- marginal_revenue(q)
p <- demand(q)
q_c <- uniroot(function(x) demand(x) - marginal_cost(x), range(x))$root
p_c <- demand(q_c)
w <- seq(0, q, 0.01)
w_2 <- seq(q, q_c, 0.01)

ggplot() +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(q, q_c), labels = c(expression(Q[M]), expression(Q[C]))) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(p_c, p), labels = c(expression(P[C]), expression(P[M]))) +
  theme_market +
  labs(x = "Q", y = "P") +
  stat_function(aes(x), fun = marginal_cost, color = purple, size = 1) +
  annotate("text", label = "MC", x = 7.25, y = 9.65, color = purple, family = "Fira Sans", size = 9) +
  stat_function(aes(x), fun = demand, color = red_pink, size = 1) +
  stat_function(aes(x), fun = marginal_revenue, color = green, size = 1) +
  annotate("text", label = "D", x = 10.1, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) +
  annotate("text", label = "MR", x = 5.75, y = 0.65, color = green, family = "Fira Sans", size = 9) +
  geom_point(aes(x = q_c, y = p_c), color = met_slate, size = 2) +
  geom_segment(aes(x = q_c, xend = q_c, y = 0, yend = p_c), color = met_slate, linetype = "dashed", size = 1) +
  geom_point(aes(x = q, y = p), color = met_slate, size = 2) +
  geom_point(aes(x = q, y = not_p), color = met_slate, size = 2) +
  geom_segment(aes(x = q, xend = q, y = 0, yend = p), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = 0, xend = q, y = p, yend = p), color = met_slate, linetype = "dashed", size = 1) +
  geom_hline(yintercept = p_c, color = met_slate, size = 1) +
  annotate("text", label = "Price Ceiling", x = 9, y = p_c - 0.3, color = met_slate, family = "Fira Sans", size = 5)
```
]

.less-right[

**Q:** Can a price ceiling at the competitive price eliminate deadweight loss?

]

---
count: false
# Regulating Monopoly

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- 0:10
marginal_cost <- function(x) 2 + 1.25*x
demand <- function(x) 10 - x
marginal_revenue <- function(x) 10 - 2*x
q <- uniroot(function(x) marginal_revenue(x) - marginal_cost(x), range(x))$root
not_p <- marginal_revenue(q)
p <- demand(q)
q_c <- uniroot(function(x) demand(x) - marginal_cost(x), range(x))$root
p_c <- demand(q_c)
w <- seq(0, q, 0.01)
w_2 <- seq(q, q_c, 0.01)
z <- seq(q_c, 10, 0.01)
not_p2 <- marginal_revenue(q_c)

ggplot() +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(q_c), labels = c(expression(Q[C]==Q[M]))) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(p_c), labels = c(expression(P[M]))) +
  theme_market +
  labs(x = "Q", y = "P") +
  stat_function(aes(x), fun = marginal_cost, color = purple, size = 1) +
  annotate("text", label = "MC", x = 7.25, y = 9.65, color = purple, family = "Fira Sans", size = 9) +
  stat_function(aes(x), fun = demand, color = red_pink, size = 1) +
  stat_function(aes(x), fun = marginal_revenue, color = green, size = 1, alpha = 0.5) +
  stat_function(aes(z), xlim = c(q_c, 10.5), fun = marginal_revenue, color = green, size = 1) +
  geom_segment(aes(x = 0, y = p_c, xend = q_c, yend = p_c), color = green, size = 1.5) +
  geom_segment(aes(x = q_c, y = p_c, xend = q_c, yend = not_p2), color = green, size = 1) +
  annotate("text", label = "D", x = 10.1, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) +
  annotate("text", label = "MR", x = 5.75, y = 0.65, color = green, family = "Fira Sans", size = 9) +
  geom_point(aes(x = q_c, y = p_c), color = met_slate, size = 2) +
  geom_segment(aes(x = q_c, xend = q_c, y = 0, yend = p_c), color = met_slate, linetype = "dashed", size = 1) +
  
  geom_point(aes(x = q, y = p), color = met_slate, size = 2, alpha = 0.5) +
  geom_point(aes(x = q, y = not_p), color = met_slate, size = 2, alpha = 0.5) +
  geom_segment(aes(x = q, xend = q, y = 0, yend = p), color = met_slate, linetype = "dashed", size = 1, alpha = 0.5) +
  geom_segment(aes(x = 0, xend = q, y = p, yend = p), color = met_slate, linetype = "dashed", size = 1, alpha = 0.5) +
  
  geom_hline(yintercept = p_c, color = met_slate, size = 1) +
  annotate("text", label = "Price Ceiling", x = 9, y = p_c - 0.3, color = met_slate, family = "Fira Sans", size = 5)
```
]

.less-right[

**Q:** Can a price ceiling at the competitive price eliminate deadweight loss?

]

---
count: false
# Regulating Monopoly

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- 0:10
marginal_cost <- function(x) 2 + 1.25*x
demand <- function(x) 10 - x
marginal_revenue <- function(x) 10 - 2*x
q <- uniroot(function(x) marginal_revenue(x) - marginal_cost(x), range(x))$root
not_p <- marginal_revenue(q)
p <- demand(q)
q_c <- uniroot(function(x) demand(x) - marginal_cost(x), range(x))$root
p_c <- demand(q_c)
w <- seq(0, q, 0.01)
w_2 <- seq(q, q_c, 0.01)
z <- seq(q_c, 10, 0.01)
not_p2 <- marginal_revenue(q_c)

ggplot() +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(q_c), labels = c(expression(Q[C]==Q[M]))) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(p_c), labels = c(expression(P[M]))) +
  theme_market +
  labs(x = "Q", y = "P") +
  stat_function(aes(x), fun = marginal_cost, color = purple, size = 1) +
  annotate("text", label = "MC", x = 7.25, y = 9.65, color = purple, family = "Fira Sans", size = 9) +
  stat_function(aes(x), fun = demand, color = red_pink, size = 1) +
  stat_function(aes(z), xlim = c(q_c, 10.5), fun = marginal_revenue, color = green, size = 1) +
  geom_segment(aes(x = 0, y = p_c, xend = q_c, yend = p_c), color = green, size = 1.5) +
  geom_segment(aes(x = q_c, y = p_c, xend = q_c, yend = not_p2), color = green, size = 1) +
  annotate("text", label = "D", x = 10.1, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) +
  annotate("text", label = "MR", x = 5.75, y = 0.65, color = green, family = "Fira Sans", size = 9) +
  geom_point(aes(x = q_c, y = p_c), color = met_slate, size = 2) +
  geom_segment(aes(x = q_c, xend = q_c, y = 0, yend = p_c), color = met_slate, linetype = "dashed", size = 1) +
  geom_hline(yintercept = p_c, color = met_slate, size = 1) +
  annotate("text", label = "Price Ceiling", x = 9, y = p_c - 0.3, color = met_slate, family = "Fira Sans", size = 5)
```
]

.less-right[

**Q:** Can a price ceiling at the competitive price eliminate deadweight loss?

]

---
count: false
# Regulating Monopoly

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- 0:10
marginal_cost <- function(x) 2 + 1.25*x
demand <- function(x) 10 - x
marginal_revenue <- function(x) 10 - 2*x
q <- uniroot(function(x) marginal_revenue(x) - marginal_cost(x), range(x))$root
not_p <- marginal_revenue(q)
p <- demand(q)
q_c <- uniroot(function(x) demand(x) - marginal_cost(x), range(x))$root
p_c <- demand(q_c)
w <- seq(0, q, 0.01)
w_2 <- seq(0, q_c, 0.01)
z <- seq(q_c, 10, 0.01)
not_p2 <- marginal_revenue(q_c)

ggplot() +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(q_c), labels = c(expression(Q[C]==Q[M]))) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(p_c), labels = c(expression(P[M]))) +
  theme_market +
  labs(x = "Q", y = "P") +
  stat_function(aes(x), fun = marginal_cost, color = purple, size = 1) +
  annotate("text", label = "MC", x = 7.25, y = 9.65, color = purple, family = "Fira Sans", size = 9) +
  stat_function(aes(x), fun = demand, color = red_pink, size = 1) +
  stat_function(aes(z), xlim = c(q_c, 10.5), fun = marginal_revenue, color = green, size = 1) +
  geom_segment(aes(x = 0, y = p_c, xend = q_c, yend = p_c), color = green, size = 1.5) +
  geom_segment(aes(x = q_c, y = p_c, xend = q_c, yend = not_p2), color = green, size = 1) +
  annotate("text", label = "D", x = 10.1, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) +
  annotate("text", label = "MR", x = 5.75, y = 0.65, color = green, family = "Fira Sans", size = 9) +
  geom_point(aes(x = q_c, y = p_c), color = met_slate, size = 2) +
  geom_segment(aes(x = q_c, xend = q_c, y = 0, yend = p_c), color = met_slate, linetype = "dashed", size = 1) +
  geom_hline(yintercept = p_c, color = met_slate, size = 1) +
  annotate("text", label = "Price Ceiling", x = 9, y = p_c - 0.3, color = met_slate, family = "Fira Sans", size = 5) +
  geom_ribbon(aes(x = w_2, ymin = marginal_cost(w_2), ymax = p_c), fill = purple, alpha = 0.2, linetype = "blank") + # PS
  geom_ribbon(aes(x = w_2, ymin = p_c, ymax = demand(w_2)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  annotate("text", label = "PS", x = q/3, y = p_c - 1, color = purple, family = "Fira Sans", size = 5) +
  annotate("text", label = "CS", x = q/3, y = p_c + 1, color = red_pink, family = "Fira Sans", size = 5)
```
]

.less-right[

**Q:** Can a price ceiling at the competitive price eliminate deadweight loss?

**A:** .pink[Yes!]
]

---
class: clear-slide

**Q:** Can a tax eliminate .orange[deadweight loss] from the monopoly?.super[.pink[<span>&#8224;</span>]]

.footnote[.pink[<span>&#8224;</span>]: Assume that there are no negative externalities.]

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- 0:10
marginal_cost <- function(x) 1 + x
demand <- function(x) 10 - x
marginal_revenue <- function(x) 10 - 2*x
q <- uniroot(function(x) marginal_revenue(x) - marginal_cost(x), range(x))$root
not_p <- marginal_revenue(q)
p <- demand(q)
q_c <- uniroot(function(x) demand(x) - marginal_cost(x), range(x))$root
p_c <- demand(q_c)
w <- seq(0, q, 0.01)
w_2 <- seq(q, q_c, 0.01)

ggplot() +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(q, q_c), labels = c(expression(Q[M]), expression(Q[C]))) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(p_c, p), labels = c(expression(P[C]), expression(P[M]))) +
  theme_market +
  labs(x = "Q", y = "P") +
  stat_function(aes(x), fun = marginal_cost, color = purple, size = 1) +
  annotate("text", label = expression(MC), x = 9.75, y = 9.65, color = purple, family = "Fira Sans", size = 9) +
  stat_function(aes(x), fun = demand, color = red_pink, size = 1) +
  stat_function(aes(x), fun = marginal_revenue, color = green, size = 1) +
  annotate("text", label = "D", x = 10.1, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) +
  annotate("text", label = "MR", x = 5.75, y = 0.65, color = green, family = "Fira Sans", size = 9) +
  geom_point(aes(x = q_c, y = p_c), color = met_slate, size = 2) +
  geom_segment(aes(x = 0, xend = q_c, y = p_c, yend = p_c), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = q_c, xend = q_c, y = 0, yend = p_c), color = met_slate, linetype = "dashed", size = 1) +
  geom_point(aes(x = q, y = p), color = met_slate, size = 2) +
  geom_point(aes(x = q, y = not_p), color = met_slate, size = 2) +
  geom_segment(aes(x = q, xend = q, y = 0, yend = p), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = 0, xend = q, y = p, yend = p), color = met_slate, linetype = "dashed", size = 1) +
  geom_ribbon(aes(x = w_2, ymin = marginal_cost(w_2), ymax = p_c), fill = orange, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w_2, ymin = p_c, ymax = demand(w_2)), fill = orange, alpha = 0.2, linetype = "blank") 
```
]

.less-right[

> **A.** Yes.

> **B.** No.

> **C.** Depends

]

---
count: false
class: clear-slide

**Q:** Can a tax eliminate .orange[deadweight loss] from the monopoly?.super[.pink[<span>&#8224;</span>]]

.footnote[.pink[<span>&#8224;</span>]: Assume that there are no negative externalities.]

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- 0:10
marginal_cost <- function(x) 1 + x
demand <- function(x) 10 - x
marginal_revenue <- function(x) 10 - 2*x
q <- uniroot(function(x) marginal_revenue(x) - marginal_cost(x), range(x))$root
not_p <- marginal_revenue(q)
p <- demand(q)
q_c <- uniroot(function(x) demand(x) - marginal_cost(x), range(x))$root
p_c <- demand(q_c)

marginal_cost_2 <- function(x) 4 + x
q_2 <- uniroot(function(x) marginal_revenue(x) - marginal_cost_2(x), range(x))$root
not_p_2 <- marginal_revenue(q_2)
p_2 <- demand(q_2)
w_2 <- seq(q_2, q_c, 0.01)

ggplot() +
  scale_x_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(q_2, q_c), labels = c(expression(Q[M]), expression(Q[C]))) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = c(p_c, p_2), labels = c(expression(P[C]), expression(P[M]))) +
  theme_market +
  labs(x = "Q", y = "P") +
  stat_function(aes(x), fun = marginal_cost, color = purple, size = 1, alpha = 0.5) +
  annotate("text", label = expression(MC), x = 9.75, y = 9.65, color = purple, family = "Fira Sans", size = 9, alpha = 0.5) +
  
  stat_function(aes(x), fun = marginal_cost_2, color = purple, size = 1) +
  annotate("text", label = expression(MC[Tax]), x = 7.1, y = 9.65, color = purple, family = "Fira Sans", size = 9) +
  
  stat_function(aes(x), fun = demand, color = red_pink, size = 1) +
  stat_function(aes(x), fun = marginal_revenue, color = green, size = 1) +
  annotate("text", label = "D", x = 10.1, y = 0.65, color = red_pink, family = "Fira Sans", size = 9) +
  annotate("text", label = "MR", x = 5.75, y = 0.65, color = green, family = "Fira Sans", size = 9) +
  geom_point(aes(x = q_c, y = p_c), color = met_slate, size = 2) +
  geom_segment(aes(x = 0, xend = q_c, y = p_c, yend = p_c), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = q_c, xend = q_c, y = 0, yend = p_c), color = met_slate, linetype = "dashed", size = 1) +
  
  geom_point(aes(x = q, y = p), color = met_slate, size = 2, alpha = 0.5) +
  geom_point(aes(x = q, y = not_p), color = met_slate, size = 2, alpha = 0.5) +
  geom_segment(aes(x = q, xend = q, y = 0, yend = p), color = met_slate, linetype = "dashed", size = 1, alpha = 0.5) +
  geom_segment(aes(x = 0, xend = q, y = p, yend = p), color = met_slate, linetype = "dashed", size = 1, alpha = 0.5) +
  
  geom_point(aes(x = q_2, y = p_2), color = met_slate, size = 2) +
  geom_point(aes(x = q_2, y = not_p_2), color = met_slate, size = 2) +
  geom_segment(aes(x = q_2, xend = q_2, y = 0, yend = p_2), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = 0, xend = q_2, y = p_2, yend = p_2), color = met_slate, linetype = "dashed", size = 1) +
  
  geom_ribbon(aes(x = w_2, ymin = marginal_cost(w_2), ymax = p_c), fill = orange, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w_2, ymin = p_c, ymax = demand(w_2)), fill = orange, alpha = 0.2, linetype = "blank") 
```
]

.less-right[

> **A.** Yes.

> .pink[**B.** No.]

> **C.** Depends.

]

---
# Regulating Natural Monopoly

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- seq(0, 8, 0.1)
atc <- function(x) (20 + 12*x)/x
marginal_cost <- function(x) 12
demand <- function(x) 37 - 5*x
marginal_revenue <- function(x) 37 - 10*x
q <- uniroot(function(x) marginal_revenue(x) - marginal_cost(x), range(x))$root
not_p <- marginal_revenue(q)
p <- demand(q)
cost <- atc(q)
z <- seq(0, q, 0.001)

ggplot() +
  scale_x_continuous(limits = c(0, 7.5), expand=c(0,0), breaks = c(q), labels = c(expression(Q[M]))) +
  scale_y_continuous(limits = c(0, 37.5), expand=c(0,0), breaks = c(cost, p), labels = c(expression(ATC[M]), expression(P[M]))) +
  theme_market +
  labs(x = "Q", y = "$") +
  stat_function(aes(x), fun = atc, color = orange, size = 1) +
  annotate("text", label = "ATC", x = 6.5, y = 17, color = orange, family = "Fira Sans", size = 9) +
  stat_function(aes(x), fun = marginal_cost, color = purple, size = 1) +
  annotate("text", label = "MC", x = 7, y = 10.5, color = purple, family = "Fira Sans", size = 9) +
  
  stat_function(aes(x), fun = demand, color = red_pink, size = 1) +
  stat_function(aes(x), fun = marginal_revenue, color = green, size = 1) +
  annotate("text", label = "D", x = 7.1, y = 4, color = red_pink, family = "Fira Sans", size = 9) +
  annotate("text", label = "MR", x = 3.9, y = 4.5, color = green, family = "Fira Sans", size = 9) +
  geom_point(aes(x = q, y = not_p), color = met_slate, size = 2) +
  geom_segment(aes(x = q, xend = q, y = 0, yend = not_p), color = met_slate, linetype = "dashed", size = 1) +
  geom_ribbon(aes(x = z, ymin = cost, ymax = p), fill = green, alpha = 0.2, linetype = "blank") +
  annotate("text", label = "Profit", x = q/3, y = (p + cost)/2, color = green, family = "Fira Sans", size = 6) +
  geom_point(aes(x = q, y = p), color = met_slate, size = 2) +
  geom_point(aes(x = q, y = cost), color = met_slate, size = 2) +
  geom_segment(aes(x = q, xend = q, y = not_p, yend = p), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = 0, xend = q, y = p, yend = p), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = 0, xend = q, y = cost, yend = cost), color = met_slate, linetype = "dashed", size = 1)
```
]

.less-right[
## Price Ceiling

Some natural monopolies are not profitable under marginal cost pricing.

]

---
count: false
# Regulating Natural Monopoly

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- seq(0, 8, 0.1)
atc <- function(x) (20 + 12*x)/x
marginal_cost <- function(x) 12
demand <- function(x) 37 - 5*x
marginal_revenue <- function(x) 37 - 10*x
q <- uniroot(function(x) marginal_revenue(x) - marginal_cost(x), range(x))$root
not_p <- marginal_revenue(q)
p <- demand(q)
cost <- atc(q)
z <- seq(0, q, 0.001)
q_c <- uniroot(function(x) demand(x) - marginal_cost(x), range(x))$root
p_c <- demand(q_c)
cost_2 <- atc(q_c)
z_2 <- seq(0, q_c, 0.001)


ggplot() +
  scale_x_continuous(limits = c(0, 7.5), expand=c(0,0), breaks = c(q, q_c), labels = c(expression(Q[M]), expression(Q[R]))) +
  scale_y_continuous(limits = c(0, 37.5), expand=c(0,0), breaks = c(p_c, cost_2, cost, p), labels = c(expression(P[R]), expression(ATC[R]), expression(ATC[M]), expression(P[M]))) +
  theme_market +
  labs(x = "Q", y = "$") +
  stat_function(aes(x), fun = atc, color = orange, size = 1) +
  annotate("text", label = "ATC", x = 6.5, y = 17, color = orange, family = "Fira Sans", size = 9) +
  stat_function(aes(x), fun = marginal_cost, color = purple, size = 1) +
  annotate("text", label = "MC", x = 7, y = 10.5, color = purple, family = "Fira Sans", size = 9) +
  
  stat_function(aes(x), fun = demand, color = red_pink, size = 1) +
  stat_function(aes(x), fun = marginal_revenue, color = green, size = 1) +
  annotate("text", label = "D", x = 7.1, y = 4, color = red_pink, family = "Fira Sans", size = 9) +
  annotate("text", label = "MR", x = 3.9, y = 4.5, color = green, family = "Fira Sans", size = 9) +
  geom_point(aes(x = q, y = not_p), color = met_slate, size = 2) +
  geom_segment(aes(x = q, xend = q, y = 0, yend = not_p), color = met_slate, linetype = "dashed", size = 1) +
  geom_ribbon(aes(x = z_2, ymin = p_c, ymax = cost_2), fill = red_pink, alpha = 0.2, linetype = "blank") +
  annotate("text", label = "Profit", x = q/3, y = (p + cost)/2, color = green, family = "Fira Sans", size = 6) +
  geom_ribbon(aes(x = z, ymin = cost, ymax = p), fill = green, alpha = 0.2, linetype = "blank") +
  annotate("text", label = "Loss", x = q/3, y = (p_c + cost_2)/2, color = red_pink, family = "Fira Sans", size = 6) +
  geom_point(aes(x = q, y = p), color = met_slate, size = 2) +
  geom_point(aes(x = q, y = cost), color = met_slate, size = 2) +
  geom_segment(aes(x = q, xend = q, y = not_p, yend = p), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = 0, xend = q, y = p, yend = p), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = 0, xend = q, y = cost, yend = cost), color = met_slate, linetype = "dashed", size = 1) +
  geom_point(aes(x = q_c, y = p_c), color = met_slate, size = 2) +
  geom_point(aes(x = q_c, y = cost_2), color = met_slate, size = 2) +
  geom_segment(aes(x = q_c, xend = q_c, y = 0, yend = cost_2), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = 0, xend = q_c, y = cost_2, yend = cost_2), color = met_slate, linetype = "dashed", size = 1)
```
]

.less-right[
## Price Ceiling

Some natural monopolies are not profitable under marginal cost pricing.

Losses <br> .mono[-->] monopolist exits <br> .mono[-->] market ceases to exist <br> .mono[-->] even less total surplus than before!
]
































