---
title: "Final Review"
subtitle: "EC 201: Principles of Microeconomics"
author: "Kyle Raze"
date: "Winter 2020"
output:
  xaringan::moon_reader:
    css: ['default', 'metropolis', 'metropolis-fonts', 'my-css.css']
    self_contained: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "https://platform.twitter.com/widgets.js"
---
class: inverse, middle

```{r Setup, include = F}
options(htmltools.dir.version = FALSE)
library(pacman)
p_load(ggthemes, readxl, viridis, knitr, dslabs, gapminder, extrafont, Ecdat, wooldridge, tidyverse, magrittr, janitor, kableExtra, scales)
# Define colors
red_pink <- "#e64173"
met_slate <- "#272822" # metropolis font color 
purple <- "#9370DB"
green <- "#007935"
light_green <- "#7DBA97"
orange <- "#FD5F00"
turquoise <- "#44C1C4"
red <- "#b92e34"
# Notes directory
dir_slides <- "~/GitHub/Intro_Microeconomics_W20/Lectures/03-Consumer_Theory/"
# Knitr options
opts_chunk$set(
  comment = "#>",
  fig.align = "center",
  fig.height = 7,
  fig.width = 10.5,
  #dpi = 300,
  #cache = T,
  warning = F,
  message = F
)  
theme_simple <- theme_bw() + theme(
  axis.line = element_line(color = met_slate),
  panel.grid = element_blank(),
  rect = element_blank(),
  strip.text = element_blank(),
  text = element_text(family = "Fira Sans", color = met_slate, size = 17),
  # axis.text.x = element_text(size = 12),
  # axis.text.y = element_text(size = 12),
  axis.ticks = element_blank()
)
theme_market <- theme_bw() + theme(
  axis.line = element_line(color = met_slate),
  panel.grid = element_blank(),
  rect = element_blank(),
  strip.text = element_blank(),
  text = element_text(family = "Fira Sans", color = met_slate, size = 17),
  axis.title.x = element_text(hjust = 1, size = 17),
  axis.title.y = element_text(hjust = 1, angle = 0, size = 17),
  axis.ticks = element_blank()
)
theme_gif <- theme_bw() + theme(
  axis.line = element_line(color = met_slate),
  panel.grid = element_blank(),
  rect = element_blank(),
  text = element_text(family = "Fira Sans", color = met_slate, size = 17),
  axis.text.x = element_text(size = 12),
  axis.text.y = element_text(size = 12),
  axis.ticks = element_blank()
)
theme_goods <- theme_bw() + theme(
  axis.line = element_line(),
  panel.grid = element_blank(),
  rect = element_rect(size = 1),
  strip.text = element_blank(),
  text = element_text(size = 17, family = "Fira Sans"), 
  axis.text.x = element_text(size = 17),
  axis.text.y = element_text(size = 17, angle = 90, hjust = 0.5),
  axis.title.x = element_text(color = purple, size = 19),
  axis.title.y = element_text(color = red_pink, size = 19),
  axis.ticks = element_blank(),
  plot.title = element_blank(),
  legend.position = "none"
)
theme_game <- theme_bw() + theme(
  axis.line = element_line(),
  panel.grid = element_blank(),
  rect = element_rect(size = 1),
  strip.text = element_blank(),
  text = element_text(size = 20, family = "Fira Sans"), 
  axis.text.x = element_text(size = 20, color = met_slate),
  axis.text.y = element_text(size = 20, angle = 90, color = met_slate, hjust = 0.5),
  axis.title.x = element_text(color = met_slate, size = 20),
  axis.title.y = element_text(color = met_slate, size = 20),
  axis.ticks = element_blank(),
  plot.title = element_blank(),
  legend.position = "none"
)
theme_histogram <- theme_bw() + theme(
  axis.line = element_line(color = met_slate),
  panel.border = element_blank(),
  panel.grid.minor = element_blank(),
  plot.caption = element_text(hjust = 0),
  text = element_text(family = "Fira Sans", color = met_slate, size = 20),
  legend.text = element_text(family = "Fira Sans", color = met_slate, size = 20)
)
shift_axis <- function(p, y=0){
  g <- ggplotGrob(p)
  dummy <- data.frame(y=y)
  ax <- g[["grobs"]][g$layout$name == "axis-b"][[1]]
  p + annotation_custom(grid::grobTree(ax, vp = grid::viewport(y=1, height=sum(ax$height))), 
                        ymax=y, ymin=y) +
    geom_hline(aes(yintercept=y), data = dummy, size = 0.5, color = met_slate) +
    theme(axis.text.x = element_blank(), 
          axis.ticks.x = element_blank())
}
wrapper <- function(x, ...) paste(strwrap(x, ...), collapse = "\n")
```

# Prologue

---
# Today

## Grades

- What do they mean?

- How does the curve work?

--

## Final Exam Review

- Topics, structure, protocol, preparation, *etc.*

--

## Q & A

- Ask me anything!

---
# Finals Week

## Office Hours

- Monday 10:00-12:00 in 412 PLC.

- Tuesday 14:00-16:00 in 412 PLC.

--

## Final Exam

- .pink[Wednesday 10:15-12:15] in Pacific 123.

--

- If you show up after 10:30, I won't let you take the exam!

---
class: inverse, middle

# Grades

---
class: clear-slide
.center[**Pre-Final Grades**]
```{r, echo = FALSE, fig.height = 7, dev = "svg"}
histogram_data <- read_csv("../../Emails/Data/grade_histogram_data_pre-final.csv")
histogram_data %>%
    ggplot(aes(x = s_combined_points, fill = s_letter_grade, color = s_letter_grade)) +
    scale_y_continuous(expand = c(0, 0), breaks = pretty_breaks()) +
    expand_limits(x = 0) +
    scale_x_continuous(expand = c(0, 0), limits = c(0, 100), breaks = pretty_breaks()) +
    geom_histogram(position = "identity", alpha = 0.65, binwidth = 1) +
    geom_hline(yintercept = 0) +
    scale_color_viridis(discrete = TRUE, option = "viridis") +
    scale_fill_viridis(discrete = TRUE, option = "viridis") +
    labs(x = "Midterm 1 + Midterm 2 Combined Score (out of 100)",
         y = "Number of Students", color = "", fill = "") +
    theme_histogram
```

---
# Grades

I will re-curve the class after the final.

```{R, echo = F, escape = F}
tab <- data.frame(
  v1 = c("Midterm Exam 1", "Midterm Exam 2", "Final Exam", "Participation", "Optional Short Essays"),
  v2 = c("25%", "25%", "40%", "10%", "up to 4% extra credit")
) %>% kable(
  escape = F,
  col.names = c("Assignment Group", "Weight"),
  align = c("l", "c")
) %>%
  kable_styling(full_width = T) %>% 
  row_spec(1:5, extra_css = "line-height: 110%;")
tab 
```

--

**Note.sub[1]:** I have dropped your lowest three paricipation scores.

--

**Note.sub[2]:** The "Total" percentage you see in Canvas is misleading until everything is graded.

---
class: inverse, middle

# Final Exam

---
# Exam Topics

**Cumulative:** Anything we've covered is fair game.

- Midterm 1 material: weeks 1-4

- Midterm 2 material: weeks 5-7.

- New material: weeks 8-10.

---
# Exam Topics

## 1. What is Economics?

**A:** The study of how humans coordinate their wants.

- Why do we need to coordinate? .pink[Scarcity!]

--

What role should economists play?

- Positive *vs.* normative statements.

--

People respond to incentives.

---
# Exam Topics

## 2. Motivating the Economic Problem

Problem? .pink[Scarcity.]

--

Organizing an economy

- **Our focus:** Capitalism.

- **Consequences:** Improved living standards .mono[+] higher inequality.

--

What is a price?

- Know the punchline!

--

Opportunity cost

---
# Exam Topics

## 3. Consumer Theory I

Total *vs.* marginal value

- *e.g.,* Diamond-Water Paradox.

--

Diminishing marginal value .mono[-->] downward-sloping demand.

--

Optimal purchase rule

- Maximizes an individual consumer's surplus.

--

Demand *vs.* quantity demanded

---
# Exam Topics

## 4. Consumer Theory II

Individual responsiveness .mono[=] price elasticity of demand.

- Relationship with revenue.

--

Determinants of demand .mono[=] demand shifters.

- Prices of other goods (substitutes *vs.* complements).

- Income (normal *vs.* inferior goods).

- Changes in tastes, expectations, *etc.*

--

Relative prices .mono[-->] relative demand.

- *e.g.,* shipping the good apples out.


---
# Exam Topics

## 5. The Market Mechanism

Pit market activity!

- Number on card .mono[=] marginal value.

- Basis of trade? Differences in marginal values.

--

Shortage and surplus *vs.* equilibrium

--

**Q:** Why do we like markets as a means to distribute goods? 

- **A:** They can maximize mutual gains .mono[=] efficiency!

---
# Exam Topics

## 6. Demand and Supply

Determinants of (market) demand .mono[=] demand shifters.

- Prices of other goods, income, tastes, expectations, number of buyers.

--

Determinants of supply .mono[=] supply shifters.

- Input prices, production technology, expectations, number of sellers, supply disruptions.

--

Simultaneous shifts .mono[-->] ambiguity.

- Know direction of change in price or quantity, but not both.

--

Market spillovers

---
# Exam Topics

## 7. Policy Levers: Taxes & Subsidies

Tax/subsidy wedge between consumer price and producer price.

- Government gets/pays the difference.

--

Tax incidence .mono[!=] statutory incidence.

- Inelastic side of the market bears more of the burden.

--

Elasticity .mono[-->] tax revenue .mono[+] deadweight loss

--

Per-unit *vs.* lump-sum

--

Equity: progressive *vs.* regressive taxes.

--

Equity-efficiency tradeoff

---
# Exam Topics

## 8. Policy Levers: Price Controls

Binding price floor .mono[-->] surplus of goods .mono[-->] deadweight loss.

- *e.g.,* agricultural price supports.

--

Binding price ceiling .mono[-->] shortage of goods .mono[-->] deadweight loss.

- *e.g.,* rent control.

--

What causes price controls to become binding?

- *e.g.,* how do price-gouging laws work?

---
# Exam Topics

## 9. How Economists Learn from Data I

**Q:** Why correlation .mono[!=] causation?

- **A:** Selection bias.

--

Structured thinking about causal *vs.* contaminated comparisons

- Ideal/hypothetical data on potential outcomes

--

How do randomized control trials help us make *internally valid* comparisons?

--

What does .pink[treatment] .mono[-] .purple[control] tell us? 

--

When is a study *externally valid*?

---
# Exam Topics

## 10. How Economists Learn from Data II

Regressions!

- Interpret slope of regression line.

- Omitted variables bias.

--

Natural experiments

- Accidental randomized control trials (*e.g.,* Oregon Medicaid Experiment)

- Difference-in-differences (*e.g.,* Card and Krueger's minimum wage study)

--

Statistical significance

---
# Exam Topics

## 11. Externalities

Market failure .mono[-->] deadweight loss.

--

Negative externality .mono[-->] overproduction.

--

Positive externality .mono[-->] underproduction.

--

Solutions?

1. Pigouvian taxes/subsidies

2. Coase theorem

---
# Exam Topics

## 12. Game Theory

Game .mono[=] players .mono[+] strategies .mono[+] payoffs.

- *e.g.,* Prisoner's dilemma, chicken, matching pennies, *etc.*

--

Dominant strategy is always a best response, but a best response isn't always a dominant strategy.

--

Nash equilibrium .mono[=] mutual best responses.

--

Make sure that you are comfortable solving a 2 .mono[&times;] 2 game!

- Don't worry about the brinkmanship material.

---
# Exam Topics

## 13. Public Goods

Types of goods

- Rival *vs.* non-rival, excludable *vs.* non-excludable.
- Private, public, club, and common-resource.

--

Markets fail with non-private goods. Why?

- Club goods .mono[-->] market power.
- Public goods and common resources .mono[-->] collective action problems.

--

Collective action problems

1. Free-rider problem
2. Tragedy of the commons

---
# Exam Topics

## 14. Producer Theory I

Accounting *vs.* economic profit

--

Production technology

--

Profit-maximizing hiring decisions

--

Fixed *vs.* variable costs

- ATC .mono[=] AVC .mono[+] AFC

--

Short-run *vs.* long-run costs

--

Returns to scale

---
# Exam Topics

## 15. Producer Theory II

Competitive markets

--

Profit maximization: pick Q where MR .mono[=] MC.

- Short-run: make decision to shut down or keep producing.

- Long-run: make entry/exit decisions.

--

Firm-level response to a change in demand and short *vs.* long-run market adjustment.

- Horizontal long-run supply curve at minimum ATC.

- Upward-sloping short-run supply curve.


---
# Exam Topics

## 16. Monopoly & Antitrust

Sources of monopoly power

--

Profit maximization: two steps!

--

Social consequences

--

Antitrust law .mono[+] regulation

---
# Exam Structure

## Multiple Choice

- 80 questions

- 1 point per question

- Multiple groups of sequential questions (*e.g.,* about graphs or tables)

---
# Exam Protocol

## Materials

- Writing utensil

- 3-inch-by-5-inch note card

- Basic or scientific calculator (no graphing or programming capabilities)

- UO ID card

- .hi[Nothing else]

---
# Exam Protocol

## Procedure

- **Randomized** seating chart

- 120 minutes from *"you may begin"* to *"pencils down"*

- First 30 minutes: .hi[quiet period] (no questions, no getting up)

- Last 90 minutes: ask lots of questions

- Show your UO ID card as you turn-in your exam

---
# Exam Preparation

## Materials

1. Lecture slides

2. Discussion worksheets

3. Midterm exams

4. Practice questions for the midterms and the final

5. Practice problems from the book

---
# Exam Preparation

## Strategies

1. .pink[**Review**] concepts .pink[.mono[<--] slides .mono[+] book]

2. .pink[**Dissect**] midterm performance .pink[.mono[<--] keys .mono[+] emails]

3. .pink[**Practice**] .pink[.mono[<--] worksheets .mono[+] book problems]

4. .pink[**Simulate**] the final .pink[.mono[<--] 25 .mono[+] 25 .mono[+] 25 .mono[=] 75 practice questions!]

    - Set a timer!

5. .purple[**Visit**] office hours

    - M 10:00-12:00 and T 14:00-16:00 in 412 PLC

6. .purple[**Prepare**] your note card


---
class: inverse, middle

# Q & A