---
title: "Game Theory"
subtitle: "EC 201: Principles of Microeconomics"
author: "Kyle Raze"
date: "Winter 2020"
output:
  xaringan::moon_reader:
    css: ['default', 'metropolis', 'metropolis-fonts', 'my-css.css']
    self_contained: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "https://platform.twitter.com/widgets.js"
---
class: inverse, middle

```{r Setup, include = F}
options(htmltools.dir.version = FALSE)
library(pacman)
p_load(ggthemes, readxl, viridis, knitr, dslabs, gapminder, extrafont, Ecdat, wooldridge, tidyverse, magrittr, janitor, kableExtra)
# Define colors
red_pink <- "#e64173"
met_slate <- "#272822" # metropolis font color 
purple <- "#9370DB"
green <- "#007935"
light_green <- "#7DBA97"
orange <- "#FD5F00"
turquoise <- "#44C1C4"
red <- "#b92e34"
# Notes directory
dir_slides <- "~/GitHub/Intro_Microeconomics_W20/Lectures/03-Consumer_Theory/"
# Knitr options
opts_chunk$set(
  comment = "#>",
  fig.align = "center",
  fig.height = 7,
  fig.width = 10.5,
  #dpi = 300,
  #cache = T,
  warning = F,
  message = F
)  
theme_simple <- theme_bw() + theme(
  axis.line = element_line(color = met_slate),
  panel.grid = element_blank(),
  rect = element_blank(),
  strip.text = element_blank(),
  text = element_text(family = "Fira Sans", color = met_slate, size = 17),
  # axis.text.x = element_text(size = 12),
  # axis.text.y = element_text(size = 12),
  axis.ticks = element_blank()
)
theme_market <- theme_bw() + theme(
  axis.line = element_line(color = met_slate),
  panel.grid = element_blank(),
  rect = element_blank(),
  strip.text = element_blank(),
  text = element_text(family = "Fira Sans", color = met_slate, size = 17),
  axis.title.x = element_text(hjust = 1, size = 17),
  axis.title.y = element_text(hjust = 1, angle = 0, size = 17),
  # axis.text.x = element_text(size = 12),
  # axis.text.y = element_text(size = 12),
  axis.ticks = element_blank()
)
theme_gif <- theme_bw() + theme(
  axis.line = element_line(color = met_slate),
  panel.grid = element_blank(),
  rect = element_blank(),
  text = element_text(family = "Fira Sans", color = met_slate, size = 17),
  axis.text.x = element_text(size = 12),
  axis.text.y = element_text(size = 12),
  axis.ticks = element_blank()
)
theme_game <- theme_bw() + theme(
  axis.line = element_line(),
  panel.grid = element_blank(),
  rect = element_rect(size = 1),
  strip.text = element_blank(),
  text = element_text(size = 17, family = "Fira Sans"), 
  axis.text.x = element_text(size = 17, color = purple),
  axis.text.y = element_text(size = 17, angle = 90, color = red_pink, hjust = 0.5),
  axis.title.x = element_text(color = purple, size = 19),
  axis.title.y = element_text(color = red_pink, size = 19),
  axis.ticks = element_blank(),
  plot.title = element_blank(),
  legend.position = "none"
)
game_box_1 <- function(x) 5 - x
game_box_2 <- function(x) 10 - x
game_box_3 <- function(x) 15 - x
w <- seq(0, 5, 0.01)
z <- seq(5, 10, 0.01)
wrapper <- function(x, ...) paste(strwrap(x, ...), collapse = "\n")
```

# Prologue

---
# Agenda

## Game Theory

- How can we model strategic interactions?

- Last topic for Midterm 2.

## Midterm 2 Overview

- One week from today!

---
class: inverse, middle

# Game Theory

---
# Game Theory

## What is it?

> A framework of analysis in which two or more players compete for payoffs.

--

A player's payoff depends on

1. Her decisions.

2. All other players' decisions.

--

Useful for understanding a variety of "games" people play.

- .pink[Competitive scenarios:] Politics, collective bargaining, war, sports, *etc.*

- .purple[Cooperative scenarios:] Resource management, public goods, *etc.*

---
# What defines a game?

.pink[Who are the **players**?]

.smallest[
- Pete Buttigieg and Bernie Sanders.
- Saudi Arabia, Venezuela, and other OPEC countries.
- USA and USSR.
]

--

.purple[What are the **strategies** available to each player?]

.smallest[
- Focus on early primary states or Super Tuesday states.
- Collude (restrict oil output) or compete (increase oil output).
- Develop nuclear arsenal or not.
]

--

.green[What are the **payoffs** associated with each combination of strategies?]

.smallest[
- Democratic nominee or just another rank-and-file progressive.
- Split higher revenues or lower revenues.
- Continue to exist or total annihilation.
]

---
# The Prisoner's Dilemma

.pull-left[
```{r, echo = FALSE, fig.height = 4.5, fig.width = 4.5, dev = "svg"}
strategies_p1 <- c("Accuse", "Deny")
strategies_p2 <- c("Deny", "Accuse")

pd_1 <- ggplot() + 
  scale_x_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p2, position = "top")+
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p1) +
  labs(x = "Prisoner 2", y = "Prisoner 1") +
  # Player 1 payoffs
  annotate("text", label = wrapper(expression("Short Jail Sentence"), 12), x = 1.75, y = 6, family = "Fira Sans", size = 5, color = red_pink) +
  annotate("text", label = wrapper(expression("Longest Jail Sentence"), 14), x = 6.75, y = 6, family = "Fira Sans", size = 5, color = red_pink) +
  annotate("text", label = wrapper(expression("No Jail Sentence"), 12), x = 1.75, y = 1, family = "Fira Sans", size = 5, color = red_pink) +
  annotate("text", label = wrapper(expression("Long Jail Sentence"), 12), x = 6.75, y = 1, family = "Fira Sans", size = 5, color = red_pink) +
  # Player 2 payoffs
  annotate("text", label = wrapper(expression("Short Jail Sentence"), 12), x = 3.25, y = 9, family = "Fira Sans", size = 5, color = purple) +
  annotate("text", label = wrapper(expression("No Jail Sentence"), 12), x = 8.25, y = 9, family = "Fira Sans", size = 5, color = purple) +
  annotate("text", label = wrapper(expression("Longest Jail Sentence"), 14), x = 3.25, y = 4, family = "Fira Sans", size = 5, color = purple) +
  annotate("text", label = wrapper(expression("Long Jail Sentence"), 12), x = 8.25, y = 4, family = "Fira Sans", size = 5, color = purple) +
  # theme stuff
  theme_game +
  geom_segment(aes(x = 0, y = 10, xend = 10, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 10, xend = 10, yend = 5), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 0, y = 5, xend = 5, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 0, xend = 5, yend = 10), linetype  = "solid", size = 1) +
  geom_segment(aes(x = 0, y = 5, xend = 10, yend = 5), linetype  = "solid", size = 1) +
  geom_ribbon(aes(x = w, ymin = 0, ymax = game_box_1(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = 5, ymax = game_box_2(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 5, ymax = game_box_3(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 0, ymax = game_box_2(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_1(w), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_2(w), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_3(z), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_2(z), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank")
pd_1
```
]

.pull-right[

## Story

An interrogator keeps .pink[Prisoner 1] and .purple[Prisoner 2] in separate cells and interviews them separately.

Each prisoner can either deny the charges or accuse the other prisoner.

]

---
# The Prisoner's Dilemma

.pull-left[
```{r, echo = FALSE, fig.height = 4.5, fig.width = 4.5, dev = "svg"}
pd_1
```
]

.pull-right[

## Analysis

**Step 1:** Quantify the payoffs.

]

---
count: false
# The Prisoner's Dilemma

.pull-left[
```{r, echo = FALSE, fig.height = 4.5, fig.width = 4.5, dev = "svg"}
strategies_p1 <- c("Accuse", "Deny")
strategies_p2 <- c("Deny", "Accuse")

pd_2 <- ggplot() + 
  scale_x_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p2, position = "top")+
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p1) +
  labs(x = "Prisoner 2", y = "Prisoner 1") +
  # Player 1 payoffs
  annotate("text", label = expression(-3), x = 1.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(-12), x = 6.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(0), x = 1.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(-8), x = 6.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  # Player 2 payoffs
  annotate("text", label = expression(-3), x = 3.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression(0), x = 8.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression(-12), x = 3.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression(-8), x = 8.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  # theme stuff
  theme_game +
  geom_segment(aes(x = 0, y = 10, xend = 10, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 10, xend = 10, yend = 5), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 0, y = 5, xend = 5, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 0, xend = 5, yend = 10), linetype  = "solid", size = 1) +
  geom_segment(aes(x = 0, y = 5, xend = 10, yend = 5), linetype  = "solid", size = 1) +
  geom_ribbon(aes(x = w, ymin = 0, ymax = game_box_1(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = 5, ymax = game_box_2(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 5, ymax = game_box_3(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 0, ymax = game_box_2(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_1(w), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_2(w), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_3(z), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_2(z), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank")
pd_2 
```
]

.pull-right[

## Analysis

**Step 1:** Quantify the payoffs.

]

---
count: false
# The Prisoner's Dilemma

.pull-left[
```{r, echo = FALSE, fig.height = 4.5, fig.width = 4.5, dev = "svg"}
pd_2
```
]

.pull-right[

## Analysis

**Step 1:** Quantify the payoffs.

**Step 2:** Find each prisoner's *best response* for each strategy the other could play.

]

---
count: false
# The Prisoner's Dilemma

.pull-left[
```{r, echo = FALSE, fig.height = 4.5, fig.width = 4.5, dev = "svg"}
strategies_p1 <- c("Accuse", "Deny")
strategies_p2 <- c("Deny", "Accuse")

ggplot() + 
  scale_x_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p2, position = "top")+
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p1) +
  labs(x = "Prisoner 2", y = "Prisoner 1") +
  # Player 1 payoffs
  annotate("text", label = expression(-3), x = 1.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(-12), x = 6.75, y = 6, family = "Fira Sans", size = 6, color = red_pink, alpha = 0.2) +
  annotate("text", label = expression(0), x = 1.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(-8), x = 6.75, y = 1, family = "Fira Sans", size = 6, color = red_pink, alpha = 0.2) +
  # Player 2 payoffs
  annotate("text", label = expression(-3), x = 3.25, y = 9, family = "Fira Sans", size = 6, color = purple, alpha = 0.2) +
  annotate("text", label = expression(0), x = 8.25, y = 9, family = "Fira Sans", size = 6, color = purple, alpha = 0.2) +
  annotate("text", label = expression(-12), x = 3.25, y = 4, family = "Fira Sans", size = 6, color = purple, alpha = 0.2) +
  annotate("text", label = expression(-8), x = 8.25, y = 4, family = "Fira Sans", size = 6, color = purple, alpha = 0.2) +
  # theme stuff
  theme_game +
  geom_segment(aes(x = 0, y = 10, xend = 10, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 10, xend = 10, yend = 5), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 0, y = 5, xend = 5, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 0, xend = 5, yend = 10), linetype  = "solid", size = 1) +
  geom_segment(aes(x = 0, y = 5, xend = 10, yend = 5), linetype  = "solid", size = 1) +
  geom_ribbon(aes(x = w, ymin = 0, ymax = game_box_1(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = 5, ymax = game_box_2(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 5, ymax = game_box_3(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 0, ymax = game_box_2(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_1(w), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_2(w), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_3(z), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_2(z), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank")

```
]

.pull-right[

## Analysis

**Step 1:** Quantify the payoffs.

**Step 2:** Find each prisoner's *best response* for each strategy the other could play.
]

---
count: false
# The Prisoner's Dilemma

.pull-left[
```{r, echo = FALSE, fig.height = 4.5, fig.width = 4.5, dev = "svg"}
strategies_p1 <- c("Accuse", "Deny")
strategies_p2 <- c("Deny", "Accuse")

ggplot() + 
  scale_x_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p2, position = "top")+
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p1) +
  labs(x = "Prisoner 2", y = "Prisoner 1") +
  # Player 1 payoffs
  annotate("text", label = expression(-3), x = 1.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(-12), x = 6.75, y = 6, family = "Fira Sans", size = 6, color = red_pink, alpha = 0.2) +
  annotate("label", label = expression(0), x = 1.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(-8), x = 6.75, y = 1, family = "Fira Sans", size = 6, color = red_pink, alpha = 0.2) +
  # Player 2 payoffs
  annotate("text", label = expression(-3), x = 3.25, y = 9, family = "Fira Sans", size = 6, color = purple, alpha = 0.2) +
  annotate("text", label = expression(0), x = 8.25, y = 9, family = "Fira Sans", size = 6, color = purple, alpha = 0.2) +
  annotate("text", label = expression(-12), x = 3.25, y = 4, family = "Fira Sans", size = 6, color = purple, alpha = 0.2) +
  annotate("text", label = expression(-8), x = 8.25, y = 4, family = "Fira Sans", size = 6, color = purple, alpha = 0.2) +
  # theme stuff
  theme_game +
  geom_segment(aes(x = 0, y = 10, xend = 10, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 10, xend = 10, yend = 5), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 0, y = 5, xend = 5, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 0, xend = 5, yend = 10), linetype  = "solid", size = 1) +
  geom_segment(aes(x = 0, y = 5, xend = 10, yend = 5), linetype  = "solid", size = 1) +
  geom_ribbon(aes(x = w, ymin = 0, ymax = game_box_1(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = 5, ymax = game_box_2(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 5, ymax = game_box_3(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 0, ymax = game_box_2(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_1(w), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_2(w), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_3(z), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_2(z), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank")

```
]

.pull-right[

## Analysis

**Step 1:** Quantify the payoffs.

**Step 2:** Find each prisoner's *best response* for each strategy the other could play.
]

---
count: false
# The Prisoner's Dilemma

.pull-left[
```{r, echo = FALSE, fig.height = 4.5, fig.width = 4.5, dev = "svg"}
strategies_p1 <- c("Accuse", "Deny")
strategies_p2 <- c("Deny", "Accuse")

ggplot() + 
  scale_x_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p2, position = "top")+
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p1) +
  labs(x = "Prisoner 2", y = "Prisoner 1") +
  # Player 1 payoffs
  annotate("text", label = expression(-3), x = 1.75, y = 6, family = "Fira Sans", size = 6, color = red_pink, alpha = 0.2) +
  annotate("text", label = expression(-12), x = 6.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(0), x = 1.75, y = 1, family = "Fira Sans", size = 6, color = red_pink, alpha = 0.2) +
  annotate("label", label = expression(-8), x = 6.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  # Player 2 payoffs
  annotate("text", label = expression(-3), x = 3.25, y = 9, family = "Fira Sans", size = 6, color = purple, alpha = 0.2) +
  annotate("text", label = expression(0), x = 8.25, y = 9, family = "Fira Sans", size = 6, color = purple, alpha = 0.2) +
  annotate("text", label = expression(-12), x = 3.25, y = 4, family = "Fira Sans", size = 6, color = purple, alpha = 0.2) +
  annotate("text", label = expression(-8), x = 8.25, y = 4, family = "Fira Sans", size = 6, color = purple, alpha = 0.2) +
  # theme stuff
  theme_game +
  geom_segment(aes(x = 0, y = 10, xend = 10, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 10, xend = 10, yend = 5), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 0, y = 5, xend = 5, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 0, xend = 5, yend = 10), linetype  = "solid", size = 1) +
  geom_segment(aes(x = 0, y = 5, xend = 10, yend = 5), linetype  = "solid", size = 1) +
  geom_ribbon(aes(x = w, ymin = 0, ymax = game_box_1(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = 5, ymax = game_box_2(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 5, ymax = game_box_3(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 0, ymax = game_box_2(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_1(w), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_2(w), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_3(z), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_2(z), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank")

```
]

.pull-right[

## Analysis

**Step 1:** Quantify the payoffs.

**Step 2:** Find each prisoner's *best response* for each strategy the other could play.
]

---
count: false
# The Prisoner's Dilemma

.pull-left[
```{r, echo = FALSE, fig.height = 4.5, fig.width = 4.5, dev = "svg"}
strategies_p1 <- c("Accuse", "Deny")
strategies_p2 <- c("Deny", "Accuse")

ggplot() + 
  scale_x_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p2, position = "top")+
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p1) +
  labs(x = "Prisoner 2", y = "Prisoner 1") +
  # Player 1 payoffs
  annotate("text", label = expression(-3), x = 1.75, y = 6, family = "Fira Sans", size = 6, color = red_pink, alpha = 0.2) +
  annotate("text", label = expression(-12), x = 6.75, y = 6, family = "Fira Sans", size = 6, color = red_pink, alpha = 0.2) +
  annotate("text", label = expression(0), x = 1.75, y = 1, family = "Fira Sans", size = 6, color = red_pink, alpha = 0.2) +
  annotate("text", label = expression(-8), x = 6.75, y = 1, family = "Fira Sans", size = 6, color = red_pink, alpha = 0.2) +
  # Player 2 payoffs
  annotate("text", label = expression(-3), x = 3.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("label", label = expression(0), x = 8.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression(-12), x = 3.25, y = 4, family = "Fira Sans", size = 6, color = purple, alpha = 0.2) +
  annotate("text", label = expression(-8), x = 8.25, y = 4, family = "Fira Sans", size = 6, color = purple, alpha = 0.2) +
  # theme stuff
  theme_game +
  geom_segment(aes(x = 0, y = 10, xend = 10, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 10, xend = 10, yend = 5), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 0, y = 5, xend = 5, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 0, xend = 5, yend = 10), linetype  = "solid", size = 1) +
  geom_segment(aes(x = 0, y = 5, xend = 10, yend = 5), linetype  = "solid", size = 1) +
  geom_ribbon(aes(x = w, ymin = 0, ymax = game_box_1(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = 5, ymax = game_box_2(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 5, ymax = game_box_3(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 0, ymax = game_box_2(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_1(w), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_2(w), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_3(z), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_2(z), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank")

```
]

.pull-right[

## Analysis

**Step 1:** Quantify the payoffs.

**Step 2:** Find each prisoner's *best response* for each strategy the other could play.
]

---
count: false
# The Prisoner's Dilemma

.pull-left[
```{r, echo = FALSE, fig.height = 4.5, fig.width = 4.5, dev = "svg"}
strategies_p1 <- c("Accuse", "Deny")
strategies_p2 <- c("Deny", "Accuse")

ggplot() + 
  scale_x_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p2, position = "top")+
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p1) +
  labs(x = "Prisoner 2", y = "Prisoner 1") +
  # Player 1 payoffs
  annotate("text", label = expression(-3), x = 1.75, y = 6, family = "Fira Sans", size = 6, color = red_pink, alpha = 0.2) +
  annotate("text", label = expression(-12), x = 6.75, y = 6, family = "Fira Sans", size = 6, color = red_pink, alpha = 0.2) +
  annotate("text", label = expression(0), x = 1.75, y = 1, family = "Fira Sans", size = 6, color = red_pink, alpha = 0.2) +
  annotate("text", label = expression(-8), x = 6.75, y = 1, family = "Fira Sans", size = 6, color = red_pink, alpha = 0.2) +
  # Player 2 payoffs
  annotate("text", label = expression(-3), x = 3.25, y = 9, family = "Fira Sans", size = 6, color = purple, alpha = 0.2) +
  annotate("text", label = expression(0), x = 8.25, y = 9, family = "Fira Sans", size = 6, color = purple, alpha = 0.2) +
  annotate("text", label = expression(-12), x = 3.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  annotate("label", label = expression(-8), x = 8.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  # theme stuff
  theme_game +
  geom_segment(aes(x = 0, y = 10, xend = 10, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 10, xend = 10, yend = 5), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 0, y = 5, xend = 5, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 0, xend = 5, yend = 10), linetype  = "solid", size = 1) +
  geom_segment(aes(x = 0, y = 5, xend = 10, yend = 5), linetype  = "solid", size = 1) +
  geom_ribbon(aes(x = w, ymin = 0, ymax = game_box_1(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = 5, ymax = game_box_2(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 5, ymax = game_box_3(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 0, ymax = game_box_2(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_1(w), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_2(w), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_3(z), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_2(z), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank")

```
]

.pull-right[

## Analysis

**Step 1:** Quantify the payoffs.

**Step 2:** Find each prisoner's *best response* for each strategy the other could play.
]

---
count: false
# The Prisoner's Dilemma

.pull-left[
```{r, echo = FALSE, fig.height = 4.5, fig.width = 4.5, dev = "svg"}
strategies_p1 <- c("Accuse", "Deny")
strategies_p2 <- c("Deny", "Accuse")

ggplot() + 
  scale_x_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p2, position = "top")+
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p1) +
  labs(x = "Prisoner 2", y = "Prisoner 1") +
  # Player 1 payoffs
  annotate("text", label = expression(-3), x = 1.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(-12), x = 6.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("label", label = expression(0), x = 1.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("label", label = expression(-8), x = 6.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  # Player 2 payoffs
  annotate("text", label = expression(-3), x = 3.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("label", label = expression(0), x = 8.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression(-12), x = 3.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  annotate("label", label = expression(-8), x = 8.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  # theme stuff
  theme_game +
  geom_segment(aes(x = 0, y = 10, xend = 10, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 10, xend = 10, yend = 5), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 0, y = 5, xend = 5, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 0, xend = 5, yend = 10), linetype  = "solid", size = 1) +
  geom_segment(aes(x = 0, y = 5, xend = 10, yend = 5), linetype  = "solid", size = 1) +
  geom_ribbon(aes(x = w, ymin = 0, ymax = game_box_1(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = 5, ymax = game_box_2(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 5, ymax = game_box_3(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 0, ymax = game_box_2(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_1(w), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_2(w), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_3(z), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_2(z), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank")

```
]

.pull-right[

## Analysis

**Step 1:** Quantify the payoffs.

**Step 2:** Find each prisoner's *best response* for each strategy the other could play.

- Both players have a *dominant strategy* to accuse.
]

---
count: false
# The Prisoner's Dilemma

.pull-left[
```{r, echo = FALSE, fig.height = 4.5, fig.width = 4.5, dev = "svg"}
strategies_p1 <- c("Accuse", "Deny")
strategies_p2 <- c("Deny", "Accuse")

ggplot() + 
  scale_x_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p2, position = "top")+
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p1) +
  labs(x = "Prisoner 2", y = "Prisoner 1") +
  # Player 1 payoffs
  annotate("text", label = expression(-3), x = 1.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(-12), x = 6.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("label", label = expression(0), x = 1.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("label", label = expression(-8), x = 6.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  # Player 2 payoffs
  annotate("text", label = expression(-3), x = 3.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("label", label = expression(0), x = 8.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression(-12), x = 3.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  annotate("label", label = expression(-8), x = 8.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  # theme stuff
  theme_game +
  geom_segment(aes(x = 0, y = 10, xend = 10, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 10, xend = 10, yend = 5), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 0, y = 5, xend = 5, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 0, xend = 5, yend = 10), linetype  = "solid", size = 1) +
  geom_segment(aes(x = 0, y = 5, xend = 10, yend = 5), linetype  = "solid", size = 1) +
  geom_ribbon(aes(x = w, ymin = 0, ymax = game_box_1(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = 5, ymax = game_box_2(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 5, ymax = game_box_3(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 0, ymax = game_box_2(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_1(w), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_2(w), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_3(z), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_2(z), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank")

```
]

.pull-right[

## Analysis

**Step 1:** Quantify the payoffs.

**Step 2:** Find each prisoner's *best response* for each strategy the other could play.

- Both players have a *dominant strategy* to accuse.

**Step 3:** Find the *Nash equilibrium*.

- Equilibrium .mono[=] (.pink[Accuse], .purple[Accuse])
]

---
# Nash Equilibrium

## Definition

>  A set of strategies such that no player has the incentive to deviate unilaterally from her chosen strategy.

--

**Q:** What logic supports the (.pink[Accuse], .purple[Accuse]) equilibrium?

--

- Given that .pink[Prisoner 1] decided to .pink[Accuse], .purple[Prisoner 2] is better off having played .purple[Accuse].

- Given that .purple[Prisoner 2] decided to .purple[Accuse], .pink[Prisoner 1] is better off having played .pink[Accuse].

---
class: clear-slide

**Q:** Cooperate or defect?

- If everyone chooses to cooperate, then everyone will receive 2 bonus participation points.

- If everyone chooses to cooperate, but one person defects, then the person who defects will receive 6 bonus participation points and everyone else will receive 0 bonus participation points.

- If more than one person chooses to defect, then everyone will receive 0 bonus participation points.

> **A.** Cooperate.

> **B.** Defect.

---
# Cournot Game

.pull-left[
```{r, echo = FALSE, fig.height = 4.5, fig.width = 4.5, dev = "svg"}
strategies_p1 <- c("Low Price", "High Price")
strategies_p2 <- c("High Price", "Low Price")

ggplot() + 
  scale_x_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p2, position = "top")+
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p1) +
  labs(x = "Nike", y = "Adidas") +
  # Player 1 payoffs
  annotate("text", label = wrapper(expression("High Profits"), 12), x = 1.75, y = 6, family = "Fira Sans", size = 5, color = red_pink) +
  annotate("text", label = wrapper(expression("Lowest Profits"), 14), x = 6.75, y = 6, family = "Fira Sans", size = 5, color = red_pink) +
  annotate("text", label = wrapper(expression("Highest Profits"), 12), x = 1.75, y = 1, family = "Fira Sans", size = 5, color = red_pink) +
  annotate("text", label = wrapper(expression("Low Profits"), 12), x = 6.75, y = 1, family = "Fira Sans", size = 5, color = red_pink) +
  # Player 2 payoffs
  annotate("text", label = wrapper(expression("High Profits"), 12), x = 3.25, y = 9, family = "Fira Sans", size = 5, color = purple) +
  annotate("text", label = wrapper(expression("Highest Profits"), 12), x = 8.25, y = 9, family = "Fira Sans", size = 5, color = purple) +
  annotate("text", label = wrapper(expression("Lowest Profits"), 14), x = 3.25, y = 4, family = "Fira Sans", size = 5, color = purple) +
  annotate("text", label = wrapper(expression("Low Profits"), 12), x = 8.25, y = 4, family = "Fira Sans", size = 5, color = purple) +
  # theme stuff
  theme_game +
  geom_segment(aes(x = 0, y = 10, xend = 10, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 10, xend = 10, yend = 5), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 0, y = 5, xend = 5, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 0, xend = 5, yend = 10), linetype  = "solid", size = 1) +
  geom_segment(aes(x = 0, y = 5, xend = 10, yend = 5), linetype  = "solid", size = 1) +
  geom_ribbon(aes(x = w, ymin = 0, ymax = game_box_1(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = 5, ymax = game_box_2(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 5, ymax = game_box_3(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 0, ymax = game_box_2(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_1(w), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_2(w), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_3(z), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_2(z), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank")
```
]

.pull-right[

**Q:** What are the equilibrium strategies?

]

---
count: false
# Cournot Game

.pull-left[
```{r, echo = FALSE, fig.height = 4.5, fig.width = 4.5, dev = "svg"}
strategies_p1 <- c("Low Price", "High Price")
strategies_p2 <- c("High Price", "Low Price")

ggplot() + 
  scale_x_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p2, position = "top")+
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p1) +
  labs(x = "Nike", y = "Adidas") +
  # Player 1 payoffs
  annotate("text", label = expression(35), x = 1.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(5), x = 6.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(50), x = 1.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(20), x = 6.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  # Player 2 payoffs
  annotate("text", label = expression(35), x = 3.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression(50), x = 8.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression(5), x = 3.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression(20), x = 8.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  # theme stuff
  theme_game +
  geom_segment(aes(x = 0, y = 10, xend = 10, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 10, xend = 10, yend = 5), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 0, y = 5, xend = 5, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 0, xend = 5, yend = 10), linetype  = "solid", size = 1) +
  geom_segment(aes(x = 0, y = 5, xend = 10, yend = 5), linetype  = "solid", size = 1) +
  geom_ribbon(aes(x = w, ymin = 0, ymax = game_box_1(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = 5, ymax = game_box_2(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 5, ymax = game_box_3(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 0, ymax = game_box_2(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_1(w), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_2(w), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_3(z), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_2(z), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank")
```
]

.pull-right[

**Q:** What are the equilibrium strategies?

]

---
count: false
# Cournot Game

.pull-left[
```{r, echo = FALSE, fig.height = 4.5, fig.width = 4.5, dev = "svg"}
strategies_p1 <- c("Low Price", "High Price")
strategies_p2 <- c("High Price", "Low Price")

ggplot() + 
  scale_x_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p2, position = "top")+
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p1) +
  labs(x = "Nike", y = "Adidas") +
  # Player 1 payoffs
  annotate("text", label = expression(35), x = 1.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(5), x = 6.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("label", label = expression(50), x = 1.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("label", label = expression(20), x = 6.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  # Player 2 payoffs
  annotate("text", label = expression(35), x = 3.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("label", label = expression(50), x = 8.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression(5), x = 3.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  annotate("label", label = expression(20), x = 8.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  # theme stuff
  theme_game +
  geom_segment(aes(x = 0, y = 10, xend = 10, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 10, xend = 10, yend = 5), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 0, y = 5, xend = 5, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 0, xend = 5, yend = 10), linetype  = "solid", size = 1) +
  geom_segment(aes(x = 0, y = 5, xend = 10, yend = 5), linetype  = "solid", size = 1) +
  geom_ribbon(aes(x = w, ymin = 0, ymax = game_box_1(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = 5, ymax = game_box_2(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 5, ymax = game_box_3(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 0, ymax = game_box_2(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_1(w), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_2(w), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_3(z), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_2(z), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank")
```
]

.pull-right[

**Q:** What are the equilibrium strategies?

**A:** Equilibrium .mono[=] (.pink[Low Price], .purple[Low Price]), even though a bilateral deviation would be mutually beneficial!

- Another dilemma!

]

---
# Prisoner's Dilemma 2.0

.pull-left[
```{r, echo = FALSE, fig.height = 4.5, fig.width = 4.5, dev = "svg"}
strategies_p1 <- c("Accuse", "Deny")
strategies_p2 <- c("Deny", "Accuse")

pd_1 <- ggplot() + 
  scale_x_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p2, position = "top")+
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p1) +
  labs(x = "Prisoner 2", y = "Prisoner 1") +
  # Player 1 payoffs
  annotate("text", label = wrapper(expression("Short Jail Sentence"), 12), x = 1.75, y = 6, family = "Fira Sans", size = 5, color = red_pink) +
  annotate("text", label = wrapper(expression("Longest Jail Sentence"), 14), x = 6.75, y = 6, family = "Fira Sans", size = 5, color = red_pink) +
  annotate("text", label = wrapper(expression("Compensate Prisoner 2"), 12), x = 1.75, y = 1, family = "Fira Sans", size = 5, color = red_pink) +
  annotate("text", label = wrapper(expression("Long Jail Sentence"), 12), x = 6.75, y = 1, family = "Fira Sans", size = 5, color = red_pink) +
  # Player 2 payoffs
  annotate("text", label = wrapper(expression("Short Jail Sentence"), 12), x = 3.25, y = 9, family = "Fira Sans", size = 5, color = purple) +
  annotate("text", label = wrapper(expression("Compensate Prisoner 1"), 12), x = 8.25, y = 9, family = "Fira Sans", size = 5, color = purple) +
  annotate("text", label = wrapper(expression("Longest Jail Sentence"), 14), x = 3.25, y = 4, family = "Fira Sans", size = 5, color = purple) +
  annotate("text", label = wrapper(expression("Long Jail Sentence"), 12), x = 8.25, y = 4, family = "Fira Sans", size = 5, color = purple) +
  # theme stuff
  theme_game +
  geom_segment(aes(x = 0, y = 10, xend = 10, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 10, xend = 10, yend = 5), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 0, y = 5, xend = 5, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 0, xend = 5, yend = 10), linetype  = "solid", size = 1) +
  geom_segment(aes(x = 0, y = 5, xend = 10, yend = 5), linetype  = "solid", size = 1) +
  geom_ribbon(aes(x = w, ymin = 0, ymax = game_box_1(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = 5, ymax = game_box_2(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 5, ymax = game_box_3(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 0, ymax = game_box_2(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_1(w), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_2(w), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_3(z), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_2(z), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank")
pd_1
```
]

.pull-right[

## Story

An interrogator keeps .pink[Prisoner 1] and .purple[Prisoner 2] in separate cells and interviews them separately.

Each prisoner can either deny the charges or accuse the other prisoner.

**Twist:** Pre-heist contract.

]

---
count: false
# Prisoner's Dilemma 2.0

.pull-left[
```{r, echo = FALSE, fig.height = 4.5, fig.width = 4.5, dev = "svg"}
strategies_p1 <- c("Accuse", "Deny")
strategies_p2 <- c("Deny", "Accuse")

ggplot() + 
  scale_x_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p2, position = "top")+
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p1) +
  labs(x = "Prisoner 2", y = "Prisoner 1") +
  # Player 1 payoffs
  annotate("text", label = expression(-3), x = 1.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(-12), x = 6.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(-4), x = 1.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(-8), x = 6.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  # Player 2 payoffs
  annotate("text", label = expression(-3), x = 3.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression(-4), x = 8.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression(-12), x = 3.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression(-8), x = 8.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  # theme stuff
  theme_game +
  geom_segment(aes(x = 0, y = 10, xend = 10, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 10, xend = 10, yend = 5), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 0, y = 5, xend = 5, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 0, xend = 5, yend = 10), linetype  = "solid", size = 1) +
  geom_segment(aes(x = 0, y = 5, xend = 10, yend = 5), linetype  = "solid", size = 1) +
  geom_ribbon(aes(x = w, ymin = 0, ymax = game_box_1(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = 5, ymax = game_box_2(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 5, ymax = game_box_3(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 0, ymax = game_box_2(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_1(w), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_2(w), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_3(z), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_2(z), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank")
```
]

.pull-right[

## Story

An interrogator keeps .pink[Prisoner 1] and .purple[Prisoner 2] in separate cells and interviews them separately.

Each prisoner can either deny the charges or accuse the other prisoner.

**Twist:** Pre-heist contract.

]

---
# Prisoner's Dilemma 2.0

.pull-left[
```{r, echo = FALSE, fig.height = 4.5, fig.width = 4.5, dev = "svg"}
strategies_p1 <- c("Accuse", "Deny")
strategies_p2 <- c("Deny", "Accuse")

ggplot() + 
  scale_x_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p2, position = "top")+
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p1) +
  labs(x = "Prisoner 2", y = "Prisoner 1") +
  # Player 1 payoffs
  annotate("text", label = expression(-3), x = 1.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(-12), x = 6.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(-4), x = 1.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(-8), x = 6.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  # Player 2 payoffs
  annotate("text", label = expression(-3), x = 3.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression(-4), x = 8.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression(-12), x = 3.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression(-8), x = 8.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  # theme stuff
  theme_game +
  geom_segment(aes(x = 0, y = 10, xend = 10, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 10, xend = 10, yend = 5), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 0, y = 5, xend = 5, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 0, xend = 5, yend = 10), linetype  = "solid", size = 1) +
  geom_segment(aes(x = 0, y = 5, xend = 10, yend = 5), linetype  = "solid", size = 1) +
  geom_ribbon(aes(x = w, ymin = 0, ymax = game_box_1(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = 5, ymax = game_box_2(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 5, ymax = game_box_3(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 0, ymax = game_box_2(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_1(w), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_2(w), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_3(z), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_2(z), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank")
```
]

.pull-right[

**Q:** What are the equilibrium strategies?

]

---
count: false
# Prisoner's Dilemma 2.0

.pull-left[
```{r, echo = FALSE, fig.height = 4.5, fig.width = 4.5, dev = "svg"}
strategies_p1 <- c("Accuse", "Deny")
strategies_p2 <- c("Deny", "Accuse")

ggplot() + 
  scale_x_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p2, position = "top")+
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p1) +
  labs(x = "Prisoner 2", y = "Prisoner 1") +
  # Player 1 payoffs
  annotate("label", label = expression(-3), x = 1.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(-12), x = 6.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(-4), x = 1.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("label", label = expression(-8), x = 6.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  # Player 2 payoffs
  annotate("label", label = expression(-3), x = 3.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression(-4), x = 8.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression(-12), x = 3.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  annotate("label", label = expression(-8), x = 8.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  # theme stuff
  theme_game +
  geom_segment(aes(x = 0, y = 10, xend = 10, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 10, xend = 10, yend = 5), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 0, y = 5, xend = 5, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 0, xend = 5, yend = 10), linetype  = "solid", size = 1) +
  geom_segment(aes(x = 0, y = 5, xend = 10, yend = 5), linetype  = "solid", size = 1) +
  geom_ribbon(aes(x = w, ymin = 0, ymax = game_box_1(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = 5, ymax = game_box_2(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 5, ymax = game_box_3(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 0, ymax = game_box_2(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_1(w), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_2(w), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_3(z), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_2(z), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank")
```
]

.pull-right[

**Q:** What are the equilibrium strategies?

**A:** There are two equilibria!

- (.pink[Accuse], .purple[Accuse])

- (.pink[Deny], .purple[Deny])

No longer a prisoner's dilemma.

- Now a coordination game.

]

---
class: clear-slide

<iframe width="900" height="600" src="https://www.youtube.com/embed/p3Uos2fzIJ0" frameborder="0" allow="accelerometer; autoplay; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

---
# Split or Steal?

.pull-left[
```{r, echo = FALSE, fig.height = 4.5, fig.width = 4.5, dev = "svg"}
strategies_p1 <- c("Steal", "Split")
strategies_p2 <- c("Split", "Steal")

ggplot() + 
  scale_x_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p2, position = "top")+
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p1) +
  labs(x = "Stephen", y = "Sarah") +
  # Player 1 payoffs
  annotate("text", label = expression("50k"), x = 1.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(0), x = 6.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression("100k"), x = 1.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(0), x = 6.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  # Player 2 payoffs
  annotate("text", label = expression("50k"), x = 3.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression("100k"), x = 8.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression(0), x = 3.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression(0), x = 8.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  # theme stuff
  theme_game +
  geom_segment(aes(x = 0, y = 10, xend = 10, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 10, xend = 10, yend = 5), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 0, y = 5, xend = 5, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 0, xend = 5, yend = 10), linetype  = "solid", size = 1) +
  geom_segment(aes(x = 0, y = 5, xend = 10, yend = 5), linetype  = "solid", size = 1) +
  geom_ribbon(aes(x = w, ymin = 0, ymax = game_box_1(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = 5, ymax = game_box_2(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 5, ymax = game_box_3(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 0, ymax = game_box_2(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_1(w), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_2(w), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_3(z), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_2(z), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank")
```
]

.pull-right[

**Observed outcome:** (.pink[Steal], .purple[Split])

**Q:** What are the equilibrium strategies?

]

---
count: false
# Split or Steal?

.pull-left[
```{r, echo = FALSE, fig.height = 4.5, fig.width = 4.5, dev = "svg"}
strategies_p1 <- c("Steal", "Split")
strategies_p2 <- c("Split", "Steal")

ggplot() + 
  scale_x_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p2, position = "top")+
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p1) +
  labs(x = "Stephen", y = "Sarah") +
  # Player 1 payoffs
  annotate("text", label = expression("50k"), x = 1.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("label", label = expression(0), x = 6.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("label", label = expression("100k"), x = 1.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("label", label = expression(0), x = 6.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  # Player 2 payoffs
  annotate("text", label = expression("50k"), x = 3.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("label", label = expression("100k"), x = 8.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("label", label = expression(0), x = 3.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  annotate("label", label = expression(0), x = 8.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  # theme stuff
  theme_game +
  geom_segment(aes(x = 0, y = 10, xend = 10, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 10, xend = 10, yend = 5), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 0, y = 5, xend = 5, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 0, xend = 5, yend = 10), linetype  = "solid", size = 1) +
  geom_segment(aes(x = 0, y = 5, xend = 10, yend = 5), linetype  = "solid", size = 1) +
  geom_ribbon(aes(x = w, ymin = 0, ymax = game_box_1(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = 5, ymax = game_box_2(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 5, ymax = game_box_3(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 0, ymax = game_box_2(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_1(w), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_2(w), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_3(z), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_2(z), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank")
```
]

.pull-right[

**Observed outcome:** (.pink[Steal], .purple[Split])

**Q:** What are the equilibrium strategies?

**A:** Three equilibria.

- (.pink[Steal], .purple[Split])

- (.pink[Split], .purple[Steal])

- (.pink[Steal], .purple[Steal])

]

---
# Chicken

.pull-left[
```{r, echo = FALSE, fig.height = 4.5, fig.width = 4.5, dev = "svg"}
strategies_p1 <- c("Straight", "Swerve")
strategies_p2 <- c("Swerve", "Straight")

ggplot() + 
  scale_x_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p2, position = "top")+
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p1) +
  labs(x = "Driver 2", y = "Driver 1") +
  # Player 1 payoffs
  annotate("text", label = expression("Tie"), x = 1.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression("Lose"), x = 6.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression("Win"), x = 1.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression("Crash"), x = 6.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  # Player 2 payoffs
  annotate("text", label = expression("Tie"), x = 3.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression("Win"), x = 8.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression("Lose"), x = 3.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression("Crash"), x = 8.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  # theme stuff
  theme_game +
  geom_segment(aes(x = 0, y = 10, xend = 10, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 10, xend = 10, yend = 5), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 0, y = 5, xend = 5, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 0, xend = 5, yend = 10), linetype  = "solid", size = 1) +
  geom_segment(aes(x = 0, y = 5, xend = 10, yend = 5), linetype  = "solid", size = 1) +
  geom_ribbon(aes(x = w, ymin = 0, ymax = game_box_1(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = 5, ymax = game_box_2(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 5, ymax = game_box_3(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 0, ymax = game_box_2(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_1(w), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_2(w), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_3(z), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_2(z), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank")
```
]

.pull-right[

**Q:** What are the equilibrium strategies?

]

---
count: false
# Chicken

.pull-left[
```{r, echo = FALSE, fig.height = 4.5, fig.width = 4.5, dev = "svg"}
strategies_p1 <- c("Straight", "Swerve")
strategies_p2 <- c("Swerve", "Straight")

ggplot() + 
  scale_x_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p2, position = "top")+
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p1) +
  labs(x = "Driver 2", y = "Driver 1") +
  # Player 1 payoffs
  annotate("text", label = expression("Tie"), x = 1.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("label", label = expression("Lose"), x = 6.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("label", label = expression("Win"), x = 1.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression("Crash"), x = 6.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  # Player 2 payoffs
  annotate("text", label = expression("Tie"), x = 3.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("label", label = expression("Win"), x = 8.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("label", label = expression("Lose"), x = 3.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression("Crash"), x = 8.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  # theme stuff
  theme_game +
  geom_segment(aes(x = 0, y = 10, xend = 10, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 10, xend = 10, yend = 5), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 0, y = 5, xend = 5, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 0, xend = 5, yend = 10), linetype  = "solid", size = 1) +
  geom_segment(aes(x = 0, y = 5, xend = 10, yend = 5), linetype  = "solid", size = 1) +
  geom_ribbon(aes(x = w, ymin = 0, ymax = game_box_1(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = 5, ymax = game_box_2(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 5, ymax = game_box_3(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 0, ymax = game_box_2(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_1(w), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_2(w), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_3(z), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_2(z), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank")
```
]

.pull-right[

**Q:** What are the equilibrium strategies

**A:** Two equilibria.

- (.pink[Swerve], .purple[Straight])

- (.pink[Straight], .purple[Swerve])

]

---
# Matching Pennies

.pull-left[
```{r, echo = FALSE, fig.height = 4.5, fig.width = 4.5, dev = "svg"}
strategies_p1 <- c("Tails", "Heads")
strategies_p2 <- c("Heads", "Tails")

ggplot() + 
  scale_x_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p2, position = "top")+
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p1) +
  labs(x = "Roommate 2", y = "Roommate 1") +
  # Player 1 payoffs
  annotate("text", label = expression(+1), x = 1.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(-1), x = 6.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(-1), x = 1.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(+1), x = 6.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  # Player 2 payoffs
  annotate("text", label = expression(-1), x = 3.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression(+1), x = 8.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression(+1), x = 3.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression(-1), x = 8.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  # theme stuff
  theme_game +
  geom_segment(aes(x = 0, y = 10, xend = 10, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 10, xend = 10, yend = 5), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 0, y = 5, xend = 5, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 0, xend = 5, yend = 10), linetype  = "solid", size = 1) +
  geom_segment(aes(x = 0, y = 5, xend = 10, yend = 5), linetype  = "solid", size = 1) +
  geom_ribbon(aes(x = w, ymin = 0, ymax = game_box_1(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = 5, ymax = game_box_2(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 5, ymax = game_box_3(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 0, ymax = game_box_2(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_1(w), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_2(w), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_3(z), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_2(z), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank")
```
]

.pull-right[

**Q:** What are the equilibrium strategies?

]

---
count: false
# Matching Pennies

.pull-left[
```{r, echo = FALSE, fig.height = 4.5, fig.width = 4.5, dev = "svg"}
strategies_p1 <- c("Tails", "Heads")
strategies_p2 <- c("Heads", "Tails")

ggplot() + 
  scale_x_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p2, position = "top")+
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p1) +
  labs(x = "Roommate 2", y = "Roommate 1") +
  # Player 1 payoffs
  annotate("label", label = expression(+1), x = 1.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(-1), x = 6.75, y = 6, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression(-1), x = 1.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("label", label = expression(+1), x = 6.75, y = 1, family = "Fira Sans", size = 6, color = red_pink) +
  # Player 2 payoffs
  annotate("text", label = expression(-1), x = 3.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("label", label = expression(+1), x = 8.25, y = 9, family = "Fira Sans", size = 6, color = purple) +
  annotate("label", label = expression(+1), x = 3.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  annotate("text", label = expression(-1), x = 8.25, y = 4, family = "Fira Sans", size = 6, color = purple) +
  # theme stuff
  theme_game +
  geom_segment(aes(x = 0, y = 10, xend = 10, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 10, xend = 10, yend = 5), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 0, y = 5, xend = 5, yend = 0), linetype  = "dashed", size = 0.5) +
  geom_segment(aes(x = 5, y = 0, xend = 5, yend = 10), linetype  = "solid", size = 1) +
  geom_segment(aes(x = 0, y = 5, xend = 10, yend = 5), linetype  = "solid", size = 1) +
  geom_ribbon(aes(x = w, ymin = 0, ymax = game_box_1(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = 5, ymax = game_box_2(w)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 5, ymax = game_box_3(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 0, ymax = game_box_2(z)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_1(w), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = w, ymin = game_box_2(w), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_3(z), ymax = 10), fill = purple, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = game_box_2(z), ymax = 5), fill = purple, alpha = 0.2, linetype = "blank")
```
]

.pull-right[

**Q:** What are the equilibrium strategies

**A:** No mutual best responses!

- The only equilibrium strategy is to randomize.

]

---
# Pascal's Wager

.pull-left[
```{r, echo = FALSE, fig.height = 5.5, fig.width = 4.5, dev = "svg"}
strategies_p1 <- c("Live as if No God", "Live as if God")
strategies_p2 <- c("God", "No God")

pascal <- ggplot() + 
  scale_x_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p2, position = "top")+
  scale_y_continuous(limits = c(0, 10), expand = c(0,0), breaks = c(2.5, 7.5), labels = strategies_p1) +
  labs(x = "Nature", y = "Pascal") +
  # Player 1 payoffs
  annotate("text", label = expression("Heaven"), x = 2.5, y = 7.5, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression("A"), x = 7.5, y = 7.5, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression("Hell"), x = 2.5, y = 2.5, family = "Fira Sans", size = 6, color = red_pink) +
  annotate("text", label = expression("B"), x = 7.7, y = 2.5, family = "Fira Sans", size = 6, color = red_pink) +
  # theme stuff
  theme_game +
  geom_segment(aes(x = 5, y = 0, xend = 5, yend = 10), linetype  = "solid", size = 1) +
  geom_segment(aes(x = 0, y = 5, xend = 10, yend = 5), linetype  = "solid", size = 1) +
  geom_ribbon(aes(x = w, ymin = 0, ymax = 10), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = z, ymin = 0, ymax = 10), fill = red_pink, alpha = 0.2, linetype = "blank")
pascal
```
]

--

.pull-right[
## Analysis

Pr[G] .mono[=] Probability that God exists.

Expected payoff from belief
<br> $\quad$ .mono[=] Pr[G] &times; Heaven + (1 .mono[-] Pr[G]) &times; A

Expected payoff from non-belief
<br> $\quad$ .mono[=] Pr[G] &times; Hell + (1 .mono[-] Pr[G]) &times; B

]

---
count: false
# Pascal's Wager

.pull-left[
```{r, echo = FALSE, fig.height = 5.5, fig.width = 4.5, dev = "svg"}
pascal
```
]

.pull-right[
## Analysis

Pr[G] .mono[=] Probability that God exists.

Expected payoff from belief
<br> $\quad$ .mono[=] Pr[G] &times; Heaven + (1 .mono[-] Pr[G]) &times; A

Expected payoff from non-belief
<br> $\quad$ .mono[=] Pr[G] &times; Hell + (1 .mono[-] Pr[G]) &times; B

.pink[**Pascal's Conclusion:**] If Pr[G] .mono[>] 0, Heaven .mono[>] Hell, and A is at least as good as B, .pink[belief dominates non-belief].

]

---
class: inverse, middle

# Brinkmanship

---
# Brinkmanship

A **threat** provides a way to make a strategic move. 

- But, you often only have one large threat available.

--
 
Some threats are so big that your opponent doesn't believe you'll carry it out. 

- *e.g.,* "give me $50 or I will kill myself" is unlikely to provide a credible threat.

---
# Brinkmanship

A **credible threat** is costly enough, but not too costly, for opponents to take you seriously.

- *e.g.,* "give me $50 or I will run across the road with my eyes closed" is more likely to induce other player to fork over money.

    - Probability of death is positive, but less than one.

---
# Brinkmanship

Brinkmanship requires that  process is at least partially out of control.

- *e.g.,* if the threatener hears traffic, then the threat wouldn't work.

    - All parties would know the threatener wouldn't run if there were a vehicle coming.

- The probability of the threat actually being carried out must be outside the threatener's control.

---
# Brinkmanship

## Cuban Missile Crisis

A striking example of successful brinkmanship?

- JFK took the world to the brink of nuclear war, and by doing so persuaded Khrushchev to remove missiles from Cuba.

--

But it seems as if game theory cannot explain this.

- Why didn't Khrushchev apply backward induction, figure out the final outcome, and decide not to start the whole process?

---
# Brinkmanship

## George Bush *vs.* Saddam Hussein

Bush played a similar strategy against Hussein: .pink[he gambled].

- **Preferred outcome:** Compliance or removal of Hussein as a response to Bush's threat.

--

- So...Bush lost.

--

Bush and JFK played essentially the same brinkmanship strategy.

- Why didnt Hussein employ backward induction and quit before the war?

---
# Brinkmanship

We can't say Bush played the game better or worse than JFK.

- Brinkmanship necessarily involves taking risks. There is a chance of success and failure.

- Something went wrong in the Persian gulf that didn't in the Cuban missile crisis, but what was it?

---
# Brinkmanship

## Simple Threat Model

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
iraq_1 <- ggplot() + 
  theme_void() +
  # Node 1
  geom_segment(aes(x = 0, y = 0, xend = 1, yend = 1), linetype  = "solid", size = 1, color = met_slate) +
  annotate("label", label = "Don't Threat", x = 0.5, y = 0.5, family = "Fira Sans", size = 5, color = red_pink) +
  geom_segment(aes(x = 0, y = 0, xend = 1, yend = -1), linetype  = "solid", size = 1, color = met_slate) +
  annotate("label", label = "Threat", x = 0.5, y = -0.5, family = "Fira Sans", size = 5, color = red_pink) +
  geom_point(aes(x = 0, y = 0), size = 3, color = red_pink) +
  annotate("text", label = "GWB", x = -0.15, y = 0, family = "Fira Sans", size = 5, color = red_pink) +
  # Subgame 1
  geom_segment(aes(x = 1, y = 1, xend = 2, yend = 1), linetype  = "solid", size = 1, color = met_slate) +
  geom_point(aes(x = 1, y = 1), size = 3, color = purple) +
  annotate("text", label = "SH", x = 1, y = 1.15, family = "Fira Sans", size = 5, color = purple) +
  annotate("text", label = "-2", x = 2.1, y = 1.07, family = "Fira Sans", size = 5, color = red_pink) +
  annotate("text", label = "2", x = 2.12, y = 0.93, family = "Fira Sans", size = 5, color = purple) +
  # Subgame 2
  geom_segment(aes(x = 1, y = -1, xend = 2, yend = 0), linetype  = "solid", size = 1, color = met_slate) +
  annotate("label", label = "Quit", x = 1.5, y = -0.5, family = "Fira Sans", size = 5, color = purple) +
  geom_segment(aes(x = 1, y = -1, xend = 2, yend = -2), linetype  = "solid", size = 1, color = met_slate) +
  annotate("label", label = "Stay", x = 1.5, y = -1.5, family = "Fira Sans", size = 5, color = purple) +
  geom_point(aes(x = 1, y = -1), size = 3, color = purple) +
  annotate("text", label = "SH", x = 1, y = -1.15, family = "Fira Sans", size = 5, color = purple) +
  annotate("text", label = "1", x = 2.11, y = 0.07, family = "Fira Sans", size = 5, color = red_pink) +
  annotate("text", label = "-4", x = 2.1, y = -0.07, family = "Fira Sans", size = 5, color = purple) +
  annotate("text", label = "-10", x = 2.1, y = -1.93, family = "Fira Sans", size = 5, color = red_pink) +
  annotate("text", label = "-8", x = 2.1, y = -2.07, family = "Fira Sans", size = 5, color = purple)
iraq_1
```
]

--

.less-right[
**Problem:** Backward induction tells that .pink[GWB] always .pink[threats] and .purple[SH] always .purple[quits].

- But SH didn't quit!

]

---
count: false
# Brinkmanship

## Simple Threat Model

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
iraq_1
```
]

.less-right[
**Problem:** Backward induction tells that .pink[GWB] always .pink[threats] and .purple[SH] always .purple[quits].

- But SH didn't quit!

- Maybe we have the payoffs wrong?

- Maybe SH values defying the US and simply prefers not to quit?

]

---
# Brinkmanship

## Take 2

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
ggplot() + 
  theme_void() +
  # Node 1
  geom_segment(aes(x = 0, y = 0, xend = 1, yend = 1), linetype  = "solid", size = 1, color = met_slate) +
  annotate("label", label = "Don't Threat", x = 0.5, y = 0.5, family = "Fira Sans", size = 5, color = red_pink) +
  geom_segment(aes(x = 0, y = 0, xend = 1, yend = -1), linetype  = "solid", size = 1, color = met_slate) +
  annotate("label", label = "Threat", x = 0.5, y = -0.5, family = "Fira Sans", size = 5, color = red_pink) +
  geom_point(aes(x = 0, y = 0), size = 3, color = red_pink) +
  annotate("text", label = "GWB", x = -0.15, y = 0, family = "Fira Sans", size = 5, color = red_pink) +
  # Subgame 1
  geom_segment(aes(x = 1, y = 1, xend = 2, yend = 1), linetype  = "solid", size = 1, color = met_slate) +
  geom_point(aes(x = 1, y = 1), size = 3, color = purple) +
  annotate("text", label = "SH", x = 1, y = 1.15, family = "Fira Sans", size = 5, color = purple) +
  annotate("text", label = "-2", x = 2.1, y = 1.07, family = "Fira Sans", size = 5, color = red_pink) +
  annotate("text", label = "2", x = 2.12, y = 0.93, family = "Fira Sans", size = 5, color = purple) +
  # Subgame 2
  geom_segment(aes(x = 1, y = -1, xend = 2, yend = 0), linetype  = "solid", size = 1, color = met_slate) +
  annotate("label", label = "Quit", x = 1.5, y = -0.5, family = "Fira Sans", size = 5, color = purple) +
  geom_segment(aes(x = 1, y = -1, xend = 2, yend = -2), linetype  = "solid", size = 1, color = met_slate) +
  annotate("label", label = "Stay", x = 1.5, y = -1.5, family = "Fira Sans", size = 5, color = purple) +
  geom_point(aes(x = 1, y = -1), size = 3, color = purple) +
  annotate("text", label = "SH", x = 1, y = -1.15, family = "Fira Sans", size = 5, color = purple) +
  annotate("text", label = "1", x = 2.11, y = 0.07, family = "Fira Sans", size = 5, color = red_pink) +
  annotate("text", label = "-8", x = 2.1, y = -0.07, family = "Fira Sans", size = 5, color = purple) +
  annotate("text", label = "-10", x = 2.1, y = -1.93, family = "Fira Sans", size = 5, color = red_pink) +
  annotate("text", label = "-4", x = 2.1, y = -2.07, family = "Fira Sans", size = 5, color = purple)
```
]

--

.less-right[
**Problem:** Backward induction tells that .purple[SH] always .purple[stays] and .pink[GWB] .pink[doesn't threat].

- But GWB didn't quit!

]

---
# Brinkmanship

At least two elements are missing in our analysis:

1. GWB didn't know SH's payoffs: "Am I in game 1 or game 2?"

    - Expensive intel to know how SH valued defying the US.

2. Neither SH nor GWB was sure GWB would invade.

    - Many unkowns (*e.g.,* opposition to GWB's plans in the UN by France, Germany, and Russia).

--

Suppose that SH is tough with probability p $\in$ (0, 1) and GB is committed to his threat with probability q $\in$ (0, 1) 

- New game, similar to Cuban missile crisis.

---
class: clear-slide

.more-left[
```{r, echo = FALSE, fig.height = 6, fig.width = 5, dev = "svg"}
iraq_2 <- ggplot() + 
  scale_x_continuous(limits = c(-1, 2.5)) +
  theme_void() +
  geom_segment(aes(x = 0, y = -1, xend = 0, yend = 2), linetype  = "dashed", size = 1, color = met_slate) +
  # Nature
  geom_segment(aes(x = -0.5, y = 0.5, xend = 0, yend = -1), linetype  = "solid", size = 1, color = met_slate) +
  annotate("label", label = "Pr[Soft SH] = 1 - p", x = -0.5, y = -0.1, family = "Fira Sans", size = 5, color = met_slate) +
  geom_segment(aes(x = -0.5, y = 0.5, xend = 0, yend = 2), linetype  = "solid", size = 1, color = met_slate) +
  annotate("label", label = "Pr[Tough SH] = p", x = -0.5, y = 1.1, family = "Fira Sans", size = 5, color = met_slate) +
  geom_point(aes(x = -0.5, y = 0.5), size = 3, color = met_slate) +
  annotate("text", label = "Nature", x = -0.85, y = 0.5, family = "Fira Sans", size = 5, color = met_slate) +
  # Game 1
  geom_segment(aes(x = 0, y = -1, xend = 1, yend = 0), linetype  = "solid", size = 1, color = met_slate) +
  annotate("label", label = "Don't Threat", x = 0.5, y = -0.5, family = "Fira Sans", size = 5, color = red_pink) +
  geom_segment(aes(x = 0, y = -1, xend = 1, yend = -2), linetype  = "solid", size = 1, color = met_slate) +
  annotate("label", label = "Threat", x = 0.5, y = -1.5, family = "Fira Sans", size = 5, color = red_pink) +
  geom_point(aes(x = 0, y = -1), size = 3, color = red_pink) +
  annotate("text", label = "GWB", x = -0.25, y = -1, family = "Fira Sans", size = 5, color = red_pink) +
  # Subgame 1
  geom_segment(aes(x = 1, y = 0, xend = 2, yend = 0), linetype  = "solid", size = 1, color = met_slate) +
  geom_point(aes(x = 1, y = 0), size = 3, color = purple) +
  annotate("text", label = "SH", x = 1, y = 0.25, family = "Fira Sans", size = 5, color = purple) +
  annotate("text", label = "-2", x = 2.1, y = 0.15, family = "Fira Sans", size = 5, color = red_pink) +
  annotate("text", label = "2", x = 2.13, y = -0.15, family = "Fira Sans", size = 5, color = purple) +
  # Subgame 2
  geom_segment(aes(x = 1, y = -2, xend = 2, yend = -1), linetype  = "solid", size = 1, color = met_slate) +
  annotate("label", label = "Quit", x = 1.5, y = -1.5, family = "Fira Sans", size = 5, color = purple) +
  geom_segment(aes(x = 1, y = -2, xend = 2, yend = -3), linetype  = "solid", size = 1, color = met_slate) +
  annotate("label", label = "Stay", x = 1.5, y = -2.5, family = "Fira Sans", size = 5, color = purple) +
  geom_point(aes(x = 1, y = -2), size = 3, color = purple) +
  annotate("text", label = "SH", x = 1, y = -2.25, family = "Fira Sans", size = 5, color = purple) +
  annotate("text", label = "1", x = 2.13, y = -0.85, family = "Fira Sans", size = 5, color = red_pink) +
  annotate("text", label = "-4", x = 2.1, y = -1.15, family = "Fira Sans", size = 5, color = purple) +
  annotate("text", label = "-2 - 8q", x = 2.3, y = -2.85, family = "Fira Sans", size = 5, color = red_pink) +
  annotate("text", label = "2 - 10q", x = 2.35, y = -3.15, family = "Fira Sans", size = 5, color = purple) +
  # Game 2
  geom_segment(aes(x = 0, y = 2, xend = 1, yend = 3), linetype  = "solid", size = 1, color = met_slate) +
  annotate("label", label = "Don't Threat", x = 0.5, y = 2.5, family = "Fira Sans", size = 5, color = red_pink) +
  geom_segment(aes(x = 0, y = 2, xend = 1, yend = 1), linetype  = "solid", size = 1, color = met_slate) +
  annotate("label", label = "Threat", x = 0.5, y = 1.5, family = "Fira Sans", size = 5, color = red_pink) +
  geom_point(aes(x = 0, y = 2), size = 3, color = red_pink) +
  annotate("text", label = "GWB", x = -0.25, y = 2, family = "Fira Sans", size = 5, color = red_pink) +
  # Subgame 1
  geom_segment(aes(x = 1, y = 3, xend = 2, yend = 3), linetype  = "solid", size = 1, color = met_slate) +
  geom_point(aes(x = 1, y = 3), size = 3, color = purple) +
  annotate("text", label = "SH", x = 1, y = 3.25, family = "Fira Sans", size = 5, color = purple) +
  annotate("text", label = "-2", x = 2.1, y = 3.15, family = "Fira Sans", size = 5, color = red_pink) +
  annotate("text", label = "2", x = 2.13, y = 2.85, family = "Fira Sans", size = 5, color = purple) +
  # Subgame 2
  geom_segment(aes(x = 1, y = 1, xend = 2, yend = 2), linetype  = "solid", size = 1, color = met_slate) +
  annotate("label", label = "Quit", x = 1.5, y = 1.5, family = "Fira Sans", size = 5, color = purple) +
  geom_segment(aes(x = 1, y = 1, xend = 2, yend = 1), linetype  = "solid", size = 1, color = met_slate) +
  annotate("label", label = "Stay", x = 1.5, y = 1, family = "Fira Sans", size = 5, color = purple) +
  geom_point(aes(x = 1, y = 1), size = 3, color = purple) +
  annotate("text", label = "SH", x = 1, y = 0.75, family = "Fira Sans", size = 5, color = purple) +
  annotate("text", label = "1", x = 2.12, y = 2.15, family = "Fira Sans", size = 5, color = red_pink) +
  annotate("text", label = "-8", x = 2.1, y = 1.85, family = "Fira Sans", size = 5, color = purple) +
  annotate("text", label = "2  8q", x = 2.3, y = 1.15, family = "Fira Sans", size = 5, color = red_pink) +
  annotate("text", label = "2  6q", x = 2.33, y = 0.85, family = "Fira Sans", size = 5, color = purple)
iraq_2
```
]

--

.less-right[

## Backward induction

Tough SH won't quit for any q.

- Since 8 .mono[>] 2 .mono[] 6q.

Soft SH quits if q .mono[>] 0.6

- *i.e.,* 2 .mono[] 10q .mono[<] 4.

]

---
class: clear-slide

.more-left[
```{r, echo = FALSE, fig.height = 6, fig.width = 5, dev = "svg"}
iraq_2
```
]

.less-right[

## GWB's options

Threaten and get <br> p(2 .mono[] 8q) .mono[+] (1 .mono[] p)(1)

Don't threaten and get <br> p(2) .mono[+] (1 .mono[] p)(-2) .mono[=] -2

Will threaten if <br> p .mono[<] 3/[3 .mono[+] 8q]

]

---
# Brinkmanship

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5, dev = "svg"}
x <- seq(0, 1, 0.0001)

data <- tibble(
  x = seq(0, 1, 0.0001)
) %>% 
  mutate(y = case_when(x >= 3/11 ~ -(3*(-1 + x))/(8*x),
                       TRUE ~ 1))

gwb_mixed <- function(x) -(3*(-1 + x))/(8*x)
a <- seq(0, 3/11, 0.0001)
b <- seq(3/11, 0.3846, 0.0001)
c <- seq(0.3846, 1, 0.0001)

ggplot() +
  scale_x_continuous(limits = c(0, 1.05), expand = c(0, 0), breaks = c(0, 0.3846, 1), labels = c(0, 0.38, 1)) +
  scale_y_continuous(limits = c(0, 1.05), expand = c(0, 0), breaks = c(0, 0.6, 1), labels = c(0, 0.6, 1)) +
  theme_market +
  labs(x = "p", y = "q") +
  geom_line(data = data, mapping = aes(x = x, y = y), size = 1, color = met_slate) +
  geom_ribbon(aes(x = c, ymin = 0, ymax = gwb_mixed(c)), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = a, ymin = 0, ymax = 0.6), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = b, ymin = 0, ymax = 0.6), fill = red_pink, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = a, ymin = 0.6, ymax = 1), fill = green, alpha = 0.2, linetype = "blank") +
  geom_ribbon(aes(x = b, ymin = 0.6, ymax = gwb_mixed(b)), fill = green, alpha = 0.2, linetype = "blank") +
  geom_segment(aes(x = 0, xend = 3/11, y = 1), yend = 1, color = met_slate, size = 1) +
  geom_point(aes(x = 0.3846, y = 0.6), color = met_slate, size = 2) +
  geom_segment(aes(x = 0, xend = 0.3846, y = 0.6, yend = 0.6), color = met_slate, linetype = "dashed", size = 1) +
  geom_segment(aes(x = 0.3846, xend = 0.3846, y = 0, yend = 0.6), color = met_slate, linetype = "dashed", size = 1) +
  annotate("text", label = wrapper("Zone where GWB would not threaten.", 20), x = 0.75, y = 0.75, family = "Fira Sans", size = 3, color = met_slate) +
  annotate("text", label = wrapper("Zone where GWB should not threaten. Either type of SH would not quit.", 25), x = 0.2, y = 0.35, family = "Fira Sans", size = 3, color = red_pink) +
  annotate("text", label = wrapper("Zone where GWB threatens. Soft SH quits. Tough SH stays.", 25), x = 0.175, y = 0.75, family = "Fira Sans", size = 3, color = green)

```
]

--

.less-right[

GWB tries a q .mono[<] 0.6. If this doesn't work and doesn't trigger conflict, he tries a slightly higher q.

Either GWB reaches q = 0.6 and a soft SH quits or we're going to war.

No different from JFK and Khrushchev, except for the outcome.
]

---
class: inverse, middle

# Midterm 2

---
# Midterm Topics

.green[Anything from lectures or discussions (weeks 5 through 7 only)] .hi-green[is fair game!]

1. Policy Levers: Taxes & Subsidies

2. Policy Levers: Price Controls

3. How Economists Learn from Data I

4. How Economists Learn from Data II

5. Market Failure: Externalities

6. Game Theory

---
# Midterm Structure

## Multiple Choice

- 50 questions

- 1 point per question

- Multiple groups of sequential questions (*e.g.,* about graphs or tables)

---
# Midterm Protocol

## Materials

- Writing utensil
- 3-inch-by-5-inch note card
- Basic or scientific calculator (no graphing or programming capabilities)
- UO ID card
- .hi[Nothing else]

--

## Procedure

- **Randomized** seating chart
- 80 minutes from *"you may begin"* to *"pencils down"*
- First 30 minutes: .hi[quiet period] (no questions, no getting up)
- Last 50 minutes: ask lots of questions
- Show your UO ID card as you turn-in your exam

---
# Midterm Preparation

1. Lecture slides

2. Discussion worksheets

3. Practice midterm questions

    - No solutions posted, but you can ask about the questions in office hours.

4. Extended office hours

    - M and T 14:00-16:00 in **412 PLC**

5. Prepare your note card
