---
title: "The Market Mechanism"
subtitle: "EC 201: Principles of Microeconomics"
author: "Kyle Raze"
date: "Winter 2020"
output:
  xaringan::moon_reader:
    css: ['default', 'metropolis', 'metropolis-fonts', 'my-css.css']
    self_contained: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "https://platform.twitter.com/widgets.js"
---
class: inverse, middle

```{r Setup, include = F}
options(htmltools.dir.version = FALSE)
library(pacman)
p_load(ggthemes, readxl, viridis, knitr, extrafont, tidyverse, RcmdrPlugin.KMggplot2, magrittr, janitor, kableExtra, scales)
# Define colors
red_pink <- "#e64173"
met_slate <- "#272822" # metropolis font color 
purple <- "#9370DB"
green <- "#007935"
light_green <- "#7DBA97"
orange <- "#FD5F00"
turquoise <- "#44C1C4"
red <- "#b92e34"
# Notes directory
dir_slides <- "~/GitHub/Intro_Microeconomics_W20/Lectures/05-Market_Mechanism/"
# Knitr options
opts_chunk$set(
  comment = "#>",
  fig.align = "center",
  fig.height = 7,
  fig.width = 10.5,
  #dpi = 300,
  #cache = T,
  warning = F,
  message = F
)  
theme_simple <- theme_bw() + theme(
  axis.line = element_line(color = met_slate),
  panel.grid = element_blank(),
  rect = element_blank(),
  strip.text = element_blank(),
  text = element_text(family = "Fira Sans", color = met_slate, size = 17),
  axis.text.x = element_text(size = 17),
  axis.text.y = element_text(size = 17),
  axis.ticks = element_blank()
)
theme_market <- theme_bw() + theme(
  axis.line = element_line(color = met_slate),
  panel.grid = element_blank(),
  rect = element_blank(),
  strip.text = element_blank(),
  text = element_text(family = "Fira Sans", color = met_slate, size = 17),
  axis.title.x = element_text(hjust = 1, size = 17),
  axis.title.y = element_text(hjust = 1, angle = 0, size = 17),
  # axis.text.x = element_text(size = 12),
  # axis.text.y = element_text(size = 12),
  axis.ticks = element_blank()
)
theme_gif <- theme_bw() + theme(
  axis.line = element_line(color = met_slate),
  panel.grid = element_blank(),
  rect = element_blank(),
  text = element_text(family = "Fira Sans", color = met_slate, size = 17),
  axis.text.x = element_text(size = 12),
  axis.text.y = element_text(size = 12),
  axis.ticks = element_blank()
)
wrapper <- function(x, ...) paste(strwrap(x, ...), collapse = "\n")
# functions
step_demand <- tibble(x = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12), mv = c(10, 10, 9, 9, 8, 8, 7, 7, 6, 6, 5, 4, 4))
step_supply <- tibble(x = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12), mv = c(2, 2, 3, 3, 4, 4, 5, 5, 6, 6, 7, 8, 8))
step_supply_tax <- step_supply %>% 
  mutate(mv = mv + 2)

data <- read_csv("05-Activity_Pit_Market/Results.csv") %>% 
  filter(!is.na(price))
```

# Prologue

---
# Agenda

## Pit Market Activity

1. Read instructions handout

2. Join a team

3. Make trades and earn extra participation points

## The Market Mechanism

- Activity debrief and discussion

## Midterm 1 Overview

- One week from today!

---
class: inverse, middle

# Pit Market Activity

---
class: clear-slide

.center[**Pit Market Results**]

.pull-left[
```{r price, echo = FALSE, fig.height = 5.25, fig.width = 4, dev = "svg"}
data %>% 
  group_by(round) %>% 
  summarize(median = median(price),
            lower = quantile(price, 1/4),
            upper = quantile(price, 3/4)) %>% 
  ungroup() %>% 
  ggplot() +
  geom_errorbar(aes(x = round, ymin = lower, ymax = upper), size = 1, color = red_pink) + 
  geom_point(aes(x = round, y = median), size = 2, color = red_pink) +
  labs(x = "Round", y = "Price") +
  scale_x_continuous(limits = c(0, 7.5), expand = c(0, 0), breaks = seq(0, 7, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  geom_vline(xintercept = 5.5, size = 1, linetype = "dashed") +
  annotate("label", label = "$2 Tax Imposed", x = 5.5, y = 10, color = met_slate, family = "Fira Sans", size = 5) +
  theme_simple
```
]

.pull-right[
```{r quantity, echo = FALSE, fig.height = 5.25, fig.width = 4, dev = "svg"}
data %>% 
  group_by(round) %>% 
  summarize(trades = n()) %>% 
  ungroup() %>% 
  ggplot(aes(x = round, y = trades)) +
  geom_col(fill = red_pink, color = red_pink, alpha = 0.5) +
  labs(x = "Round", y = "Number of Trades") +
  scale_x_continuous(limits = c(0, 7.5), expand = c(0, 0), breaks = seq(0, 7, 1)) +
  scale_y_continuous(limits = c(0, 12.5), expand = c(0, 0), breaks = seq(0, 12, 1)) +
  geom_vline(xintercept = 5.5, size = 1, linetype = "dashed") +
  annotate("label", label = "$2 Tax Imposed", x = 5.5, y = 12, color = met_slate, family = "Fira Sans", size = 5) +
  theme_simple
```
]

---
class: inverse, middle

# The Market Mechanism

---
# Pit Market

.less-right[
## .pink[Buyers]

.pink[Maximum willingness to pay:]
.pull-left[
10 of .red[`r emo::ji("diamond_suit")`] <br>
9 of .red[`r emo::ji("diamond_suit")`] <br>
8 of .red[`r emo::ji("diamond_suit")`] <br>
7 of .red[`r emo::ji("diamond_suit")`] <br>
6 of .red[`r emo::ji("diamond_suit")`] <br>
5 of .red[`r emo::ji("diamond_suit")`] <br>
4 of .red[`r emo::ji("diamond_suit")`]
]
.pull-right[
10 of .red[`r emo::ji("heart_suit")`] <br>
9 of .red[`r emo::ji("heart_suit")`] <br>
8 of .red[`r emo::ji("heart_suit")`] <br>
7 of .red[`r emo::ji("heart_suit")`] <br>
6 of .red[`r emo::ji("heart_suit")`] 
]
]

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5.4, dev = "svg"}
ggplot() +
  #geom_step(data = step_supply, mapping = aes(x = x, y = mv), color = met_slate, size = 1) +
  geom_step(data = step_demand, mapping = aes(x = x, y = mv), color = red_pink, size = 1) +
  annotate("text", label = "D", x = 12.5, y = 4, color = red_pink, family = "Fira Sans", size = 9) +
  #geom_step(data = step_supply_tax, mapping = aes(x = x, y = mv), color = met_slate, size = 1, alpha = 0.5) +
  labs(x = "Q", y = "P") +
  scale_x_continuous(limits = c(0, 14), expand = c(0, 0), breaks = seq(0, 12, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market
```
]

---
# Pit Market

.less-right[
## .purple[Sellers]

.purple[Minimum willingness to accept:]
.pull-left[
2 of `r emo::ji("spade_suit")` <br>
3 of `r emo::ji("spade_suit")` <br>
4 of `r emo::ji("spade_suit")` <br>
5 of `r emo::ji("spade_suit")` <br>
6 of `r emo::ji("spade_suit")` <br>
7 of `r emo::ji("spade_suit")` <br>
8 of `r emo::ji("spade_suit")`
]
.pull-right[
2 of `r emo::ji("club_suit")` <br>
3 of `r emo::ji("club_suit")` <br>
4 of `r emo::ji("club_suit")` <br>
5 of `r emo::ji("club_suit")` <br>
6 of `r emo::ji("club_suit")`
]
]

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5.4, dev = "svg"}
ggplot() +
  geom_step(data = step_supply, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  annotate("text", label = "S", x = 12.5, y = 8, color = purple, family = "Fira Sans", size = 9) +
  #geom_step(data = step_demand, mapping = aes(x = x, y = mv), color = red_pink, size = 1) +
  #annotate("text", label = "D", x = 12.5, y = 4, color = red_pink, family = "Fira Sans", size = 9) +
  #geom_step(data = step_supply_tax, mapping = aes(x = x, y = mv), color = purple, size = 1, alpha = 0.5) +
  labs(x = "Q", y = "P") +
  scale_x_continuous(limits = c(0, 14), expand = c(0, 0), breaks = seq(0, 12, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market
```
]

---
# Pit Market

.less-right[
## .purple[Sellers]

The tax increases each seller's cost by .mono[$]2.

]

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5.4, dev = "svg"}
ggplot() +
  geom_step(data = step_supply, mapping = aes(x = x, y = mv), color = purple, size = 1, alpha = 0.5) +
  annotate("text", label = "S", x = 12.5, y = 8, color = purple, family = "Fira Sans", size = 9, alpha = 0.5) +
  #geom_step(data = step_demand, mapping = aes(x = x, y = mv), color = red_pink, size = 1) +
  #annotate("text", label = "D", x = 12.5, y = 4, color = red_pink, family = "Fira Sans", size = 9) +
  geom_step(data = step_supply_tax, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  annotate("text", label = expression(S[Tax]), x = 13, y = 10, color = purple, family = "Fira Sans", size = 9) +
  labs(x = "Q", y = "P") +
  scale_x_continuous(limits = c(0, 14), expand = c(0, 0), breaks = seq(0, 12, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market
```
]

---
# Pit Market

.less-right[
## Equilibrium

Theory predicts that a price of .mono[$]6 clears the market..super[.pink[<span>&#8224;</span>]]

**Q:** Who trades? Who doesn't?

]

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5.4, dev = "svg"}
ggplot() +
  geom_step(data = step_supply, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  annotate("text", label = "S", x = 12.5, y = 8, color = purple, family = "Fira Sans", size = 9) +
  geom_step(data = step_demand, mapping = aes(x = x, y = mv), color = red_pink, size = 1) +
  annotate("text", label = "D", x = 12.5, y = 4, color = red_pink, family = "Fira Sans", size = 9) +
  # geom_step(data = step_supply_tax, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  # annotate("text", label = expression(S[Tax]), x = 13, y = 10, color = purple, family = "Fira Sans", size = 9) +
  labs(x = "Q", y = "P") +
  scale_x_continuous(limits = c(0, 14), expand = c(0, 0), breaks = seq(0, 12, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market
```
]

.footnote[.pink[<span>&#8224;</span>]: *Clears the market* .mono[<=>] Q.sub[D] .mono[=] Q.sub[S].]

---
count: false
# Pit Market

.less-right[
## Equilibrium

Theory predicts that a price of .mono[$]6 clears the market..super[.pink[<span>&#8224;</span>]]

**Q:** Who trades? Who doesn't?

]

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5.4, dev = "svg"}
ggplot() +
  geom_step(data = step_supply, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  annotate("text", label = "S", x = 12.5, y = 8, color = purple, family = "Fira Sans", size = 9) +
  geom_step(data = step_demand, mapping = aes(x = x, y = mv), color = red_pink, size = 1) +
  annotate("text", label = "D", x = 12.5, y = 4, color = red_pink, family = "Fira Sans", size = 9) +
  # geom_step(data = step_supply_tax, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  # annotate("text", label = expression(S[Tax]), x = 13, y = 10, color = purple, family = "Fira Sans", size = 9) +
  labs(x = "Q", y = "P") +
  scale_x_continuous(limits = c(0, 14), expand = c(0, 0), breaks = seq(0, 12, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  geom_point(aes(x = 10, y = 6), color = met_slate, size = 2) +
  geom_segment(aes(x = 10, y = 0, xend = 10, yend = 6), linetype  = "dashed", color = met_slate, size = 1) + 
  geom_segment(aes(x = 0, y = 6, xend = 10, yend = 6), linetype  = "dashed", color = met_slate, size = 1)
```
]

.footnote[.pink[<span>&#8224;</span>]: *Clears the market* .mono[<=>] Q.sub[D] .mono[=] Q.sub[S].]

---
count: false
# Pit Market

.less-right[
## Equilibrium

Theory predicts that a price of .mono[$]6 clears the market..super[.pink[<span>&#8224;</span>]]

**Q:** Who trades? Who doesn't?

]

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5.4, dev = "svg"}
ggplot() +
  geom_step(data = step_supply, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  annotate("text", label = "S", x = 12.5, y = 8, color = purple, family = "Fira Sans", size = 9) +
  geom_step(data = step_demand, mapping = aes(x = x, y = mv), color = red_pink, size = 1) +
  annotate("text", label = "D", x = 12.5, y = 4, color = red_pink, family = "Fira Sans", size = 9) +
  # geom_step(data = step_supply_tax, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  # annotate("text", label = expression(S[Tax]), x = 13, y = 10, color = purple, family = "Fira Sans", size = 9) +
  labs(x = "Q", y = "P") +
  scale_x_continuous(limits = c(0, 14), expand = c(0, 0), breaks = seq(0, 12, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  geom_point(aes(x = 8, y = 6), color = met_slate, size = 2) +
  geom_segment(aes(x = 8, y = 0, xend = 8, yend = 6), linetype  = "dashed", color = met_slate, size = 1) + 
  geom_segment(aes(x = 0, y = 6, xend = 8, yend = 6), linetype  = "dashed", color = met_slate, size = 1)
```
]

.footnote[.pink[<span>&#8224;</span>]: *Clears the market* .mono[<=>] Q.sub[D] .mono[=] Q.sub[S].]

---
count: false
# Pit Market

.less-right[
## Equilibrium

Theory predicts that a price of .mono[$]6 clears the market..super[.pink[<span>&#8224;</span>]]

**Q:** Who trades? Who doesn't?

]

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5.4, dev = "svg"}
ggplot() +
  geom_step(data = step_supply, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  annotate("text", label = "S", x = 12.5, y = 8, color = purple, family = "Fira Sans", size = 9) +
  geom_step(data = step_demand, mapping = aes(x = x, y = mv), color = red_pink, size = 1) +
  annotate("text", label = "D", x = 12.5, y = 4, color = red_pink, family = "Fira Sans", size = 9) +
  # geom_step(data = step_supply_tax, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  # annotate("text", label = expression(S[Tax]), x = 13, y = 10, color = purple, family = "Fira Sans", size = 9) +
  labs(x = "Q", y = "P") +
  scale_x_continuous(limits = c(0, 14), expand = c(0, 0), breaks = seq(0, 12, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  geom_point(aes(x = 9, y = 6), color = met_slate, size = 2) +
  geom_segment(aes(x = 9, y = 0, xend = 9, yend = 6), linetype  = "dashed", color = met_slate, size = 1) + 
  geom_segment(aes(x = 0, y = 6, xend = 9, yend = 6), linetype  = "dashed", color = met_slate, size = 1)
```
]

.footnote[.pink[<span>&#8224;</span>]: *Clears the market* .mono[<=>] Q.sub[D] .mono[=] Q.sub[S].]

---
# Pit Market

.less-right[
## Market Surplus

Theory predicts that prices above .mono[$]6 create a market surplus..super[.pink[<span>&#8224;</span>]]

]

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5.4, dev = "svg"}
ggplot() +
  geom_step(data = step_supply, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  annotate("text", label = "S", x = 12.5, y = 8, color = purple, family = "Fira Sans", size = 9) +
  geom_step(data = step_demand, mapping = aes(x = x, y = mv), color = red_pink, size = 1) +
  annotate("text", label = "D", x = 12.5, y = 4, color = red_pink, family = "Fira Sans", size = 9) +
  # geom_step(data = step_supply_tax, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  # annotate("text", label = expression(S[Tax]), x = 13, y = 10, color = purple, family = "Fira Sans", size = 9) +
  labs(x = "Q", y = "P") +
  scale_x_continuous(limits = c(0, 14), expand = c(0, 0), breaks = seq(0, 12, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  geom_hline(yintercept = 7.5, linetype  = "dashed", color = met_slate, size = 1) #+
  # geom_point(aes(x = 11, y = 7.5), color = met_slate, size = 2) +
  # geom_point(aes(x = 6, y = 7.5), color = met_slate, size = 2) +
  # geom_segment(aes(x = 6, y = 0, xend = 6, yend = 7.5), linetype  = "dashed", color = met_slate, size = 1) + 
  # geom_segment(aes(x = 11, y = 0, xend = 11, yend = 7.5), linetype  = "dashed", color = met_slate, size = 1)
```
]

.footnote[.pink[<span>&#8224;</span>]: *Market surplus* .mono[<=>] Q.sub[D] .mono[<] Q.sub[S].]

---
count: false
# Pit Market

.less-right[
## Market Surplus

Theory predicts that prices above .mono[$]6 create a market surplus..super[.pink[<span>&#8224;</span>]]

]

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5.4, dev = "svg"}
ggplot() +
  geom_step(data = step_supply, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  annotate("text", label = "S", x = 12.5, y = 8, color = purple, family = "Fira Sans", size = 9) +
  geom_step(data = step_demand, mapping = aes(x = x, y = mv), color = red_pink, size = 1) +
  annotate("text", label = "D", x = 12.5, y = 4, color = red_pink, family = "Fira Sans", size = 9) +
  # geom_step(data = step_supply_tax, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  # annotate("text", label = expression(S[Tax]), x = 13, y = 10, color = purple, family = "Fira Sans", size = 9) +
  labs(x = "Q", y = "P") +
  scale_x_continuous(limits = c(0, 14), expand = c(0, 0), breaks = seq(0, 12, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  geom_hline(yintercept = 7.5, linetype  = "dashed", color = met_slate, size = 1) +
  geom_point(aes(x = 11, y = 7.5), color = met_slate, size = 2) +
  geom_point(aes(x = 6, y = 7.5), color = met_slate, size = 2) +
  geom_segment(aes(x = 6, y = 0, xend = 6, yend = 7.5), linetype  = "dashed", color = met_slate, size = 1) +
  geom_segment(aes(x = 11, y = 0, xend = 11, yend = 7.5), linetype  = "dashed", color = met_slate, size = 1)
```
]

.footnote[.pink[<span>&#8224;</span>]: *Market surplus* .mono[<=>] Q.sub[D] .mono[<] Q.sub[S].]

---
# Pit Market

.less-right[
## Shortage

Theory predicts that prices below .mono[$]6 create a shortage..super[.pink[<span>&#8224;</span>]]

]

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5.4, dev = "svg"}
ggplot() +
  geom_step(data = step_supply, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  annotate("text", label = "S", x = 12.5, y = 8, color = purple, family = "Fira Sans", size = 9) +
  geom_step(data = step_demand, mapping = aes(x = x, y = mv), color = red_pink, size = 1) +
  annotate("text", label = "D", x = 12.5, y = 4, color = red_pink, family = "Fira Sans", size = 9) +
  # geom_step(data = step_supply_tax, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  # annotate("text", label = expression(S[Tax]), x = 13, y = 10, color = purple, family = "Fira Sans", size = 9) +
  labs(x = "Q", y = "P") +
  scale_x_continuous(limits = c(0, 14), expand = c(0, 0), breaks = seq(0, 12, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  geom_hline(yintercept = 4.5, linetype  = "dashed", color = met_slate, size = 1)
```
]

.footnote[.pink[<span>&#8224;</span>]: *Shortage* .mono[<=>] Q.sub[D] .mono[>] Q.sub[S].]

---
count: false
# Pit Market

.less-right[
## Shortage

Theory predicts that prices below .mono[$]6 create a shortage..super[.pink[<span>&#8224;</span>]]

]

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5.4, dev = "svg"}
ggplot() +
  geom_step(data = step_supply, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  annotate("text", label = "S", x = 12.5, y = 8, color = purple, family = "Fira Sans", size = 9) +
  geom_step(data = step_demand, mapping = aes(x = x, y = mv), color = red_pink, size = 1) +
  annotate("text", label = "D", x = 12.5, y = 4, color = red_pink, family = "Fira Sans", size = 9) +
  # geom_step(data = step_supply_tax, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  # annotate("text", label = expression(S[Tax]), x = 13, y = 10, color = purple, family = "Fira Sans", size = 9) +
  labs(x = "Q", y = "P") +
  scale_x_continuous(limits = c(0, 14), expand = c(0, 0), breaks = seq(0, 12, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  geom_hline(yintercept = 4.5, linetype  = "dashed", color = met_slate, size = 1) +
  geom_point(aes(x = 11, y = 4.5), color = met_slate, size = 2) +
  geom_point(aes(x = 6, y = 4.5), color = met_slate, size = 2) +
  geom_segment(aes(x = 6, y = 0, xend = 6, yend = 4.5), linetype  = "dashed", color = met_slate, size = 1) +
  geom_segment(aes(x = 11, y = 0, xend = 11, yend = 4.5), linetype  = "dashed", color = met_slate, size = 1)
```
]

.footnote[.pink[<span>&#8224;</span>]: *Shortage* .mono[<=>] Q.sub[D] .mono[>] Q.sub[S].]

---
# Pit Market

.less-right[
## Tax

The tax increases each seller's cost by .mono[$]2:

1. Supply curve shifts.
2. Market adjusts to a new equilibrium.

**Q:** By how much does the equilibrium price increase?

]

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5.4, dev = "svg"}
ggplot() +
  geom_step(data = step_supply, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  annotate("text", label = "S", x = 12.5, y = 8, color = purple, family = "Fira Sans", size = 9) +
  geom_step(data = step_demand, mapping = aes(x = x, y = mv), color = red_pink, size = 1) +
  annotate("text", label = "D", x = 12.5, y = 4, color = red_pink, family = "Fira Sans", size = 9) +
  # geom_step(data = step_supply_tax, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  # annotate("text", label = expression(S[Tax]), x = 13, y = 10, color = purple, family = "Fira Sans", size = 9) +
  labs(x = "Q", y = "P") +
  scale_x_continuous(limits = c(0, 14), expand = c(0, 0), breaks = seq(0, 12, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market
```
]

---
count: false
# Pit Market

.less-right[
## Tax

The tax increases each seller's cost by .mono[$]2:

1. Supply curve shifts.
2. Market adjusts to a new equilibrium.

**Q:** By how much does the equilibrium price increase?

]

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5.4, dev = "svg"}
ggplot() +
  geom_step(data = step_supply, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  annotate("text", label = "S", x = 12.5, y = 8, color = purple, family = "Fira Sans", size = 9) +
  geom_step(data = step_demand, mapping = aes(x = x, y = mv), color = red_pink, size = 1) +
  annotate("text", label = "D", x = 12.5, y = 4, color = red_pink, family = "Fira Sans", size = 9) +
  # geom_step(data = step_supply_tax, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  # annotate("text", label = expression(S[Tax]), x = 13, y = 10, color = purple, family = "Fira Sans", size = 9) +
  labs(x = "Q", y = "P") +
  scale_x_continuous(limits = c(0, 14), expand = c(0, 0), breaks = seq(0, 12, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  geom_point(aes(x = 9, y = 6), color = met_slate, size = 2) +
  geom_segment(aes(x = 9, y = 0, xend = 9, yend = 6), linetype  = "dashed", color = met_slate, size = 1) + 
  geom_segment(aes(x = 0, y = 6, xend = 9, yend = 6), linetype  = "dashed", color = met_slate, size = 1)
```
]

---
count: false
# Pit Market

.less-right[
## Tax

The tax increases each seller's cost by .mono[$]2:

1. Supply curve shifts.
2. Market adjusts to a new equilibrium.

**Q:** By how much does the equilibrium price increase?

]

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5.4, dev = "svg"}
ggplot() +
  geom_step(data = step_supply, mapping = aes(x = x, y = mv), color = purple, size = 1, alpha = 0.5) +
  annotate("text", label = "S", x = 12.5, y = 8, color = purple, family = "Fira Sans", size = 9, alpha = 0.5) +
  geom_step(data = step_demand, mapping = aes(x = x, y = mv), color = red_pink, size = 1) +
  annotate("text", label = "D", x = 12.5, y = 4, color = red_pink, family = "Fira Sans", size = 9) +
  geom_step(data = step_supply_tax, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  annotate("text", label = expression(S[Tax]), x = 13, y = 10, color = purple, family = "Fira Sans", size = 9) +
  labs(x = "Q", y = "P") +
  scale_x_continuous(limits = c(0, 14), expand = c(0, 0), breaks = seq(0, 12, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  geom_point(aes(x = 9, y = 6), color = met_slate, size = 2, alpha = 0.5) +
  geom_segment(aes(x = 9, y = 0, xend = 9, yend = 6), linetype  = "dashed", color = met_slate, size = 1, alpha = 0.5) + 
  geom_segment(aes(x = 0, y = 6, xend = 9, yend = 6), linetype  = "dashed", color = met_slate, size = 1, alpha = 0.5)
```
]

---
count: false
# Pit Market

.less-right[
## Tax

The tax increases each seller's cost by .mono[$]2:

1. Supply curve shifts.
2. Market adjusts to a new equilibrium.

**Q:** By how much does the equilibrium price increase?

]

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5.4, dev = "svg"}
ggplot() +
  geom_step(data = step_supply, mapping = aes(x = x, y = mv), color = purple, size = 1, alpha = 0.5) +
  annotate("text", label = "S", x = 12.5, y = 8, color = purple, family = "Fira Sans", size = 9, alpha = 0.5) +
  geom_step(data = step_demand, mapping = aes(x = x, y = mv), color = red_pink, size = 1) +
  annotate("text", label = "D", x = 12.5, y = 4, color = red_pink, family = "Fira Sans", size = 9) +
  geom_step(data = step_supply_tax, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  annotate("text", label = expression(S[Tax]), x = 13, y = 10, color = purple, family = "Fira Sans", size = 9) +
  labs(x = "Q", y = "P") +
  scale_x_continuous(limits = c(0, 14), expand = c(0, 0), breaks = seq(0, 12, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  geom_point(aes(x = 9, y = 6), color = met_slate, size = 2, alpha = 0.5) +
  geom_segment(aes(x = 9, y = 0, xend = 9, yend = 6), linetype  = "dashed", color = met_slate, size = 1, alpha = 0.5) + 
  geom_segment(aes(x = 0, y = 6, xend = 9, yend = 6), linetype  = "dashed", color = met_slate, size = 1, alpha = 0.5) +
  geom_point(aes(x = 7, y = 7), color = met_slate, size = 2, alpha = 1) +
  geom_segment(aes(x = 7, y = 0, xend = 7, yend = 7), linetype  = "dashed", color = met_slate, size = 1, alpha = 1) + 
  geom_segment(aes(x = 0, y = 7, xend = 7, yend = 7), linetype  = "dashed", color = met_slate, size = 1, alpha = 1)
```
]

---
count: false
# Pit Market

.less-right[
## Tax

The tax increases each seller's cost by .mono[$]2:

1. Supply curve shifts.
2. Market adjusts to a new equilibrium.

**Q:** By how much does the equilibrium price increase?

]

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5.4, dev = "svg"}
ggplot() +
  geom_step(data = step_supply, mapping = aes(x = x, y = mv), color = purple, size = 1, alpha = 0.5) +
  annotate("text", label = "S", x = 12.5, y = 8, color = purple, family = "Fira Sans", size = 9, alpha = 0.5) +
  geom_step(data = step_demand, mapping = aes(x = x, y = mv), color = red_pink, size = 1) +
  annotate("text", label = "D", x = 12.5, y = 4, color = red_pink, family = "Fira Sans", size = 9) +
  geom_step(data = step_supply_tax, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  annotate("text", label = expression(S[Tax]), x = 13, y = 10, color = purple, family = "Fira Sans", size = 9) +
  labs(x = "Q", y = "P") +
  scale_x_continuous(limits = c(0, 14), expand = c(0, 0), breaks = seq(0, 12, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  geom_point(aes(x = 7, y = 7), color = met_slate, size = 2, alpha = 1) +
  geom_segment(aes(x = 7, y = 0, xend = 7, yend = 7), linetype  = "dashed", color = met_slate, size = 1, alpha = 1) + 
  geom_segment(aes(x = 0, y = 7, xend = 7, yend = 7), linetype  = "dashed", color = met_slate, size = 1, alpha = 1)
```
]

---
# Mutual Gains from Trade

.less-right[

.pink[Consumer surplus] .mono[=] sum of buyers' earnings.

.purple[Producer surplus] .mono[=] sum of sellers' earnings.

.green[Total surplus] .mono[=] .pink[consumer surplus] .mono[+] .purple[producer surplus].


]

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5.4, dev = "svg"}
ggplot() +
  geom_step(data = step_supply, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  annotate("text", label = "S", x = 12.5, y = 8, color = purple, family = "Fira Sans", size = 9) +
  geom_step(data = step_demand, mapping = aes(x = x, y = mv), color = red_pink, size = 1) +
  annotate("text", label = "D", x = 12.5, y = 4, color = red_pink, family = "Fira Sans", size = 9) +
  # geom_step(data = step_supply_tax, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  # annotate("text", label = expression(S[Tax]), x = 13, y = 10, color = purple, family = "Fira Sans", size = 9) +
  labs(x = "Q", y = "P") +
  scale_x_continuous(limits = c(0, 14), expand = c(0, 0), breaks = seq(0, 12, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  geom_stepribbon(data = step_demand, aes(ymax = ifelse(x < 9, mv, 6), ymin = 6, x = x), fill = red_pink, alpha = 0.2, linetype = "blank") +
  annotate("text", label = "CS", x = 2, y = 7.5, color = red_pink, family = "Fira Sans", size = 9) +
  geom_stepribbon(data = step_supply, aes(ymax = ifelse(x < 9, mv, 6), ymin = 6, x = x), fill = purple, alpha = 0.2, linetype = "blank") +
  annotate("text", label = "PS", x = 2, y = 4.5, color = purple, family = "Fira Sans", size = 9) +
  geom_point(aes(x = 9, y = 6), color = met_slate, size = 2) +
  geom_segment(aes(x = 9, y = 0, xend = 9, yend = 6), linetype  = "dashed", color = met_slate, size = 1) + 
  geom_segment(aes(x = 0, y = 6, xend = 9, yend = 6), linetype  = "dashed", color = met_slate, size = 1)
```
]

---
count: false
# Mutual Gains from Trade

.less-right[

.pink[Consumer surplus] .mono[=] sum of buyers' earnings.

.purple[Producer surplus] .mono[=] sum of sellers' earnings.

.green[Total surplus] .mono[=] .pink[consumer surplus] .mono[+] .purple[producer surplus].

]

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5.4, dev = "svg"}
ggplot() +
  geom_step(data = step_supply, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  annotate("text", label = "S", x = 12.5, y = 8, color = purple, family = "Fira Sans", size = 9) +
  geom_step(data = step_demand, mapping = aes(x = x, y = mv), color = red_pink, size = 1) +
  annotate("text", label = "D", x = 12.5, y = 4, color = red_pink, family = "Fira Sans", size = 9) +
  # geom_step(data = step_supply_tax, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  # annotate("text", label = expression(S[Tax]), x = 13, y = 10, color = purple, family = "Fira Sans", size = 9) +
  labs(x = "Q", y = "P") +
  scale_x_continuous(limits = c(0, 14), expand = c(0, 0), breaks = seq(0, 12, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  geom_stepribbon(data = step_demand, aes(ymax = ifelse(x < 9, mv, 6), ymin = 6, x = x), fill = green, alpha = 0.2, linetype = "blank") +
  annotate("text", label = "TS", x = 1.5, y = 7, color = green, family = "Fira Sans", size = 9) +
  geom_stepribbon(data = step_supply, aes(ymax = ifelse(x < 9, mv, 6), ymin = 6, x = x), fill = green, alpha = 0.2, linetype = "blank") +
  geom_point(aes(x = 9, y = 6), color = met_slate, size = 2) +
  geom_segment(aes(x = 9, y = 0, xend = 9, yend = 6), linetype  = "dashed", color = met_slate, size = 1) + 
  geom_segment(aes(x = 0, y = 6, xend = 9, yend = 6), linetype  = "dashed", color = met_slate, size = 1)
```
]

---
# Mutual Gains from Trade

.less-right[
## Market Surplus

P .mono[=].mono[$]7.50 .mono[==>] TS .mono[=] .mono[$]36.

]

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5.4, dev = "svg"}
ggplot() +
  geom_step(data = step_supply, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  annotate("text", label = "S", x = 12.5, y = 8, color = purple, family = "Fira Sans", size = 9) +
  geom_step(data = step_demand, mapping = aes(x = x, y = mv), color = red_pink, size = 1) +
  annotate("text", label = "D", x = 12.5, y = 4, color = red_pink, family = "Fira Sans", size = 9) +
  # geom_step(data = step_supply_tax, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  # annotate("text", label = expression(S[Tax]), x = 13, y = 10, color = purple, family = "Fira Sans", size = 9) +
  labs(x = "Q", y = "P") +
  scale_x_continuous(limits = c(0, 14), expand = c(0, 0), breaks = seq(0, 12, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  geom_stepribbon(data = step_demand, aes(ymax = ifelse(x < 6, mv, 7.5), ymin = 7.5, x = x), fill = green, alpha = 0.2, linetype = "blank") +
  annotate("text", label = "TS", x = 2, y = 6, color = green, family = "Fira Sans", size = 9) +
  geom_stepribbon(data = step_supply, aes(ymax = ifelse(x < 6, mv, 7.5), ymin = 7.5, x = x), fill = green, alpha = 0.2, linetype = "blank") +
  geom_hline(yintercept = 7.5, linetype  = "dashed", color = met_slate, size = 1) +
  geom_point(aes(x = 11, y = 7.5), color = met_slate, size = 2) +
  geom_point(aes(x = 6, y = 7.5), color = met_slate, size = 2) +
  geom_segment(aes(x = 6, y = 0, xend = 6, yend = 7.5), linetype  = "dashed", color = met_slate, size = 1) +
  geom_segment(aes(x = 11, y = 0, xend = 11, yend = 7.5), linetype  = "dashed", color = met_slate, size = 1)
```
]

---
# Mutual Gains from Trade

.less-right[
## Shortage

P .mono[=].mono[$]4.50 .mono[==>] TS .mono[=] .mono[$]36.

]

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5.4, dev = "svg"}
ggplot() +
  geom_step(data = step_supply, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  annotate("text", label = "S", x = 12.5, y = 8, color = purple, family = "Fira Sans", size = 9) +
  geom_step(data = step_demand, mapping = aes(x = x, y = mv), color = red_pink, size = 1) +
  annotate("text", label = "D", x = 12.5, y = 4, color = red_pink, family = "Fira Sans", size = 9) +
  # geom_step(data = step_supply_tax, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  # annotate("text", label = expression(S[Tax]), x = 13, y = 10, color = purple, family = "Fira Sans", size = 9) +
  labs(x = "Q", y = "P") +
  scale_x_continuous(limits = c(0, 14), expand = c(0, 0), breaks = seq(0, 12, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  geom_hline(yintercept = 4.5, linetype  = "dashed", color = met_slate, size = 1) +
  geom_point(aes(x = 11, y = 4.5), color = met_slate, size = 2) +
  geom_point(aes(x = 6, y = 4.5), color = met_slate, size = 2) +
  geom_vline(xintercept = 6, linetype  = "dashed", color = met_slate, size = 1) +
  geom_segment(aes(x = 11, y = 0, xend = 11, yend = 4.5), linetype  = "dashed", color = met_slate, size = 1) +
  geom_stepribbon(data = step_demand, aes(ymax = ifelse(x < 6, mv, 7.5), ymin = 7.5, x = x), fill = green, alpha = 0.2, linetype = "blank") +
  annotate("text", label = "TS", x = 2, y = 6, color = green, family = "Fira Sans", size = 9) +
  geom_stepribbon(data = step_supply, aes(ymax = ifelse(x < 6, mv, 7.5), ymin = 7.5, x = x), fill = green, alpha = 0.2, linetype = "blank")
```
]

---
# Mutual Gains from Trade

.less-right[
## Equilibrium

P .mono[=].mono[$]6 .mono[==>] TS .mono[=] .mono[$]40.

]

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5.4, dev = "svg"}
ggplot() +
  geom_step(data = step_supply, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  annotate("text", label = "S", x = 12.5, y = 8, color = purple, family = "Fira Sans", size = 9) +
  geom_step(data = step_demand, mapping = aes(x = x, y = mv), color = red_pink, size = 1) +
  annotate("text", label = "D", x = 12.5, y = 4, color = red_pink, family = "Fira Sans", size = 9) +
  # geom_step(data = step_supply_tax, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  # annotate("text", label = expression(S[Tax]), x = 13, y = 10, color = purple, family = "Fira Sans", size = 9) +
  labs(x = "Q", y = "P") +
  scale_x_continuous(limits = c(0, 14), expand = c(0, 0), breaks = seq(0, 12, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  geom_stepribbon(data = step_demand, aes(ymax = ifelse(x < 9, mv, 6), ymin = 6, x = x), fill = green, alpha = 0.2, linetype = "blank") +
  annotate("text", label = "TS", x = 1.5, y = 7, color = green, family = "Fira Sans", size = 9) +
  geom_stepribbon(data = step_supply, aes(ymax = ifelse(x < 9, mv, 6), ymin = 6, x = x), fill = green, alpha = 0.2, linetype = "blank") +
  geom_point(aes(x = 9, y = 6), color = met_slate, size = 2) +
  geom_segment(aes(x = 9, y = 0, xend = 9, yend = 6), linetype  = "dashed", color = met_slate, size = 1) + 
  geom_segment(aes(x = 0, y = 6, xend = 9, yend = 6), linetype  = "dashed", color = met_slate, size = 1)
```
]

---
# Mutual Gains from Trade

.less-right[
## Intervention

**Q:** True or false? The government can increase TS by mandating 12 trades.

> **A.** True

> **B.** False

]

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5.4, dev = "svg"}
ggplot() +
  geom_step(data = step_supply, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  annotate("text", label = "S", x = 12.5, y = 8, color = purple, family = "Fira Sans", size = 9) +
  geom_step(data = step_demand, mapping = aes(x = x, y = mv), color = red_pink, size = 1) +
  annotate("text", label = "D", x = 12.5, y = 4, color = red_pink, family = "Fira Sans", size = 9) +
  # geom_step(data = step_supply_tax, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  # annotate("text", label = expression(S[Tax]), x = 13, y = 10, color = purple, family = "Fira Sans", size = 9) +
  labs(x = "Q", y = "P") +
  scale_x_continuous(limits = c(0, 14), expand = c(0, 0), breaks = seq(0, 12, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  #geom_stepribbon(data = step_demand, aes(ymax = ifelse(x < 9, mv, 6), ymin = 6, x = x), fill = green, alpha = 0.2, linetype = "blank") +
  #annotate("text", label = "TS", x = 2, y = 6, color = green, family = "Fira Sans", size = 9) +
  #geom_stepribbon(data = step_supply, aes(ymax = ifelse(x < 9, mv, 6), ymin = 6, x = x), fill = green, alpha = 0.2, linetype = "blank") +
  geom_vline(xintercept = 12, linetype  = "dashed", color = met_slate, size = 1)
```
]

---
count: false
# Mutual Gains from Trade

.less-right[
## Intervention

**Q:** True or false? The government can increase TS by mandating 12 trades.

> **A.** True

> .pink[**B.** False]

]

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5.4, dev = "svg"}
ggplot() +
  geom_step(data = step_supply, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  annotate("text", label = "S", x = 12.5, y = 8, color = purple, family = "Fira Sans", size = 9) +
  geom_step(data = step_demand, mapping = aes(x = x, y = mv), color = red_pink, size = 1) +
  annotate("text", label = "D", x = 12.5, y = 4, color = red_pink, family = "Fira Sans", size = 9) +
  # geom_step(data = step_supply_tax, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  # annotate("text", label = expression(S[Tax]), x = 13, y = 10, color = purple, family = "Fira Sans", size = 9) +
  labs(x = "Q", y = "P") +
  scale_x_continuous(limits = c(0, 14), expand = c(0, 0), breaks = seq(0, 12, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  #geom_stepribbon(data = step_demand, aes(ymax = ifelse(x < 9, mv, 6), ymin = 6, x = x), fill = green, alpha = 0.2, linetype = "blank") +
  #annotate("text", label = "TS", x = 2, y = 6, color = green, family = "Fira Sans", size = 9) +
  #geom_stepribbon(data = step_supply, aes(ymax = ifelse(x < 9, mv, 6), ymin = 6, x = x), fill = green, alpha = 0.2, linetype = "blank") +
  geom_vline(xintercept = 12, linetype  = "dashed", color = met_slate, size = 1)
```
]

---
count: false
# Mutual Gains from Trade

.less-right[
## Intervention

**Q:** True or false? The government can increase TS by mandating 12 trades.

> **A.** True

> .pink[**B.** False]

TS would decrease from .mono[$]40 to .mono[$]34

]

.more-left[
```{r, echo = FALSE, fig.height = 5, fig.width = 5.4, dev = "svg"}
ggplot() +
  geom_step(data = step_supply, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  annotate("text", label = "S", x = 12.5, y = 8, color = purple, family = "Fira Sans", size = 9) +
  geom_step(data = step_demand, mapping = aes(x = x, y = mv), color = red_pink, size = 1) +
  annotate("text", label = "D", x = 12.5, y = 4, color = red_pink, family = "Fira Sans", size = 9) +
  # geom_step(data = step_supply_tax, mapping = aes(x = x, y = mv), color = purple, size = 1) +
  # annotate("text", label = expression(S[Tax]), x = 13, y = 10, color = purple, family = "Fira Sans", size = 9) +
  labs(x = "Q", y = "P") +
  scale_x_continuous(limits = c(0, 14), expand = c(0, 0), breaks = seq(0, 12, 1)) +
  scale_y_continuous(limits = c(0, 10.5), expand = c(0, 0), breaks = seq(0, 10, 1)) +
  theme_market +
  geom_stepribbon(data = step_demand, aes(ymax = ifelse(x < 9, mv, 6), ymin = 6, x = x), fill = green, alpha = 0.2, linetype = "blank") +
  annotate("text", label = "+TS", x = 2, y = 6, color = green, family = "Fira Sans", size = 9) +
  geom_stepribbon(data = step_supply, aes(ymax = ifelse(x < 9, mv, 6), ymin = 6, x = x), fill = green, alpha = 0.2, linetype = "blank") +
  
  geom_stepribbon(data = step_demand, aes(ymax = ifelse(x > 9, mv, 6), ymin = 6, x = x), fill = orange, alpha = 0.2, linetype = "blank") +
  annotate("text", label = "-TS", x = 11, y = 6, color = orange, family = "Fira Sans", size = 9) +
  geom_stepribbon(data = step_supply, aes(ymax = ifelse(x > 9, mv, 6), ymin = 6, x = x), fill = orange, alpha = 0.2, linetype = "blank") +
  
  geom_vline(xintercept = 12, linetype  = "dashed", color = met_slate, size = 1)
```
]

---
class: clear-slide

**Q:** What is the source of mutual benefits from voluntary trade?

--

**A:** .pink[Differences in marginal values.]

--

**Example:** You own a pair of sneakers. You value them at .mono[$]20. I value them at .mono[$]50. Could we benefit from trade?

--

- **A:** Yes! Marginal value differential of .mono[$]30 .mono[=] .mono[$]30 in mutual benefits!.super[.pink[<span>&#8224;</span>]]

.footnote[.pink[<span>&#8224;</span>]: The distribution of the benefits depends on the price we negotiate.]

---
class: clear-slide

**Q:** What role do "middlemen".super[.pink[<span>&#8224;</span>]] play?

.footnote[.pink[<span>&#8224;</span>]: *e.g.,* real estate agents, stock brokers, grocers, retailers, *etc.*]
--

**A:** .pink[They play match-maker in exchange for part of the marginal-value differential.]

---
# Efficiency

## Definition 1 (market-specific)

> The outcome that maximizes mutual benefits from voluntary trade (*i.e.,* total surplus).

--

## Definition 2 (more general)

> The absence of waste.

--

## Definition 3 (.pink[*Pareto Efficiency*])

> The state in which no person can be made better off without harming someone else.

---
class: clear-slide

**Podcast Question:** What problem(s) did Alaska's food bank have before *Feeding America* established an auction market?

> **A.** It received too few pickles.

> **B.** It received so much fresh produce that it would rot before people could eat it.

> **C.** It received too many potatoes.

> **D.** It received too few potatoes.

> **E.** It received too many pickles.

---
count: false
class: clear-slide

**Podcast Question:** What problem(s) did Alaska's food bank have before *Feeding America* established an auction market?

> **A.** It received too few pickles.

> **B.** It received so much fresh produce that it would rot before people could eat it.

> **C.** It received too many potatoes.

> **D.** It received too few potatoes.

> .pink[**E.** It received too many pickles.]

---
class: clear-slide

**Podcast Question:** Which food banks received the most "fake money" to use in the auction market?

> **A.** Those in relatively affluent areas.

> **B.** Those in relatively poor areas.

> **C.** Those with the most donors.

> **D.** Those most willing to pay.

> **E.** Those with the most food stockpiled.

---
count: false
class: clear-slide

**Podcast Question:** Which food banks received the most "fake money" to use in the auction market?

> **A.** Those in relatively affluent areas.

> .pink[**B.** Those in relatively poor areas.]

> **C.** Those with the most donors.

> **D.** Those most willing to pay.

> **E.** Those with the most food stockpiled.

---
class: clear-slide

**Q:** What problems did the *Feeding America* auction market solve?

--

1. .pink[Local information problem:] Food bank directors knew exactly what they wanted, but *Feeding America* executives had to make guesses.

2. .pink[Shortages and surpluses:] Food banks had insufficient quantities of sought-after items but abundant quantities of unwanted items.

--

**Q:** Are food banks and their clients better off now than they were before? 

--

**A:** Yes. 

- Less food waste .mono[-->] more food for people in need.

- .pink[*Pareto improving:*] More mutually beneficial transactions..super[.pink[<span>&#8224;</span>]]

.footnote[.pink[<span>&#8224;</span>]: *Pareto improvement* .mono[=] A transaction that makes at least one person better off without harming others.]

---
# Efficient Markets

## Conditions

1. Well-defined property rights.

2. No market power.

3. *Symmetric* information.

--

## Result 

> If all three conditions hold, then an *unregulated* market maximizes total surplus.

---
# Efficient Markets

## Result 

> If all three conditions hold, then an *unregulated* market maximizes total surplus.

## Implications

1. When all three conditions hold, government intervention in a market can generate inefficient outcomes.

2. When at least one condition fails, an *unregulated* market can generate inefficient outcomes. 

    - Well-designed government intervention *can* redress inefficiencies.

---
# Efficient Markets

**Q:** What kinds of markets are most likely to yield efficient outcomes?

--

**A:** Markets that feature 

1. Transactions that don't hurt or help those outside the transaction.

2. Many buyers and sellers.

3. Goods with easily observable quality.

--

Examples?

---
class: inverse, middle

# Midterm 1

---
# Midterm Topics

.green[Anything from the lectures, discussions, or podcasts] .hi-green[is fair game!]

1. What is Economics?

2. Motivating the Economic Problem

3. Consumer Theory I

4. Consumer Theory II

5. The Market Mechanism

6. Demand and Supply

---
# Midterm Structure

## Multiple Choice

- 50 questions
- 1 point per question

---
# Midterm Protocol

## Materials

- Writing utensil
- 3-inch-by-5-inch note card
- Basic or scientific calculator (no graphing or programming capabilities)
- .hi[Nothing else]

--

## Procedure

- **Randomized** seating chart
- 80 minutes from *"you may begin"* to *"pencils down"*
- First 30 minutes: .hi[quiet period] (no questions, no getting up)
- Last 50 minutes: ask lots of questions
- Show your UO ID card as you turn-in your exam

---
# Midterm Preparation

1. Lecture slides

2. Discussion worksheets

3. Practice midterm questions

    - No solutions posted, but you can ask about the questions in office hours.

4. Extended office hours 

    - M 14:00-15:00 and T 14:30-17:30 in **412 PLC**
    
5. Textbook (additional practice problems)

6. Prepare your note card